### 2019-07-29 tex/Drafts/Draft-2.tex
Introduction revisions

  \documentclass{amsart}   
  \usepackage{../RepStyle} 
  % Colors links
  \hypersetup{
  colorlinks,
  citecolor=blue,
  linkcolor=orange!50!red,
  urlcolor=blue}

  \usepackage{../NatMacros}

  % Emulates \subsection* but doesn't add to ToC (I don't know why amsart is weird 
  \newcommand{\fakesubsection}[1]{
    \vspace{7pt}
    \noindent \textbf{#1.}
  }

  \makeatletter
  %Table of Contents
  \setcounter{tocdepth}{3}

  % Add bold to \section titles in ToC and remove . after numbers
  \renewcommand{\tocsection}[3]{%
    \indentlabel{\@ifnotempty{#2}{\bfseries\ignorespaces#1 #2\quad}}\bfseries#3}
  % Remove . after numbers in \subsection
  \renewcommand{\tocsubsection}[3]{%
    \indentlabel{\@ifnotempty{#2}{\ignorespaces#1 #2\quad}}#3}
  \let\tocsubsubsection\tocsubsection% Update for \subsubsection
  %...

  \newcommand\@dotsep{4.5}
  \def\@ocline#1#2#3#4#5#6#7{\relax
    \ifnum #1>\c@tocdepth % then omit
    \else
      \par \addpenalty\@secpenalty\addvspace{#2}%
      \begingroup \hyphenpenalty\@M
      \@ifempty{#4}{%
        \@tempdima\csname r@tocindent\number#1\endcsname\relax
      }{%
        \@tempdima#4\relax
      }%
      \parindent\z@ \leftskip#3\relax \advance\leftskip\@tempdima\relax
      \rightskip\@pnumwidth plus1em \parfillskip-\@pnumwidth
      #5\leavevmode\hskip-\@tempdima{#6}\nobreak
      \leaders\hbox{$\m@th\mkern \@dotsep mu\hbox{.}\mkern \@dotsep mu$}\hfill
      \nobreak
      \hbox to\@pnumwidth{\@tocpagenum{\ifnum#1=1\bfseries\fi#7}}\par% <-- \bfseries for \section page
      \nobreak
      \endgroup
    \fi}

  \def\l@subsection{\@tocline{2}{0pt}{2.5pc}{5pc}{}}
  \makeatother

  \begin{document}

  \title[Some Graphical Realizations of Two-Row Specht Modules of Hecke Algebras]{Some Graphical Realizations of Two-Row Specht Modules of Iwahori--Hecke Algebras of the Symmetric Group}
  \author[Miles Johnson and Natalie Stewart]{Miles Johnson and Natalie Stewart\\
    Mentor : Oron Propp\\
  Project Suggested by : Roman Bezrukavnikov\\ \; \\
  July 31, 2019
  }

   \begin{titlepage}
    \maketitle
    \begin{abstract}
      We consider the Iwahori--Hecke algebra of the symmetric group on $2n + r$ letters with parameter $q$.
      Let $e$ be the smallest positive integer such that the $q$-number $\brk{e}_q = 0$, or set $e = \infty$ if none exist.
      We modify Khovanov's crossingless matchings to include $2n$ ``nodes'' and $r$ ``anchors,'' and prove in the case $e > n + r + 1$ that the associated module is isomorphic to the Specht module $S^{(n+r,n)}$ which corresponds to the partition $(n + r,n) \vdash 2n + r$.
      We then give heuristics in support of the general case, including explicit composition series for $e = n + r + 1$ and for $2n + r \leq 7$. 
      Lastly, when $e = 5$, we prove an isomorphism between the irreducible quotient $D^{(n+r,n)}$ with $r \leq 3$ and some subrepresentations of Jordan--Shor's Fibonacci representation.
      We provide explicit transition matrices between this representation and the crossingless matchings representation for $2n + r \leq 6$.
    \end{abstract}

  \begingroup
  \hypersetup{linkcolor=black} % no colored hyperlinks in toc
  \tableofcontents
  \endgroup

  \end{titlepage}

\section{Introduction} 
  Let $S_{2n+r}$ be the symmetric group on $2n+r$ letters with $2n + r \geq 2$, let $\SH := \SH_{k,q}(S_{2n+r})$ be the corresponding Hecke algebra over a field $k$ with parameter $q \in k^\times$ having a fixed square root $q^{1/2}$, and let $\cbr{T_i}$ be the simple reflections generating $\SH$.

  Let $\brk{m}_q = 1 + q + \dots + q^{m-1}$ be the $q$-number of $m$. 
  Let $e$ be the smallest positive integer such that $\brk{e}_q = 0$, and set $e = \infty$ if no such integer exists.
  Note that either $q = 1$ and $e$ is the characteristic of $k$ (with $0$ replaced by $\infty$), or $q \neq 1$ and $q$ is a primitive $e$th root of unity.

  \vspace{5pt}
  When $q = 1$, the Hecke algebra $\SH$ is isomorphic to the group algebra $k\brk{S_{2n+r}}$;
  hence the representation theory of $\SH$ generalizes the representation theory of the symmetric group.
  The Hecke algebra is also well known to be connected to the representation theory of the general linear group over a finite field.\cite{Mathas-book}

  It is a classical result that $\SH$ is semisimple precisely when $e > 2n + r$, in which case the irreducible representations of $\SH$ are given by \emph{Specht modules} $S^\lambda$, which are indexed by the partitions $\lambda$ of $2n + r$.
  Further, $\SH$ admits a cellular basis with cell modules given by $S^\lambda$.
  In particular, these admit quotients $D^\lambda$ such that the set $\cbr{D^\lambda \mid D^\lambda \neq 0, \lambda \vdash n}$ is a complete set of pairwise-nonisomorphic irreducible $\SH$-modules.
  This set is indexed by the partitions $\lambda \vdash n$ which are $e$-restricted.\cite{Murphy1,Murphy2}

  These representations $D^\lambda$ have explicit constructions, but many of their properties are unknown.
  For instance, the dimension of these modules is unknown outside of some special cases.\cite{Mathas-book}
  However, there does exist an algorithm due to Lascoux--Leclerc--Thibon--Ariki which computes the decomposition matrices of the Iwahori--Hecke Algebra $\SH_{\CC,q}(S_{2n + r})$ for $q$ an $e$th root of unity.\cite{LLT,Ariki}

  The cellular basis for $S^\lambda$ and associated construction for $D^\lambda$ tend to be complicated and often computationally intractable.
  We aim to give simple graphical realizations of $S^\lambda$ and $D^\lambda$ in some cases that $\lambda = (n+r,n)$.
  These realizations expose the structure of irreducibility and branching in an intuitive and computationally simple way.

  \vspace{5pt}
  Note that we follow the convention of Murphy and Kleshchev, which is dual to the conventions of Dipper, James, and Mathas;
  one may translate our results to the latter convention by transposing all partitions.\cite{Kleshchev,Mathas-book,Murphy1}

  Throughout this paper, we analyze the \emph{two-row partitions} $(n+r,n) \vdash 2n + r$ and their corresponding modules $S^{(n+r,n)}$ and $D^{(n+r,n)}$.

  \fakesubsection{Crossingless Matchings}
  The following definition modifies the crossingless matchings defined by Khovanov.\cite{Khovanov}

  \begin{definition}
  Define a \emph{crossingless matching on $2n + r$ nodes and $r$ anchors} to be an isotopy class of $n + r$ nonintersecting paths in the slice $\RR \times \brk{0,1}$ connecting $2n + r$ disctinct points of $\RR \times \cbr{0}$ and $r$ points of $\RR \times \cbr{1}$ such that none of the latter points are conected. 
  Let $M_{2n + r}^r$ have basis given by these matchings.
  This is illustrated in Figure \ref{S5 Basis}.

  Let the length of an arc $(i,j)$ be $j - i + 1$.

  We endow $M_{2n + r}^r$ with an action by specifying $(1 + T_i)w_j$ for some basis element $w_j$ of $M_{2n + r}$.
  We do so by concatenating in ``vertical lines'' below each point other than the $i$th and $i+1$st, concatenating paths between the $i$th and $i+1$st points as well as points below them, removing any ``loops'' this forms, and taking isotopy to some matching $w_l$ if possible;
  if this is not possible, then there are anchors at $i,i+1$ and we set $(1 + T_i)w_j := 0$;
  if this is possible and there is a ``loop,'' set $(1 + T_i)w_j := (1 + q)w_j$ and otherwise set $(1 + T_i)w_j := q^{1/2}w_l$.
  This action is illustrated in Figure \ref{Action}, and we verify that this is well-defined in Appendix \ref{Cross Relations}.
  \end{definition}
  \begin{figure} 
    \def\cbasisspacing{5mm}
    \begin{adjustbox}{width=\textwidth}
    $\cbr{
      \begin{gathered}
        \GeneralizedMatching{5}{1/2, 3/4}{1}{5/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{1/4, 2/3}{1}{5/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{1/2, 4/5}{1}{3/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{2/3, 4/5}{1}{1/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{2/5, 3/4}{1}{1/1}{3/4}, \hspace{\cbasisspacing}
       \end{gathered}}$ 
     \end{adjustbox}
       \caption{The basis for $M_5^1$.}
    \label{S5 Basis}
   \end{figure}

  \begin{figure}
    \begin{adjustbox}{width=\textwidth}
      \begin{tabular}{l l}
        \GeneralizedAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{2}{1/4, 2/3}{5/1,6/2}{(1+q)}
        \hspace{20pt}
        &
        \GeneralizedAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{3}{1/2, 3/4}{5/1,6/2}{q^{1/2}}\\
        \GeneralizedAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{4}{2/3, 4/5}{1/1,6/2}{q^{1/2}}
        &
        \GeneralizedZeroAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{5}
      \end{tabular}
  \end{adjustbox}

    \caption{Illustration of the actions $(1 + T_i)w_{\abs{M^2_6}}$.
      In general, we act by deleting loops, isotoping onto a new crossingless matching, and scaling by either $q^{1/2}$, $(q + 1)$, or 0.
    }
    \label{Action}
  \end{figure}

  Note that the representations $M_{0 + r}^r$ and $S^{(r)}$ are isomorphic to the sign representation.
  In fact, we will prove that $M_{2n + r}^r \simeq S^{(n+r,n)}$ whenever $e > n + r + 1$.
  Additionally, we will prove that $M_{2n + r}^r$ is irreducible everywhere that $e > n$ and $S^{(n+r,n)}$ is irreducible.
  We will conjecture that $M_{2n + r}^r \simeq S^{(n+r,n)}$ with $e$ unrestricted as well.

  This proves a graphical characterization of $S^\lambda$ in many cases, and hence $D^\lambda$ in the cases where $S^\lambda \simeq D^\lambda$.
  However, the representation theory of $\SH$ is tied to the modules $D^\lambda$ even in cases where $S^\lambda$ is not irreducible;
  we will present a characterization of the module $D^{(n+r,n)}$ when $e = 5$ and $r \leq 3$.

  \def\vara{\alpha_1}
  \def\varb{\alpha_2}
  \def\varc{\varepsilon_1}
  \def\vard{\delta}
  \def\vare{\varepsilon_2}
  \def\vs{\texttt{*}}
  \def\vp{\texttt{0}}
  \fakesubsection{Fibonacci Representation}
  Now suppose that $e = 5$ and $k$ contains the algebraic number $\sqrt{q + q^4}$ for reasons which will be apparent soon.
  For convenience set $m := 2n + 4$.
  The following is a modification of Shor--Jordan's Fibonacci representation of the braid group.\cite{Shor}

  \begin{definition}
  Let $V^m$ be the $k$-vector space with basis given by the strings $\cbr{\vs,\vp}^{n+1}$ such that the bit $\vs$ never appears twice consecutively. 
  We will refer to $V^m$ as the \emph{Fibonacci representation} and suppress the superscript whenever it is clear from context.

  We wish to endow this with a $\SH$-action which acts on a basis vector only dependent on bits $i,i+1,i+2$, sending each basis vector to a combination of the other basis vectors having the same bits $1,\dots,i,i+2,\dots,n+1$ as follows:

  \begin{equation} 
    \begin{split}
      T_1 \, \prn{\vs\vp\vp} &:= \vara\prn{\vs\vp\vp}\\
      T_1 \, \prn{\vp\vp\vs} &:= \vara\prn{\vp\vp\vs}\\
      T_1 \, \prn{\vs\vp\vs} &:= \varb\prn{\vs\vp\vs}\\
      T_1 \, \prn{\vp\vs\vp} &:= \varc\prn{\vp\vs\vp} + \vard\prn{\vp\vp\vp}\\
      T_1 \, \prn{\vp\vp\vp} &:= \vard\prn{\vp\vs\vp} + \vare\prn{\vp\vp\vp}
  \end{split} \label{Fib Action} 
  \end{equation}
  for constants
  \begin{equation}
    \begin{split}
    \tau  &:= q + q^4\\
    \vara &:= -1\\
    \varb &:= q\\
    \varc &:= \tau(q\tau - 1)\\
    \vard &:= \tau^{3/2}(q + 1)\\
    \vare &:= \tau(q-\tau)
  \end{split} \label{Fib Constants} 
  \end{equation}
  with $T_i$ acting similarly on the substring $i,i+1,i+2$.
  We will verify that this is a representation of $\SH$ in Appendix \ref{Fib Relations}
  \end{definition}

  This contains four subrepresentations spanned by strings with beginning and ending with specified bits.
  Label the subrepresentation spanned by strings $(\vs\dots\vs)$ by $V_{\vs\vs}$, $V_{\vs\vp}$, $V_{\vp\vs}$, and $V_{\vp\vp}$;
  these bits are preserved by the action \eqref{Fib Action}.
  We will show that these are isomorphic to particular $D^\lambda$.

  \fakesubsection{Overview of Paper}

  In Section \ref{Specht Modules Section} we give corollaries to standard theorems concerning Specht modules.
  First, James-Mathas provides a sharp characterization of the irreducibility $S^\lambda$ for $\lambda \vdash 2n + r$ which is $e$-regular, called the \emph{Carter criterion}.\cite{Mathas-book}
  We specialize this to the case that $\lambda = (n+r,n)$ to give a combinatorial condition for irreducibility of $S^{(n+r,n)}$.
  We note that this irreducibility depends only on $e$ when $e > n$;
  otherwise it depends on both $e$ and the characteristic of the field.
  Further, we use Kleshchev--Brundan's modular branching rules to prove our first significant statement: if $S^\lambda \simeq D^\lambda$ and $e > n$, then a particular length-2 composition series uniquely determines $\lambda$;
  further, an irreducible restriction to $D^{(n,n-1)}$ determines $\lambda$ as well.\cite{Kleshchev, Brundan}

  \def\thmcolor{black!60!orange}
  \vspace{5pt}
  In Section \ref{Crossingless Matchings Section}, we first prove that every basis vector in $M$ is cyclic, then that $M$ contains no sign subrepresentation.
  Then, we prove the following theorem.
  \begingroup
  \hypersetup{linkcolor=\thmcolor} % no orange theorem label
  \begin{customthm}{\ref{Irreducibility Theorem}}
    Suppose $e > n$ and $S^{(n+r,n)}$ is irreducible.
    Then $M_{2n + r}^r$ is irreducible.
  \end{customthm}
  \endgroup
  Following this, we prove the existence of a particular filtration with factors given by other crossingless matchings representations;
  using irreducibility, this becomes a composition series, and an inductive argument combined with the branching of Section \ref{Specht Modules Section} allow us to prove the following:
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{Correspondence Theorem}}
    Suppose $e > n + r + 1$.
    Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.
  \end{customthm}
  \endgroup
  Last, we finish the section by proving more general statements concerning sign subrepresentations of $M$, which contribute to conjectures later in the paper.{\color{orange} Fill in with specific information later.}

  \vspace{5pt}
  In Section \ref{Fibonacci Section}, we begin by establishing isomorphisms as in \ref{Fibonacci Theorem} for the subrepresentations of $V^2$, as well as irreducibility of $V^3_{\vs\vp}$.
  We then use these cases to prove that $V^m_{\vs\vp}$ is irreducibile for all $m$, which implies that that $V^m_{\vs\vs}$ is irreducible.
  From this, we inductively prove the following theorem:
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{Fibonacci Theorem}}
    The following isomorphisms characterize the subrepresentations of the Fibonacci representation:
    \begin{align*}  
       V^{2n}_{\vs\vs} &\simeq D^{(n,n)}\\ 
       V^{2n-1}_{\vs\vs} &\simeq D^{(n+1,n-2)}\\
       V^{2n}_{\vs\vp} &\simeq D^{(n+1,n-1)}\\
       V^{2n-1}_{\vs\vp} &\simeq D^{(n,n-1)}.
    \end{align*}
  \end{customthm}
  \endgroup
  {\color{orange} Overview of Conjecture and Empirics goes here.}

  % Acknowledgements should be given section-level heading in toc, subsection-level in text.
  \addcontentsline{toc}{section}{\textbf{Acknowledgements}}
  \fakesubsection{Acknowledgements}
  The authors thank Prof. Roman Bezrukavnikov for suggesting this project, as well as Dr. Slava Gerovitch for organizing the SPUR+ program.
  We would also like to thank Profs. David Jerison and Ankur Moitra for their role in SPUR+ as well as general advice.
  We would also thank Professor Alexander Kleshchev for our helpful conversations concerning branching theorems.
  Lastly, we would like to express our gratitude to our mentor Oron Propp for his help and advice in both acquiring background knowledge and in executing the mathematics in this paper;
  this project would not be possible without him.

\section{Preliminaries on Specht Modules}\label{Specht Modules Section}
  For this section and the rest of the paper, assume $n > 0$.

  In the following section, we cite a theorem of James-Mathas which precisely characterizes the irreducibility of $S^\lambda$ in the case that $\lambda$ is $e$-regular, and we specialize this result to the case of two-row Specht modules.
  This falls into two cases: either $e > n$, where $S^{(n+r,n)}$ is irreducible iff $e \nmid r+2,\dots,n+r+1$, or $e \leq n$, where the irreducibility of $S^{(n+r,n)}$ is complicated and depends on the characteristic of $k$.
  We will focus primarily on the former case.

  Following this, we cite the branching theorems of Kleshchev-Brundan, which allow us to fully characterize the Socle of $\Res S^{\lambda}$.
  This and some combinatorial arguments allow us to come to the main result of this section, which may loosely be stated as follows:
  the composition series of $D^{(n+r,n)}$ is restrictive enough in many cases with $S^{(n+r,n)}$ irreducible and $e > n$ case that such a composition series uniquely characterizes irreducibles.
  This will be immensely useful later for characterizing $M$ via Specht modules, as it will include all cases with $e > n + r + 1$.

  \subsection{Irreducibility of $S^\lambda$}
  Let $k$ have characteristic $\ell$;
  then, set
  \[
    p := \begin{cases}
      \ell & \ell > 0\\
      \infty & \ell = 0
    \end{cases}.
  \]
  Note that $p = e$ when $q = 1$.
  For $h$ a natural number, let $\nu_p(h)$ be the $p$-adic evaluation of $h$.
  As a convention, set $\nu_\infty(h) = 0$ for all $h$.
  Define the function $\nu_{e,p} : \NN \rightarrow \cbr{-1} \cup \NN$ by
  \[
    \nu_{e,p}(h) := \begin{cases}
      \nu_p(h) & e \mid h\\
      -1 & e \nmid h
    \end{cases}.
  \]
  Lastly, let $h_{ab}^\lambda$ be the hook length of node $(a,b)$ in $\brk \lambda$.
  With this language, we may express the following theorem, part (ii) of which is named the \emph{carter criterion}, due to James-Mathas.\cite{Mathas-book}
  \begin{theorem}[James-Mathas]
    The following are equivalent:
    \begin{enumerate}[label={(\roman*)}]
      \item $S^{\lambda} \simeq D^{\lambda}$.
      \item $\lambda$ is $e$-regular and $S^\lambda$ is irreducible.
      \item $\nu_{e,p}\prn{h_{ab}^\lambda} = \nu_{e,p}\prn{h_{ac}^\lambda}$ for all nodes $(a,b)$ and $(a,c)$ in $\brk \lambda$.
    \qed\end{enumerate}
  \end{theorem}
  This result gives information solely on $e$-regular partitions, and the general irreducibility of $S^\lambda$ away from $p=2$ is not well understood.
  We will henceforth specialize slightly to the case that $(n+r,n)$ is $e$-restricted.

  \begin{corollary}\label{S irreducibility}
    If $r = 0$, assume $e > 2$.
    \begin{enumerate}[label={(\roman*)}]
      \item Suppose $e > n$.
        Then, $S^{(n+r,n)}$ is irreducible iff $e \nmid l$ for all $r + 2 \leq l \leq n + r + 1$.
      \item Suppose $e \leq n$.
        Then, $S^{(n+r,n)}$ is irreducible iff $e | r + 1$ and $\nu_{e,p}(h_{ab}^\lambda)$ acquires exactly two values.
    \end{enumerate}  
  \end{corollary}
  In the case that $\lambda = (n+r,n)$ and $e \nmid r+2,r+3,\dots,n+r+1$, say that $\lambda$ is \emph{$e$-top-indivisible}.
  Since no multiples of $e$ are in an interval of length $n$ in $\NN$, we have $e > n$, so this is precisely case (i) above.
  \begin{proof}
    The beginning condition implies that $\lambda$ is $e$-regular, which we will use below.

    \textbf{(i)}
    Note that $\nu_p(l) \neq -1$ for all $l$ and only hook lengths in the top row may vanish mod $e$;
    hence we may equivalently prove that $e$ divides no hook lengths in the leftmost $n$ columns of the second column.
    These hook lengths are precisely $r + 2,\dots,n+r+1$.

    \textbf{(ii)}
    Note that we have $\nu_{e,p}\prn{h^\lambda_{n-e+1,2}} \neq -1$.
    Suppose that $e \nmid r + 1$.
    Then, \[\nu_{e,p}\prn{h^\lambda_{n-e+1,1}} = \nu_{e,p}\prn{h^\lambda_{n-e,2} + r + 1} = -1,\]
    giving $S^{(n+r,n)}$ reducible.

    \def\corsecsep{5pt}
    \vspace{\corsecsep}
    Now, suppose that $\nu$ acquires at least three values, and \[0 \leq \nu_{e,p}(h^{\lambda}_{ab}) < \nu_{e,p}\prn{h^{\lambda}_{a'b'}}\] and $S^{(n+r,n)}$ is irreducible.
    If $b=b'$ then we have reducibility, so assume $b \neq b'$.
    Further, if $a = a'$, then we may replace $a$ with the other column;
    hence we may assume WLOG that $a \neq a'$ as well.

    Note that $p$-adic valuation is monotonic, so $h^\lambda_{ab} < h^\lambda_{a'b'}$.
    If $(a,b)$ is in the rightmost $r$ columns, then we may $0 \leq \nu_{e,p}\prn{h^\lambda_{a''b''}} \leq \nu_{e,p}\prn{h^\lambda_{ab}}$ for some $(a'',b'')$ in the rightmost $n$ columns;
    this has the same hook length as a cell in the leftmost $n$ columns and bottom row, so we may assume $(a,b)$ is not in the rightmost $r$ columns.
    Similar logic allows us to assume $(a',b')$ is in the leftmost $n$ columns as well.

    If $b < b'$, then $\nu_p(h_{ab}) = \nu_p(h_{ab'}) < \nu(h_{a'b'})$ while $h_{ab'} > h_{a'b'}$, a contradiction.
    If $b > b'$, then $\nu_p(h_{ab}) = \nu_p(h_{ab'})$ and $h_{ab} < h_{ab'}$, so there is some $c < b$ with $\nu_{e,b}(h_{ab}) = \nu_{e,p}(h_{ac})$; we may replace $b$ with $c$, and repeat until $b = b'$ to reach contradiction.

    \vspace{\corsecsep}
    Finally, assume $e | r + 1$ and $\nu_{e,p}$ acquires two values.
    Since $e | r + 1$, $e \mid h_{a1}^\lambda$ iff $e \mid h_{a2}$;
    since $\nu_{e,p}$ acquires only two values, this proves that $\nu_{e,p}(h_{a1}) = \nu_{e,p}(h_{a2})$ across these rows, giving the lemma.
  \end{proof}

  From condition (ii) we see that irreducibiliy at $e > n$ is not dependent on $p$, and we may cover many modular cases without reference to the characteristic of $k$.
  We will finish our discussion of irreducibilty of $S^\lambda$ via the following special cases of (ii) above.
  \begin{corollary}
    If $r = 0$, assume $e > 2$.
    Suppose $e \leq n$.
    \begin{enumerate}[label={(\roman*)}]
      \item Suppose $p > n + r + 1$. Then, $S^{(n + r,n)}$ is irreducible iff $e | r + 1$.
      \item Suppose $p = 2$. Then, $S^{(n+r,n)}$ is not irreducible.
    \end{enumerate}
  \end{corollary}
  \begin{proof}
    (i) is clear.
    (ii) is given by noting that $e + r + 1 \geq 2e$;
    then, $h_{n-e+1,1} - h_{n-e+1,2} \geq 1$, giving reducibility.
  \end{proof}
  \subsection{Branching}
  In this section as well as later sections, we will consider the restruction of representations of $\SH$ to particular subalgebras isomorphic to $\SH_{k,q}\prn{S_{2n + r - 1}}$.
  We verify in Appendix \ref{Algebra} that any two subalgebras of $\SH$ generated by $2n + r - 2$ simple transpositions give isomorphic restrictions of representations (through a particular isomorphism of the subalgebras).
  We will hence abuse notation, pick one such subalgebra $\SH'$, and notate $\Res_{\SH'}^\SH W$ by $\Res W$ for any $\SH$-module $W$.

  Throughout the rest of the paper, it will be useful to have precise notation;
  identify each partition with a tuple $\lambda = (\lambda_1^{a_1},\dots,\lambda_l^{a_l})$ having $\lambda_i > \lambda_{i+1}$, $a_i > 0$, and $\sum_i a_i\lambda_i = 2n + r$.
  Identify each of these with a subset $\brk \lambda \subset \NN^2$ as is standard, and define $\lambda(i) = (\lambda_1^{a_1},\dots,\lambda_{i-1}^{a_{i-1}},\lambda_i^{a_i - 1},\lambda_i-1,\lambda_{i+1}^{a_{i+1}},\dots,\lambda_l^{a_l})$ to be the partition with the $i$th row removed.\cite{Kleshchev}

  Fixing some partition $\lambda \vdash 2n + r$, for $1 \leq i \leq j \leq l$, let $\beta_\lambda(i,j)$ and $\gamma_\lambda$ be the quantities
  \begin{align*}
    \beta_\lambda(i,j) &= \lambda_i - \lambda_j + \sum_{t = i}^j a_t\\
    \gamma_\lambda(i,j) &= \lambda_i - \lambda_j + \sum_{t = i+1}^j a_t.
  \end{align*}
  Results due to Kleshchev and Brundan refer to \emph{normal} and \emph{good} numbers;
  for these, we will use the facts that 1 is always normal and that $j$ is normal when $\beta_\lambda(i,j) \not\equiv 0 \pmod e$ for all $i \leq j$.
  Further, we will use that $j$ is good if and only if $j$ is normal and $\gamma_\lambda(j,j') \not\equiv 0 \pmod e$ for all $j' \geq j$ normal.\cite{Brundan,Kleshchev}

  The following statements, collectively called \emph{modular branching rules} of $D^\lambda$, were originally written by Kleshchev for Specht modules of the group algebra $k\brk{S_n}$, then generalized to the Hecke algebra case by Brundan.\cite{Kleshchev}\cite{Brundan} 
  They entirely characterize the Socle of $\Res D^\lambda$, as well as the condition that $\soc\prn{\Res D^\lambda} \simeq \Res D^\lambda$.
  \begin{theorem}[Kleshchev-Brundan]
    We have the following isomorphisms of vector spaces
    \begin{align*}
      \Homarg{\SH'}{S^\mu,\Res D^\lambda} 
      &\simeq \begin{cases}
        k & \mu \xrightarrow{\text{normal}} \lambda\\
        0 & \text{otherwise}
      \end{cases}\\
      \Homarg{\SH'}{D^\mu,\Res D^\lambda} 
      &\simeq \begin{cases}
        k & \mu \xrightarrow{\text{good}} \lambda\\
        0 & \text{otherwise}
      \end{cases}
    \end{align*}
    and $\Res D^\lambda$ is semisimple if and only if every normal number in $\lambda$ is good.\qed
  \end{theorem}

  Using this, we immediately see that, for any rectangular partition $(m^\ell)$, we have
  \[
    \Res D^{(m^\ell)} \simeq D^{(m^{\ell-1},m-1)}.
  \]
  The non-rectangular two-row case is more complicated, but we may still describe it fully as follows.
  \begin{corollary}\label{D Restrictions}
    Suppose $r > 0$.
    Then, we may characterize the socle of $\Res \, D^\lambda$ as follows:
    \[
      \soc\prn{\Res \, D^{(n+r,n)}} \simeq \begin{cases}
        D^{(n + r - 1,n)} & e \mid r+2\\
        D^{(n + r,n-1)} & e \nmid r+2, \; e \mid r\\
        D^{(n+r-1,n)} \oplus D^{(n+r,n-1)} & e \nmid r+2,r
      \end{cases}.
    \]
    Further, when $e \nmid r$ or $e \mid r + 2$, $\Res D^{(n+r,n)}$ is semisimple.
  \end{corollary}
  \begin{proof}
    This amounts to computations of the hook lengths $\beta(1,2)$ and $\gamma(1,2)$:
    \begin{align*}
      \beta_\lambda(1,2) &= r + 2\\
      \gamma_\lambda(1,2) &= r
    \end{align*}
    Since $2$ is the largest removable number, $D^{(n+r,n-1)} \subset D^{(n+r,n)}$ iff $e \nmid r + 2$.
    Further, if $e \nmid r + 2$, then $D^{(n+r-1,n)} \subset D^{(n+r,n)}$ iff 1 is good iff $e \nmid r$
  \end{proof}

  Now that we've characterized how these restrict, we can describe how strongly these restrictions characterize irreducibles.
  Namely, we will prove that a composition series consistent with $(n+r,n)$ and $e$ top-indivisibility of $(n+r-1,n+1)$ is sufficient to determine that an irreducible is $D^{(n+r,n)}$ when either $r \neq 0$ or $e \neq 4$.
  \begin{proposition}\label{Combinatorics}
    Let $\lambda$ be an $e$-regular partition of $2n + r$.
    \begin{enumerate}[label={(\roman*)}]
      \item Suppose $r > 0$, suppose $e \nmid r+1, r + 2,\dots,n + r + 1$, and suppose either $e \mid r$ or $e \nmid r$.
        If $D^\lambda$ has the composition series 
        \begin{equation}\label{D Composition Series}
          0 \subset D^{(n+r-1,n)} \subset \Res \, D^{\lambda}
        \end{equation}
        with factor $\Res \, D^{\lambda} / D^{(n+r-1,n)} \simeq D^{(n+r,n)}$, then $\lambda = (n+r,n)$.
      \item Suppose $r = 0$, suppose $e \nmid 4$, and suppose $D^{(n,n-1)} \simeq \Res \, D^\lambda$.
        Then $\lambda = (n,n)$. 
    \end{enumerate}
  \end{proposition}
  \begin{proof}
    Note that $e > n$.

    \textbf{(i)}
    Let $\varpi := (n+r-1,n,1)$, let $\varsigma := (n+r-1,n+1)$, and let $\mu := (n + r,n)$.
    Since $D^{(n+r-1,n)} \subset \Res \, D^\lambda$, we have $(n + r - 1,n) \longrightarrow \lambda$, implying $\lambda = \varpi,\varsigma,\mu$.
    We will show that $\varpi, \varsigma$ do not have socle compatible with \eqref{D Composition Series};
    then, we will have $\lambda = \mu$.

    Suppose that $\lambda = \varpi$.
    We will break into cases with $r$.
    \begin{itemize}
      \item Suppose that $r > 1$.
        Note that $e \nmid r + 1 = \beta_\varpi(1,2)$, so $2$ is normal.
        Further, $\gamma_\varpi(2,3) = n \not\equiv 0 \pmod e$, so 2 is good and $D^{(n+r-1,n-1,1)} \subset D^\varpi$, which is not a composition factor in \eqref{D Composition Series}.
        Hence, by Jordan-H\"older, $\lambda \neq \varpi$.\cite{Etingof}

      \item Suppose that $r = 1$.
        Then, $\varpi = (n,n,1)$ has $\gamma_{\varpi}(1,2) = n \not\equiv 0 \pmod e$, giving $D^{(n,n-1,1)} \subset D^\varpi$ and hence $\lambda \neq \varpi$.
    \end{itemize} 
  \iffalse % keeping just in case
   \begin{align*}
    \beta_{\varpi}(1,2)     &= r + 1 \\
    \beta_{\varpi}(1,3)     &= n + r + 1 \\
    \beta_{\varpi}(2,3)     &= n + 1\\
    \gamma_{\varpi}(1,2)    &= r\\
    \gamma_{\varpi}(1,3)    &= n + r\\
    \gamma_{\varpi}(2,3)    &= n\\
  \end{align*}
  \fi

  \vspace{5pt}
  Now suppose that $\lambda = \varsigma$ and break into cases with $r$:
  \begin{itemize}
    \item Suppose $r > 2$.
      Then, by Corollary \ref{D Restrictions}, we require that $e \nmid r$ and $e \mid r - 2$;
      these are not satisfied, so $\lambda \neq \varsigma$.
    \item Suppose $r = 2$.
      Then $\Res \, D^\varsigma \simeq D^{(n+1,n)}$ is irreducible, contradicting \eqref{D Composition Series}. 
    \item Suppose $r < 2$. 
      Then $\varsigma$ is not a partition.
  \end{itemize}

  \textbf{(ii)}
  This result is easier than the previous result;
  since the socle of $D^\lambda$ is irreducible, we require that 1 is the only normal number and $\lambda(1) = (n,n-1)$.
  This reduces to the cases of $\varsigma := (n + 1, n - 1)$ and $\mu := (n,n)$;
  if $\lambda = \varsigma$, then we have that $e \mid \beta_\varsigma(1,2) = 4$, a contradiction.
  Hence $\lambda = \mu$.
\end{proof}

\iffalse
\begin{proposition}
  Suppose that $n > 0$ and $e \nmid n + r$.
  \begin{enumerate}[label={(\roman*)}]
    \item Suppose $r > 0$, $n > 1$, and either $e \nmid n + 1$ or $e \nmid n + r + 1, n + r - 1$.
      If $D^\lambda$ has the composition series 
      \begin{equation}\label{D Composition Series}
        0 \subset D^{(n+r-1,n)} \subset \Res \, D^{\lambda}
      \end{equation}
      with factor $\Res \, D^{\lambda} / D^{(n+r-1,n)} \simeq D^{(n+r,n)}$, then $\lambda = (n+r,n)$.
    \item Suppose $n = 1$ and $e \nmid r+2$, and suppose that we have composition series \eqref{D Composition Series}.
      Then $\lambda=(r+1,r+2)$.
    \item Suppose $r = 0$, either $e \nmid n+2$ or $e > 3$, and $D^{(n,n-1)} \simeq \Res \, D^\lambda$.
      Then $\lambda = (n,n)$.
  \end{enumerate}
  The above conditions are unilaterally met whenever $e > n + r + 1$ and $(n,r) \neq (1,0)$.
\end{proposition}
\begin{proof}
  \textbf{(i)}
  Let $\lambda_1 := (n+r-1,n,1)$, let $\lambda_2 := (n+r-1,n+1)$, and let $\lambda_3 := (n + r,n)$.
  Since $D^{(n+r-1,n)} \subset \Res \, D^\lambda$, we have $(n + r - 1,n) \longrightarrow \lambda$, implying $\lambda = \lambda_i$ for some $i = 1,2,3$.
  We will show that $\lambda_1,\lambda_2$ each have good numbers other than 1, leading to elements of their socle that are not factors in \eqref{D Composition Series};
  then, we will have $\lambda = \lambda_3$.

  We may straightforwardly compute this.
  First assume $n \geq 2$, so that
  \begin{align*}
    \beta_{\lambda_1}(1,2) &= n + 1\\
    \beta_{\lambda_1}(1,3) &= n + r + 1\\
    \beta_{\lambda_1}(2,3) &= n + r - 1\\
    \beta_{\lambda_2}(1,2) &= n + r
  \end{align*}

  \textbf{(ii)}
  The choices are the same as above, but now $\lambda_1$ only has two removable nodes, so that
  \begin{align*}
    \beta_{\lambda_1}(1,3) &= r + 2\\
    \beta_{\lambda_2}(1,2) &= r + 1
  \end{align*}

  \textbf{(iii)}
  Let $\lambda_1 := (n,n-1,1)$, let $\lambda_2 := (n+1,n-1)$, and let $\lambda_3 := (n,n)$;
  as before we know that $\lambda = \lambda_i$ for $i = 1,2,3$, so we will eliminate $\lambda_1,\lambda_2$.
  First, we will show that $\lambda_1$ has a good number other than 1:
  \begin{align*}
    \beta_{\lambda_1}(1,2) &= n\\
    \beta_{\lambda_1}(1,3) &= n + 2\\
    \beta_{\lambda_1}(2,3) &= n\\
  \end{align*}
  Hence we may rule out $\lambda_1$ when $e | n$.

  Suppose $D^{\lambda_2}$ satisfies the isomorphism;
  then the number 2 is normal.
  We may make the following computations:
   \begin{align*}
    \beta_{\lambda_2}(1,2) &= n + 2\\
    \gamma_{\lambda_2}(1,2) &= 3.
  \end{align*}
  When $e | n + 2$, this brings contradiction;
  when $e \nmid n + 2$ and $e > 3$, 2 is a good number (recalling that $e > 2$), giving non-semisimple restriction, a contradiction.
  Hence $\lambda = \lambda_3$ and the proposition is proven.
\end{proof}
\fi

\section{Crossingless Matchings and Specht Modules}\label{Crossingless Matchings Section}
In the following section, we will analyze the crossingless matchings representation $M_{2n + r}^r$ with the goal of proving $M_{2n + r}^r \simeq S^{(n+r,n)}$ whenever possible.
We begin by proving irreducibility of $M_{2n + r}^r$ in all cases that $(n+r,n)$ is $e$ top-indivisible;
when $p = 2$, this is all cases where $S^{(n+r,n)}$ is irreducible.
This is proven via an inductive process; if $e \nmid n + r + 1$, then $M$ contains no sign subrepresentation, and this allows us to ``project'' down to the case $M_{2(n-1) + r}^r$, and prove irreducibility of $M$ from irreducibility of this space.

\subsection{Irreducibility of $M$}
\begin{lemma}
  \label{Cyclic}
  Every basis vector in $M_{2n + r}^r$ is cyclic.
\end{lemma}
\begin{proof}
  We have already proven this in the $r = 0$ case, so suppose that $r > 0$.

    Note that, between anchors $a<a'$ having no arc $b$ with $a < b < a'$, the $M_{a'-a}^0$ case allows us to generate the vector with all length-2 arcs between $a,a'$ and identical arcs/anchors outside of this sub-matching.\footnote{At the ends, we apply the $M_a^0$ case or the $M_{2n + r - a}^{0}$ case in the same way for the first $a$ or last $2n + r - a$ indices.}

  Applying this between each arc gives us a vector with length-2 arcs and anchors, and we may use the appropriate $(1+T_i)$ to move anchors to any positions, and the reverse process from above to generate the correct matchings between arcs and generate any other basis vector.
\end{proof}

Let $K := \bigcap_{i=1}^{2n+r-1} \ker (1 + T_i) = \ker \bigoplus_{i = 1}^{2n + r - 1} (1 + T_i)$.
This will be a large technical tool in our proof of irreducibility.

\iffalse
\begin{lemma}\label{Represented}
  Let $w_j$ be the basis vector with anchors $1,\dots,r$ and all arcs of maximal length.
  Suppose $w \in K \backslash \cbr 0$.
  Then, $w_j$ is represented in $w$.
\end{lemma}
\begin{proof}
  We will show this in steps; first, we show that, given that a vector is represented with anchors $1,\dots,s$, there must be a vector represented in $w$ with $(s+1)$st anchor, including when $s = 0$;
  this implies that a vector is represented with anchors $1,\dots,r$.
  Then, we will show that, given a vector is represented with anchors $1,\dots,r$ and first $s$ arc-lengths $n,n-2,\dots,n-2s$, there is a vector represented with these and the $(s+1)$st arc-length $n-2s-2$.
  This implies that $w_j$ is represented.

  \def\lemsep{7pt}
  \vspace{\lemsep}
  \textit{Step 1.}
  Suppose that $s < r$ is the maximal number such that a vector with anchors $1,\dots,s$ is represented.
  Take the vector $w_i$ which, among vectors represented in $w$ with anchors $1,\dots,s$, has $(s+1)$st anchor at minimal index $t > s + 1$,
  Then, $q^{-1/2}(1 + T_{t-1})w_i$ has anchors $1,\dots,s$ and a earlier index than $t$, so it was not represented before;
  further, for any other basis vector $w_l \neq w_i$ to map onto $q^{-1/2}(1 + T_{t-1})w_i$, we would require that $w_l$ has anchors $1,\dots,s$ and some other anchor at index $t' < t$, so it is not represented.
  Hence $w_i$ is unique among the vectors represented mapping onto $q^{-1/2}(1 + T_{t-1})w_i$, and $(1 + T_{t-1})w$ represents this vector, giving $w \notin \ker (1 + T_{t-1})$.

  When $s = 0$, this is similar, and we simply perform this logic on the 1st anchor.
  Each lead to contradiction, so we must have $s = r$.

  \vspace{\lemsep}
  \textit{Step 2.}
  This step is similar;
  suppose that $s < n$ is the maximal number such that a vector with anchors $1,\dots,r$ and first $s$ arc-lengths $n,\dots,n-2s$ is represented.
  Take the vector $w_i$ which, among vectors represented in $w$ with anchors $1,\dots,r$ and first $s$ arc-lenths $n,\dots,n-2s$, has maximal length $t$ of the arc beginning at index $r + s + 1$.
  Then, $q^{1/2}(1 + T_{r + s + t})w_i$ is mapped to only by $w_i$ and vectors having anchors $1,\dots,r$ and first $s+1$ arc-lengths $n,\dots,n-2s,t'$ with $t' > t$, which are not represented in $w$;
  hence $q^{-1/2}(1 + T_{r + s + t})w_i$ is represented in $(1 + T_{r+s+t})w$, giving $w \notin \ker (1 + T_{r + s + t})$.
  The $s = 0$ case is similar, and implies that $s = n$.
\end{proof}
\fi

\begin{lemma}\label{Trivial}
  Suppose $e \nmid n + r + 1$.
  Then, $K = 0$.
\end{lemma}
\begin{proof}[Proof 1.]
  Consider the matrix $A = \bigoplus (1 + T_i)$ having kernel $K$.
  It is sufficient by lemma \ref{Represented} to show that $A$ includes a row $[0,\dots,0,1,0,\dots,0]$ with a nonzero entry only on the row $j$.

  Now, we may characterize the rows of $A$ as follows;
  if the row corresponding to $(1 + T_i)$ and mapping onto the element $w_l \in W$ is nonzero, then it is of the form $\brk{a_1,\dots,a_{\abs W}}$ where $a_l = 1 + q$, $a_m = q^{1/2}$ whenever $(1 + T_i)w_m = q^{1/2}w_l$, and $a_m = 0$ otherwise.

  Seeing this, the row corresponding to $(1 + T_{n + r})$ and $w_j$ has nonzero entries $q^{1/2}$ at $w_j$ and $(1 + q)$ at the vector $w$ agreeing with $w_j$ at all indices except having arcs at $(n+r-1,n+r)$ and $(n+r+1,n+r+2)$.
  Similar justification leads the row corresponding to $(1 + T_{n+r-1})$ at $w$ to have nonzero entries $q^{1/2}$ at $w$ and $(1 + q)$ at $w_j$ and the vector with anchors $1,\dots,r$ , arc $(n+r-3,n+r-2)$, and all other arcs maximum length.

  \begin{figure}
   \[
     \Action{10}{1/2,3/10,4/9,5/8,6/7}{2}{1/10,2/3,4/9,5/8,6/7}{q^{1/2}}
   \]
   \[
     \Action{10}{1/10,2/3,4/9,5/8,6/7}{2}{1/10,2/3,4/9,5/8,6/7}{q+1}
   \]
   \[
     \Action{10}{1/10,2/9,3/4,5/8,6/7}{2}{1/10,2/3,4/9,5/8,6/7}{q^{1/2}}
   \]
   \caption{Illustrated is the row constructed for transposition $(1 + T_2)$; clearly these are the only basis elements mapping to multiples of the desired element, and they relate to each other.
     replacing the outermost and/or innermost arc with an anchor typifies the rows constructed with three nonzero coefficients.}
 \label{Submatrix}
\end{figure}

  We may iterate this process as illustrated in Figure \ref{Submatrix}, eventually ending at a row with two nonzero entries, either an arc $(1,2)$ or an arc $(2,3)$, and all anchors otherwise left-aligned and arcs of maximum length.
  These rows together form an $(n + r) \times \abs{M_{2n + r}^r}$ submatrix of $A$ which has a nonzero row in the row corresponding to $j$, and has (by removing zero rows) the same row space as the following square matrix:
  \newcommand*\bigzero{\makebox(0,15){\text{\Huge0}}}
  \newcommand*\bigzerotwo{\makebox(-20,15){\text{\Huge0}}}
  \begin{equation}\label{Bnr}
    B_{n + r} := \begin{bmatrix}
      \begin{matrix}
      q + 1 & q^{1/2} & \\
      q^{1/2} & q + 1 & q^{1/2}\\
      & q^{1/2} & q + 1 & q^{1/2}\\
      &  & \ddots & \ddots
      \end{matrix}
      &   \bigzero\\
      \bigzerotwo & \begin{matrix}
        q^{1/2} & q+1 & q^{1/2}\\
        &  q^{1/2} & q + 1
       \end{matrix}
    \end{bmatrix}.
  \end{equation}
  We will show that this matrix is invertible;
  then, a sequence of elementary row operations will yield the identity, and in particular, when applied to $A$, will yield a row with a nonzero entry only on row $j$, giving $K = 0$.

  We may prove invertibility of this matrix by proving that $\det B_{n + r} = \brk{n + r + 1}_q$ inductively on $n+r$.
  This is satisfied for our base case $n + r = 1$, so suppose that it is true for each $m < n + r$.
  Then,
  \begin{align*}
    \det B_{n+r} &= (q + 1) \det B_{n + r - 1} - q \det B_{n + r - 2}\\
    &= (q + 1)(1 + \dots + q^{n + r - 1}) - (q + \dots + q^{n + r - 1})\\
    &= 1 + \dots + q^{n + r}\\
    &= \brk{n + r + 1}_q.
  \end{align*}
  Hence $\det B_{n + r} \neq 0$, and $K = 0$.
\end{proof}
\begin{theorem}\label{Irreducibility Theorem}
  Suppose that $(n+r,n)$ is $e$ top-indivisible.
  Then, the representation $M_{2n + r}^r$ is irreducible. 
\end{theorem}
\begin{proof}
  We proceed by induction on $n$.
  Note that, by identification with the trivial and sign representation, the base case $n = 0$ is already proven.

  Take an arbitrary vector $w \in W$.
  By Lemma \ref{Trivial} there exists some $(1 + T_i) \in \SH$ such that $(1 + T_i)w$.
  Note that \[\ima (1 + T_i) = \Span\cbr{w_j \mid w_j \text{ contains arc }(i,i+1)}.\]
  Hence, as vector spaces, there is an isomorphism $\varphi:\ima(1 + T_i) \rightarrow M_{2(n-1) + r}^r$ ``deleting'' the arc $(i,i+1)$.

  We will show that, for every action $(1 + T_j') \in \SH(S_{2(n-1) + r})$, there is some action $h_j \in M_{2n + r}^r$ such that the following commutes:
  \[
    \begin{tikzcd}
      \ima (1 + T_i) \arrow[r,"\varphi" above, "\sim" below] \arrow[d,"h_j"] & M_{2(n-1) + r}^r \arrow[d,"1 + T_j"]\\
      \ima (1 + T_i) \arrow[r,"\varphi" above, "\sim" below] & M_{2(n-1) + r}^r
    \end{tikzcd}
  \]
  Indeed, when $i \neq j$ this is given by $h_j = 1 + T_j$, and we have $h_i = q^{-1}(1 + T_i)(1 + T_{i+1})(1 + T_{i-1})$, as given by Figure \ref{bigloop}.

  Note that, by $e$ top-indivisibility of $(n+r,n)$, we have $e \nmid r+2,\dots,n+r$, and hence $(n+r-1,n-1)$ is $e$ top-indivisible as well.
  Then, due to the inductive hypothesis, there is some action $h' \in \SH(S_{2(n-1) + r})$ sending $\varphi((1 + T_i)w)$ to the image of a basis vector;
  then, the action $\SH$ generates the endomorphism $\varphi^{-1}h'\varphi$ sending $(1 + T_i)w$ to a basis vector, giving $w$ cyclic and hence $M_{2n + r}^r$ irreducible.
  \begin{figure}
  \begin{adjustbox}{width=\textwidth}
    \Action{6}{1/6, 2/5, 3/4}{2}{1/6, 2/3, 4/5 }{q^{1/2}} \hspace{20pt} 
    \IsoAction{8}{1/8, 2/7, 3/4, 5/6}{3}{1/8, 2/5, 3/4, 6/7}{q^{1/2}}
  \end{adjustbox}
  \begin{adjustbox}{width=\textwidth}
    \Action{4}{1/4, 2/3}{2}{1/4,2/3}{(1 + q)} \hspace{20pt}
    \IsoAction{6}{1/6, 2/5, 3/4}{3}{1/6, 2/5, 3/4}{(1 + q)}
\end{adjustbox}
  \caption{The correspondence between the action of $(1 + T_2)$ on $w'_5 \in M^0_6$ and the action of $q^{-1}(1 + T_3)(1 + T_4)(1 + T_2)$ on the corresponding vector in $M^0_8$ having arc $(3,4)$ first, then on $w'_2 \in M^0_4$.
  This demonstrates that the action works with and without creating a loop.
  }
  \label{bigloop}
  \end{figure}
\end{proof}

\subsection{Correspondence}
The following theorem of Mathas generalizes the characteristic-free version of the classical branching theorem.\cite{Mathas-article}
It will not be necessary for our present proof of the correspondence, but analogy with $M$ is suggestive.
\begin{theorem}[Characteristic-Free Classical Branching Theorem]
  Let $\lambda$ be a partition of $m$ with $\ell$ removable nodes.
  Then, $\Res \, S^\lambda$ has an $\SH_{k,q}(S_{m-1})$-module filtration
  \[
    0 = S^{0,\lambda} \subset S^{1,\lambda} \subset \cdots \subset S^{\ell,\lambda} = \Res S^\lambda
  \]
  such that $S^{t,\lambda} / S^{t-1,\lambda} \simeq S^{\lambda(t)}$ for all $1 \leq t \leq \ell$.\qed
\end{theorem}

If we replace $S^\lambda$ with the appropriate $M_{2n + r}^r$ above, we find the statement of the following proposition.
\begin{proposition}Suppose that $n > 0$.

  \begin{enumerate}[label={(\roman*)}]
    \item
    Suppose that $r > 0$.
    Then, a filtration of $\Res M_{2n + r}^r$ is given by
    \begin{equation}
      0 \subset M_{2n + r - 1}^{r-1} \subset \Res \, M_{2n + r}^r \label{Filtration}
     \end{equation}
     with $\Res \, M_{2n + r}^r / M_{2n + r - 1}^{r-1} \simeq M_{2n + r - 1}^{r + 1}$. 
    \item
      We have the following isomorphism of representations:
      \begin{equation}
        M_{2n - 1}^1 \simeq \Res \; M_{2n}^0 \label{0 Restriction}
       \end{equation}
  \end{enumerate}
    When the case is type (i) from the irreduibility lemma, this is a composition series.
\end{proposition}
\begin{proof}
  \textbf{(i)}
  Note that we may identify the subrepresentation of $\Res \; M_{2n + r}^r$ having anchor $n$ with $M_{2n + r - 1}^{r-1}$.

  Let $U := \Res \; M_{2n + r}^r / M_{2n + r - 1}^{r - 1}$. 
  Let $\phi:U \rightarrow M_{2n + r - 1}^{r + 1}$ be the $k$-linear map which regards the arc $(i,2n + r)$ in $U$ as an anchor at $i$ in $M_{2n + r - 1}^{r + 1}$.
  It is not hard to verify that this is a well-defined isomorphism of vector spaces, so we must show that it is $\SH$-linear.

  Given a basis vector $w_j$ with arc $(i,2n + r)$, $\phi$ is clearly compatible with $T_{i'}$ with $i' \neq i,i-1$.
  Further, it's easy to verify that $\phi$ is compatible with $T_{i}$ and $T_{i-1}$, as actions on one anchor were designed for this deformation.
  When there are anchors $(i,i+1)$, then $\phi(T_iw_j) = T_i\phi(w_j) = 0$, and similar for $T_{i -1}$.
  Hence $\phi$ is an isomorphism of representations, and the statement is proven.

  \textbf{(ii)}
  This follows with the above proof, defining $M_{2n - 1}^{-1} := 0$
\end{proof}

We've now assembled the basic pieces necessary to prove our correspondence in the case $e > n + r  1$.
Due to some problematic phenomena in the $r = 0$ case, we will initially make an assumption that $e \neq 4$ or $r \neq 0$ so that we may use the combinatorics of Proposition \ref{Combinatorics}.
We will then assume the remaining case $e = 4, r = 0$, and prove our correspondence using process of elimination via the other surrounding cases for $e = 4$.

\begin{theorem}\label{Correspondence Theorem}
  Suppose $e > n + r + 1$.
  Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.
\end{theorem}
\begin{proof}
  The case $n = 0$ is already proven, so suppose $n > 0$.
  In order to use Proposition \ref{Combinatorics}, suppose for now that either $e \nmid 4$ or $r \neq 0$.

  We will prove this inductively; by identification with the trivial and sign representations, the $2n + r = 2$ caseholds, so suppose this is true for $M_{2m + s}^s$ whenever $2m + s < 2n + r$ and $m + s \leq n + r$ (i.e. $e > m + s + 1$).

  By irreducibility, we know that $M_{2n + r}^r \simeq D^\lambda$ for some $e$-restricted partition $\lambda$.
  By the inductive hypothesis and irreducibility, we have a composition series given by the exact sequence
  \begin{equation}
    \label{Composition Series} 0 \longrightarrow D^{(n + r - 1,n)} \longrightarrow \Res \, D^\lambda \longrightarrow D^{(n+r,n-1)} \longrightarrow 0
   \end{equation}
  Hence the theorem is given by Proposition \ref{Combinatorics}.

  Now, suppose $e = 4$ and $r = 0$;
  then $4 > n + 1$, so $n \leq 2$.
  We've already proven the $n = 1$ case via the trivial representation, so suppose $n = 2$.
  Then, from the proof of Proposition \ref{Combinatorics}, we know that $M_{4}^0 \simeq D^\lambda$, where $\lambda \in \cbr{(n,n), (n+1,n-1)}$.
  We have already proven that $M_4^2 \simeq D^{(n+1,n-1)}$, and we may verify that $\dim M_4^0 = 2 \neq 3 = \dim M_4^2$, so we have that $\lambda = (n,n)$ and the theorem is proven for $e = 4$.
\end{proof}

\iffalse
We may reuse the above structure of proof as above;
suppose $M_{2n + r}^r$ and $S^{(n+r,n)}$ are irreducible, and $\lambda = (n+r,n)$ is the unique diagram having composition series \eqref{Composition Series}.
Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.

We have covered all irreducible cases for $r = 0$ and $n = 0$ above.
For example, another case where we may show all irreducible cases is $n = 1$, which we will do now.
\begin{lemma}\label{n=1 irreducibility}
    Set $n = 1$. 
    Then $M_{2 + r}^r$ is irreducible iff $e \nmid r + 2$.
\end{lemma}
\begin{proof}
  A basis element $w_i \in W$ is determined entirely by the position of the arc, of which there are $r + 1$ options.
  Hence $\dim W = r + 1$.

  Further, $\dim\prn{\ima (1 + T_i)} = 1$ for all $i$;
  by lemma \ref{Cyclic} $W$ is irreducible iff $K := \bigcap_{i = 1}^{r+1} \ker (1 + T_i) = \ker \bigoplus (1 + T_i)$ is trivial.
  By \ref{Trivial}, this is true iff $e \nmid r + 2$.
\end{proof}

\begin{theorem}
  Set $n = 1$ and suppose $e \nmid r + 2$.
  Then $M_{2 + r}^r \simeq S^{(1+r,1)}$.
\end{theorem}
\begin{proof}
  Using lemma \ref{n=1 irreducibility} and \ref{S irreducibility} it is sufficient to prove that, for $D^\lambda$ having composition series \eqref{Composition Series}, we have $\lambda = (1+r,1)$.
  Similar to before, we consider the options $\lambda_1 := (r,1,1)$, $\lambda_2 := (r,2)$, and $\lambda_3 = (1+r,1);$.
  We may straightforwardly check that $\lambda_1$ and $\lambda_2$ have socle containing partitions other than the composition factors in \eqref{Composition Series} by showing that each have a good number 2 as follows:
  \begin{align*}
    \beta_{\lambda_1}(1,2) &= 3 - 1 + r = r+2\\
    \beta_{\lambda_2}(1,2) &= 2 - 1 + r + 1 = r+2.
  \end{align*}
  These are each nonvanishing mod $e$, so they are good and $\lambda = \lambda_3$, giving the theorem.
\end{proof}
\fi

\subsection{Kernels and Further Work}
\begin{proposition}\label{Represented}
	Let $w$ be an arbitrary vector in $W_{2n+r}^r$. I claim that if $w\in \cap\text{ker}(1+T_i)$, the coordinate $c$ of the rainbow element $R$ in $w$ is nonzero.
\end{proposition}	
\begin{proof}
  Let $Y$ be the set of basis elements with nonzero coordinate in $w$. Let $k$ be the greatest integer such that there exists $y\in Y$ where $y(1)=...=y(k)=0$ {\color{magenta} should this be $y(1)-1 = \dots = y(k) - k = 0$? Also, we should avoid using $k$ as an integer, as it's used elsewhere as a field.}, and let $U\subset Y$ be the set of such $y$. In other words, $U$ is the set of basis elements in $Y$ which have the most anchors to the far left.

	Suppose $k<r$. Then for each $y\in U$ there exists a minimal $i_y>k+2$ such that $y(i_y)=0$. In other words, $i_y$ is the position of the next leftmost anchor in $y$.  Fix $\tilde{y}$ such that $i_{\tilde{y}}\leq i_y$ for all $y$. Then I claim the basis element $y':=q^{-1/2}(1+T_{i_{\tilde{y}}-1})\tilde{y}$ has nonzero coordinate in $(1+T_{i_{\tilde{y}}-1})w$, implying $w\not\in \cap\text{ker}(1+T_i)$. To see this, we can show that $\tilde{y}$ is the only element in $Y$ such that $q^{-1/2}(1+T_{i_{\tilde{y}}-1})\tilde{y}\sim y'$. $y'$ still has $k$ anchors on the left, and $i_{y'}<i_{\tilde{y}}$, so $y'\not\in Y$. If $x\in Y,\not\in U$, the basis element proportional to $(1+T_{i_{\tilde{y}}-1})x$ will have $k$ anchors at the far left only if the next anchor is at a position $i_{x'}>i_{\tilde{y}}$, so it cannot be $y'$. If $x\in U$ the basis element proportional to $(1+T_{i_{\tilde{y}}-1})x$ will have anchor at $i_{y'}$ if and only if $i_x=i_{\tilde{y}}$ and $x(i_{\tilde{y}})=\tilde{y}(i_{\tilde{y}})$. Since this is the only match altered by action $(1+T_{i_{\tilde{y}}-1})$ on $x$, if $(1+T_{i_{\tilde{y}}-1})x\sim y'$ this implies $x=\tilde{y}$. So if $k<r$ $w$ is not in the desired kernel.

	Suppose $k=r$ but $R\not\in U$ (so $R\not\in Y$). Let us define a sequence of subsets of $U$ in the following way: $U_0:=U$, $U_{i+1}:=\{u\in U_i| u(r+i+1)=2n+2r-i+1 \}$. Since $R\not\in U$, $\exists t<n-1$ such that $U_{t+1}=\emptyset$. Choose $\tilde{u}\in U_t$ such that $\tilde{u}(r+t+1)\geq u(r+t+1)$ for all $u\in U_t$. Consider the basis element $u':=q^{-1/2}(1+T_{\tilde{u}(r+t+1)})\tilde{u}$. I claim that $\tilde{u}$ is the only element in $Y$ such that $(1+T_{\tilde{u}(r+t+1)})\tilde{u}\sim u'$, again implying that $w$ is not in the desired kernel. $u'$ still has $k$ anchors on the left, $u'(r+i)=2n+2r-i+2$, $1\leq i\leq t$, and $u'(r+t+1)>\tilde{u}(r+t+1)$, so $u'\not\in Y$. If $x\in Y,\not\in U$,  the basis element $x'$ proportional to $(1+T_{\tilde{u}(r+t+1)})x$ will have $r$ leftmost anchors only if $x'(r+t+1)<\tilde{u}(r+t+1)$, so $x'\not=u'$. Similarly, if $x\in U,\not\in U_t$, the basis element $x'$ will have the property $x'(r+t)=2n+2r-t+2$ only if $x'(r+t+1)<\tilde{u}(r+t+1)$, so $x'\not=u'$. If $x\in U_t$, $x'(r+t+1)=u'(r+t+1)$ if and only if $x(r+t+1)=\tilde{u}(r+t+1)$ and $x(x(r+t+1)+1)=\tilde{u}(\tilde{u}(r+t+1)+1)$ (since $u'\not \in Y$). These are the only matches altered by the action $(1+T_{\tilde{u}(r+t+1)})$, so this implies $x=\tilde{u}$. Thus we have proved that if $R\not\in Y$, $w$ is not in the desired kernel.
\end{proof}

{\color{orange} Necessary characterization of the Kernel goes here.}

\begin{proposition}
  Suppose $e = n + r + 1$.
  Then $K$ is nontrivial.
\end{proposition}

\section{Fibonacci Representations and Quotients of Specht Modules}\label{Fibonacci Section}
We have difficulty characterizing the Specht module above when it is irreducible.
We may instead attempt to characterize the irreducible quotient $D^\lambda$ of the Specht module.
In particular, note that, whenever $e \leq r + 2$, the restriction $\Res D^{(n+r,n)}$ decomposes into a direct sum of partitions $(m+s,m)$ such that $e \leq s + 2$ and $2m + s = 2n + r - 1$.
This forms a recurrence among these representations, allowing a combinatorial description of their dimension.
In this section we henceforth assume $e = 5$ and note that this recurrence resembles the Fibonacci recurrence;
we follow this to characterize the restriction $D^{(n+r,n)}$ with $r \leq 3$.

\begin{remark}
  Let $d_{m}^{0,3}$ be the dimension $\dim D^{(n,n)}$ with $2n = m$ when $m$ is even, and $\dim D^{(n+3,n)}$ with $2n + 3 = m$ when $m$ is odd.
  Similarly define $d_m^{1,2}$.
  Then, Corollary \ref{D Restrictions} gives
  \begin{align*}
    d_m^{0,3} &= d_{m-1}^{1,2}\\
    d_m^{1,2} &= d_{m-1}^{1,2} + d_{m-1}^{0,3}\\
    &= d_{m-1}^{0,3} + d_{m-2}^{0,3}.
  \end{align*}
  Carefully following this and noting the base cases $d_2^{0,3} = d_2^{1,2} = 1$, one may note that this recurrences proves that $d_m^{1,2} = d_{m+1}^{0,3} = f_n$, where $f_n$ is the $n$th \emph{Fibonacci number}.
  This matches the dimension our Fibonacci subrepresentations, motivating their definition.\qed
\end{remark}

We can start our study of $V$ by studying low-dimensional cases.
First, note that $V_{\vs\vp}^2$ is the sign representation $D^{(1^2)}$ and $V_{\vs\vp}^2$ is the trivial representation $D^{(2)}$.

$V_{\vp\vp}^2$ is a 2-dimensional representation of a semisimple commutative algebra, and hence decomposes into a direct sum of two subrepresentations.
In particular, we can use the basis $\cbr{(\vp\vs\vp),(\vp\vp\vp)}$ and explicitly write the matrix
\[
  \rho_{T_1} = \begin{bmatrix}
    \varc & \vard\\
    \vard & \vare
  \end{bmatrix}
\]
having characteristic polynomial $(\varc - \lambda)(\vare - \lambda) - \vard^2 = \lambda^2 - (\varc + \vare)\lambda +(\varc\vare - \vard^2)$.
We may verify that, for $\lambda = -1$, this evaluates to
\[
  -((-1 + q + q^2) (1 + q^3 + q^4 + q^5 + 2 q^6 + q^7))\brk{5}_q = 0
\]
and for $\lambda = q$ this evaluates to 
\[
-(q^2 (-1 + q + q^2) (1 + q + q^2 + q^3 + 2 q^4 + q^5)) \brk{5}_q = 0
\]
hence $\rho_{T_1}$ has eigenvalues $-1$ and $q$.

The eigenspaces with eigenvalues $-1$ and $q$ are subrepresentations isomorphic to the sign and trivial representation, hence $V_{\vp\vp}$ is isomorphic to a direct sum of the trivial and sign representations: $V^2_{\vp\vp} \simeq V^2_{\vs\vp} \oplus V^2_{\vs\vs}$.

Now we may prove that $V^3_{\vs\vp}$ is irreducible;
this has basis $\cbr{\prn{\vs\vp\vs\vp}, \prn{\vs\vp\vp\vp}}$, and the following matrices:
\[
  \rho_{T_1} = \begin{bmatrix}
    \varb & 0\\
    0 & \vara
  \end{bmatrix}; \hspace{20pt}
  \rho_{T_2} = \begin{bmatrix}
    \varc & \vard\\
    \vard & \vare
  \end{bmatrix}.
\]
A proper nontrivial subrepresentation of $V_{\vs\vp}^3$ must be one-dimensional, and hence an eigenspace of each of these matrices;
since $\varb \neq \vara$, the first has eigenspaces given by the spans of basis elements, and since $\vard \neq 0$, these are not eigenspaces of the second.
Hence $V^3_{\vs\vp}$ is irreducible.

These establish the low-dimensional behavior that we will use as base cases below.
The rest of this section will proceed first by proving that $V_{\vs\vp}$ and $V_{\vs\vs}$ are irreducible;
then, we will use combinatorial arguments to prove that these are isomorphic to the correct two-row Specht module quotients.

\begin{proposition}
  The representation $V_{\vs\vp} := V^m_{\vs\vp}$ is irreducible.
\end{proposition}
\begin{proof}
  We will prove this inductively in $m$.
  We've already proven it for $V^2_{\vs\vp}$ and $V^3_{\vs\vp}$, so suppose that $V^{m-2}_{\vs\vp}$ is irreducible.

  Let $\cbr{v_i}$ be the basis for $V_{\vs\vp}$.
  Then, each $v_i$ is cyclic; indeed, we can transform every basis vector into $(\vs\vp \dots \vp)$ by multiplying by the appropriate $\frac{1}{\vard-\varc}(T_i - \varc)$, and we can transform $(\vs\vp \dots \vp)$ into any basis vector by multiplying be the appropriate $\frac{1}{\vard-\vare}(T_i - \vare)$.
  Hence it is sufficient to show that each $v \in V_{\vs\vp}$ generate some basis element.

  Let $v'$ be the basis element $(\vs\vp\vs\vp\dots \vp)$, which is many copies of $\vs\vp$, followed by an extra $\vp$ if $m$ is odd.
  We will show that each $v \in F$ generates $v'$.

  Suppose that no elements beginning $(\vs\vp\vs\vp)$ are represented in $v_i$;
  then, all such elements are represented in $T_3v$, so we may assume that at least one is represented in $v$.

  Note that $\ima (T_2 - \vara) = \Span\cbr{\text{Basis vectors beginning }(\vs\vp\vs\vp)}$ and $(T_2 - \vara)v \neq 0$.
  Further, note that $\Res_{\SH(S_{m-2})}^{\SH(S_m)} \ima (T_2 - \vara) \simeq V^{m-2}_{\vs\vp}$ as representations.
  Hence irreducibility of $V^{m-2}_{\vs\vp}$ implies that $v'$ is generated by $(T_2 - \vara)v$, and $V^m_{\vs\vp}$ is irreducible.
\end{proof}

Now, we may begin considering restrictions:
\begin{lemma}
  The following branching rules hold:
  \begin{align*}
    V^{m-1}_{\vp\vp} \simeq \Res V_{\vs\vp}^m &\simeq V^{m-1}_{\vs\vs} \oplus V^{m-1}_{\vs\vp}\\
    \Res V_{\vs\vs}^m &\simeq V_{\vs\vp}^{m-1}.
  \end{align*}
\end{lemma}
\begin{proof}
  The first line follows by considering the last two $m-2$ transpositions for the left isomorphism, then the first two for the right isomorphism.
  This is well-behaved by Appendix \ref{Algebra}.

  Similarly, the second line follows by considering the last $m-2$ transpositions.
\end{proof}
This immediately gives a rather strong characterization of $V$.
\begin{corollary}
  The representation $V_{\vs\vs}$ is irreducible.\qed
\end{corollary}
\begin{corollary}
  The representation $V$ decomposes into a direct sum of irreducible representations as follows:
  \[
    V \simeq 3V_{\vs\vp} \oplus 2V_{\vs\vs}.
  \]\qed
\end{corollary}

Now we may use these in order to apply Young Tableau to characterize $V$.
\begin{theorem}\label{Fibonacci Theorem}
  The irreducible components of $V$ are given by the following isomorphisms:
     \begin{align*}  
      V^{2n}_{\vs\vs} &\simeq D^{(n,n)}\\ 
      V^{2n-1}_{\vs\vs} &\simeq D^{(n+1,n-2)}\\
      V^{2n}_{\vs\vp} &\simeq D^{(n+1,n-1)}\\
      V^{2n-1}_{\vs\vp} &\simeq D^{(n,n-1)}.
     \end{align*}
\end{theorem}
\begin{proof}
  We will prove this by induction on $n$;
  we have already proven the base case $V^{2}$, so suppose that we have proven these isomorphisms for $V^{2n-2}$.
  We will prove the isomorphisms for $V^{2n-1}$ and $V^{2n}$.

  By irreducibility, $V^{2n-1}_{\vs\vs} \simeq D^{\lambda}$ and $V^{2n-1}_{\vs\vp} \simeq D^{\mu}$ for some diagrams $\lambda$ and $\mu$.
  We will show that $\lambda = (n+1,n-2)$ and $\mu = (n+1,n-1)$.

  First, note that we have \[\Res \; D^{\lambda} \simeq D^{(n,n-2)} \simeq \Res \; D^{(n+1,n-2)}\] and \[\Res \; D^{\mu} \simeq D^{(n,n-2)} \oplus D^{(n-1,n-1)} \simeq \Res \; D^{(n,n-1)}.\]

  By irreducibility of $\Res D^{\lambda}$, the only normal number in $\lambda$ is 1.
  Further, the only tableaux which can remove a cell to yield $D^{(n,n-2)}$ are $(n+1,n-2)$, $(n,n-1)$, and $(n,n-2,1)$ as illustrated in Figure \ref{OddRes};
  we have already seen that $D^{(n,n-1)}$ does not have irreducible restriction, so we are left with $(n+1,n-2)$ and $\varsigma = (n,n-2,1)$.
  We may directly check that $\varsigma$ doesn't satisfy this, as we have the following:
  \begin{align*} 
    \beta_\varsigma(1,2) &= 3 - 2 + (n-2) = n-1\\
    \beta_\varsigma(1,3) &= 3 - 1 + n = n+2\\
    \beta_\varsigma(2,3) &= 2 - 1 + 3 = 4.
   \end{align*} 
  At least one of $\beta_\varsigma(1,2)$ and $\beta_\varsigma(1,3)$ is nonzero, since $\beta_\varsigma(1,3) - \beta_\varsigma(1,2) = 3 \not\equiv 0 \pmod e$, and hence at least one of $M_2$ and $M_3$ is empty.
  Hence at least one of 2 or 3 is normal in $\varsigma$, and $\lambda = (n+1,n-2)$.

  For $\mu$, we immediately see from Figure \ref{OddRes} that the only option is $(n,n-1)$.
  \begin{figure}
  \[
      \cbr{
        \begin{gathered}
        \ydiagram{6,3}. \; \;
        \ydiagram{5,4}, \; \;
        \ydiagram{5,3,1}
      \end{gathered} 
    }\longrightarrow \begin{gathered}
      \ydiagram{5,3}
  \end{gathered}\]
      \[\cbr{
        \begin{gathered}
          \ydiagram{5,4}
        \end{gathered}
      }\longrightarrow
      \begin{gathered}
        \ydiagram{5,3},\;\;
        \ydiagram{4,4}.
      \end{gathered}
    \]
      \caption{
      Illustration of the partitions of $9$ which can, via row removal, yield $(n,n-2)$ alone, or both $(n,n-2)$ and $(n-1,n-1)$.
    }\label{OddRes}
  \end{figure}

  We can perform a similar argument for the $V^{2n}$ case, finding now that 
  \[\Res \; D^{\lambda'} \simeq D^{(n,n-1)} \simeq \Res \; D^{(n,n)}\] and 
  \[\Res \; D^{\mu'} \simeq D^{(n,n-1)} \oplus D^{(n+1,n-2)} \simeq \Res \; D^{(n+1,n-1)}.\]

  Through a similar process, we see that $\mu' = (n+1,n-1)$.
  We narrow down $\lambda'$ to one of $(n,n)$ or $\varpi := (n,n-1,1)$, and note that
  \begin{align*} 
    \beta_\varpi(1,2) &= 3 - 2 + (n-1) = n\\
    \beta_\varpi(1,3) &= 3 - 1 + n = n+2\\
    \beta_\varpi(2,3) &= 2 - 1 + 2 = 3
  \end{align*}
  and hence at least one of 2 or 3 is normal, $\Res D^{\varpi}$ is not irreducible, and $\lambda' = (n,n)$, finishing our proof.
\end{proof}
\begin{corollary}
  We have the following isomorphisms of representations:
  \begin{align*} 
    V^{2n} &\simeq 3D^{(n+1,n-1)} \oplus 2D^{(n,n)}\\
    V^{2n - 1} &\simeq 3D^{(n,n-1)} \oplus 2D^{(n+1,n-2)}
  \end{align*}
\end{corollary}

\section{Conjecture}\label{Conjecture Section}
Recall that $K_{2n + r}^r := K$ is the direct sum of all copies of the sign representation in $W$.
Hence the following characterises sign subrepresentations of $W$ completely:
\begin{proposition}
  $K \subset M_{2n + r}^r$ is trivial when $e \neq n + r + 1$, and $\dim \, K = 1$ when $e = n + r + 1$.\qed
\end{proposition} 
\begin{proposition}
  Suppose $e < n + r + 1$, and suppose $n'$ is such that $e = n' + r + 1$.
  Note that $h := (1 + T_1)(1 + T_3)\dots(1 + T_{n - n'})$ maps $M_{2n + r}^r$ onto $M_{2n' + r}^r$.
  Then, the preimage $h^{-1}(K_{2n + r}^r)$ is a subrepresentation of $M_{2n + r}^r$, and the series
  \[
    0 \subset h^{-1}(K_{2n + r}^r) \subset M_{2n + r}^r
  \]
  is a composition series of $M_{2n + r}^r$.\qed
\end{proposition}
\begin{proposition}
  Denote the composition factor $M_{2n + r}^r / h^{-1}(K_{2n + r}^r)$ by $U_{2n + r}^r$.
  Then, there exist some naturals $m,s$ satisfying $2m + s = 2n + r$ and $m + s > n + r$ such that the following is an isomorphism of $\SH$-modules \[h^{-1}\prn{K_{2n + r}^r} \simeq U_{2m + s}^s\].
  \qed
\end{proposition}
\begin{proposition}
  For the same $m,s$ as above, we have the following composition series of Specht modules:
  \[
    0 \longrightarrow D^{(m+s,m)} \longrightarrow S^{(n+r,n)} \longrightarrow D^{(n+r,n)} \longrightarrow 0.
  \]
\end{proposition}
\begin{proposition}
  $M_{2n + r}^r \simeq S^{(n+r,n)}$ and $U_{2m +s}^s \simeq D^{(m+s,m)}$.
\end{proposition}

\section{Empirical Results}\label{Empirics Section}
In this section we aim both to support conjecture with data and to provide transition matrices wherever possible between our two graphical representations.\cite{Github}

\begin{remark}
  These results are restricted by two things; first and with less recourse, this was computed in Magma, a closed source language with bugs that could not be resolved in certain cases.
  Second, the memory necessary to store the Hecke algebra via basis elements and structure constants grows with $(n!)^3$;
  this is prohibitively large when $m > 7$.
  An implementation as a quotient by a free algebra or only specifying multiplication by generators may fix this, however, this is not possible with the Magma language, and the closed source renders modification of the language impossible.
\end{remark}

We give in Figure \ref{Fib Data} some isomorphisms between $M/K$ and $V$ for $e \geq n + r + 1$ with $p = \infty$;
all but one of these are cases with $e > n + r + 1$, so $K=0$ and this is an isomorphism with our crossingless matchings representation.
The case $n = 1$, $r = 3$ gives an example of an isomorphism not proven in general, but proven via our computation.
All of these computations are done for $q$ a primitive 5th root of unity in the algebraic extension of the Cyclotomic field $\QQ(\zeta_{10})$ by a root of the polynomial $x^2 - \tau$.

We give in Figure \ref{Empirical Kernel} some data supporting a conjecture concerning sign subrepresentations of $M_{2n + r}^r$.
The computations to support this were done over $\CC$ with $q$ a primitive 5th root of unity.

It is known that, for small $2n + r$, each Specht module $S^{(n+r,n)}$ has a composition series of length 2.
We give in Figures \ref{Composition series e=3} through \ref{Composition series e=6} the map $M_{2n + r}^r \twoheadrightarrow U_{2n + r}^r$, which conjecturally illustrates the quotient $S^{(n+r,n)} \twoheadrightarrow U_{2n + r}^r$ for all $2n + r \leq 7$.

\begin{figure}[h]
\iffalse
\def\tho{\Phi_3(q^{1/2})}
\def\tht{\brk{3}_{q^{1/2}}}
\def\to{q^{1/2}\brk{2}_{q^{1/2}}}
Define the following coefficients:
\begin{align*}
  \tho &:= q^{3/2} + 1\\
  \tht &:= q^{3/2} + q + q^{1/2} + 1\\
  \to &:= q + q^{1/2}
\end{align*}
\fi

\def\tho{q^{3/2} + 1}
\def\tht{\brk{4}_{q^{1/2}}}
\def\to{q^{1/2}\prn{q^{1/2} + 1}}
\begin{align*}
  \varphi_6^0, \varphi_5^1 &= \begin{bmatrix}
    0 & 0 & -\tho & 0 & 0\\
    0 & -\tho & \tht & 0 & 0\\
    0 & 0 & \tht & 0 & -\tho\\
    -\tht & \tht & \to & 0 & \tht\\
    \tht & 0 & \tht & -\tht & 0
  \end{bmatrix}\\
  \varphi_6^2 &= \begin{bmatrix}
    0 & 0 & - q - 1 & 0 & 0 & 0 & 0 & 0\\
    0 & -q - 1 & -q^{1/2} & 0 & 0 & 0 & 0 & 0\\
    0 & q + 1 & 0 & q^{1/2} & 0 & 0 & 0 & 0\\
    0 & 0 & -q^{1/2} & 0 & 0 & 0 & -q^2 - 1 & 0\\
    q^{1/2} & -q^{1/2} & -\brk{3}_{q^{1/2}} & 0 & 0 & 0 & -q^{1/2} & 0\\
   -q^{1/2} & q^{1/2} & 0 & \brk{3}_{q^{1/2}} &0 & 0 & 0 & q^{1/2}\\
   -q^{1/2} & 0 & -q^{1/2} & 0 & q^{1/2} & 0 & 0 & 0\\
    q^{1/2} & 0 & 0 & 0 & -q^{1/2} & -\brk{3}_{q^{1/2}} & 0 & -q^{1/2}
  \end{bmatrix}\\
  \varphi_5^3 &= \begin{bmatrix}
    -\to & \tht & q^{1/2}\\
    q^{1/2} & q^{3/2} & 0\\
    0 & -q - 1
  \end{bmatrix}  \hspace{30pt} 
  \varphi_4^0 = \begin{bmatrix}
    0 & -\tht\\
    -1 & 1
  \end{bmatrix}\\
  \varphi_3^1 &= \begin{bmatrix}
    0 & -\tht\\
    \to & -\to
  \end{bmatrix} \hspace{27pt}
  \varphi_4^2 = \begin{bmatrix}
    0 & \tht & 0\\
    1 & -1 & 0\\
    -1 & 0 & 1
  \end{bmatrix}
\end{align*}

\caption{We define representations $\varphi_{2n + r}^r:M_{2n + r}^r/K_{2n + r}^r \rightarrow V^{2n + r}_{s}$, where $s = \vs\vs$ if $r = 0,3$ and $s = \vs\vp$ otherwise.
These are given with respect to the basis on $M_{2n + r}^r$ induced by the increasing lexicographic order basis on $M_{2n + 2r}^0$ and the quotient $M_{2n + 2r}^0 \twoheadrightarrow M_{2n + r}^r$.
These are also given with respect to the basis on $V$ given by increasing lexicographic order $\vs < \vp$.
  }\label{Fib Data}
\end{figure}

\begin{figure}[h]
  \[
    \ENRGrid{13/7}{2/1,5/1,8/1,11/1,1/2,4/2,7/2,10/2}
  \]\[
    \ENRGrid{13/7}{3/1,7/1,11/1,2/2,6/2,10/2,1/3,5/3,9/3}
  \]
  \caption{
    Illustration of the modules $M_{2n + r}^r$ having sign submodules for $p = \infty$ and $e = 3,4$ respectively.
    The value $n + r + 1$ is filled in the squares, and modules having sign submodules are colored magenta.
    For $p = \infty$ and $2n + r \leq 14$, it has been verified, through a combination of theorems here and empirical computations, that $K_{2n + r}^r$ is nontrivial if and only if $e | n+r+1$ and $e < n$.
  }\label{Empirical Kernel}
\end{figure}

\def\qqo{\alpha}
\def\tto{\beta}
\def\vtv{\gamma}
\def\fft{\delta}
\def\fto{\varepsilon}
\setcounter{MaxMatrixCols}{20}

\begin{figure}[h]
\begin{align*} 
  \pi_{3,3}^1 = \pi_{3,4}^0 &=  
\begin{bmatrix}
  1 & 1
\end{bmatrix}^\intercal\\
\pi_{3,5}^1 = \pi_{3,6}^0&=  
\begin{bmatrix}
  1 & 0 & 0 & 0 & 0\\
  0 & 0 & 1 & 1 & 0\\
  0 & 1 & 0 & 1 & 0\\
  0 & 0 & 0 & 1 & 1
\end{bmatrix}^\intercal\\
\pi_{3,6}^4&=  
\begin{bmatrix}
  -1 & -1 & 0 & 1 & 1
\end{bmatrix}^\intercal\\
\pi_{3,7}^3&=  
\begin{bmatrix}
 1 & 0 &-1 &-1 & 1 & 0 & -1 & -1 & 0 & 0 & 0 & 0 & 1 & 1
\end{bmatrix}^\intercal\\
\pi_{3,7}^1&=  
\begin{bmatrix}
 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1
\end{bmatrix}^\intercal
\end{align*}
\caption{The maps of representations $\pi_{e,2n+r}^r:M_{2n + r}^r \twoheadrightarrow U_{2n+r}^r$ for $e = 3$.}
\label{Composition series e=3}
\end{figure}

\begin{figure}[h]
\begin{align*}
  \pi_{4,4}^2&=  
\begin{bmatrix}
  1 & \qqo & 1
\end{bmatrix}^\intercal\\
\pi_{4,4}^1&=  
\begin{bmatrix}
  1 & \frac{1}{2}\qqo & \frac{1}{2}\qqo & 1 & \frac{1}{2}\qqo
\end{bmatrix}^\intercal\\
\pi_{4,6}^2 &=
\begin{bmatrix}
  1& 0& 0& 0& 0& 0& \qqo & 1& 0\\
  0& 1& 0& 0& 0& -1& -1& -\qqo &0\\
  0& 0& 1& 0& 0& \qqo& 0 &1 &0\\
  0& 0& 0& 1& 0& 0& -1& -\qqo &-1\\
  0& 0& 0& 0& 1& \frac{1}{2}\qqo& \frac{1}{2}\qqo &1 &\frac{1}{2}\qqo
\end{bmatrix}^\intercal\\
\pi_{4,6}^0 &=
\begin{bmatrix}
  \qqo & 1 & 1 & \qqo & 1
\end{bmatrix}^\intercal\\
\pi_{4,7}^1 &=
\begin{bmatrix}
1& 0& 0& 0& \frac{1}{2}\qqo& 0& -\frac{1}{2}& -\frac{1}{2}& 0& 0& 1& 1/2& \qqo& \frac{1}{2}\\
0& 1& 0& 0& -1& 0& \frac{1}{2}\qqo& \frac{1}{2}\qqo& 1& 0& \qqo& \frac{1}{2}\qqo& -1 & \frac{1}{2}\qqo\\
0& 0& 1& 0& -1& 0& 0& \qqo& 1& 0& -\qqo& 0& -1& 0\\
0& 0& 0& 1& \qqo& 0& 0& -1& -\qqo& 0& 1& -1& 0& 0\\
0& 0& 0& 0& 0& 1& \frac{1}{2}\qqo& \frac{1}{2}\qqo& 0& 0& 0& \frac{1}{2}\qqo& -1& \frac{1}{2}\qqo\\
0& 0& 0& 0& 0& 0& 0& 0& 0& 1& -\frac{1}{2}\qqo& \frac{1}{2}\qqo& 1& \frac{1}{2}\qqo
\end{bmatrix}^\intercal\\
\end{align*}
\caption{The maps of representations $\pi_{e,2n+r}^r:M_{2n + r}^r \twoheadrightarrow U_{2n+r}^r$ for $e = 4$.}
\label{Composition series e=4}
\end{figure}

\begin{figure}[h]
\begin{align*}
  \pi_{5,5}^3 &=
\begin{bmatrix}
  1 & \tto & \tto & 1
\end{bmatrix}^\intercal\\
\pi_{5,6}^2 &=
\begin{bmatrix}
  \tto& \tto& 1& 1& \tto+1& \tto & \tto & \tto & 1
\end{bmatrix}^\intercal\\
\pi_{5,7}^3 &=
\begin{bmatrix}
1& \tto& \tto& 1& 0& 0& 0& 0& 0& 0& 0& 0& 0& 0\\
0& 0& 0& 0& 1& \tto& \tto& 1& 0& 0& 0& 0& 0& 0\\
1& 0& 0& 0& 0& 0& 0& 0& \tto& \tto& 1& 0& 0& 0\\
\tto& \tto& 1& 0& 1& \tto+1& \tto& 0& \tto& \tto& 0& 1& 0& 0\\
-\tto-1& -\tto& -\tto& 0& -\tto& -\tto-1& -\tto-1& 0& -\tto-1& -\tto-1& 0& 0& 1& 0\\
0& 0& 1& 0& 0& 0& \tto& 0& 0& \tto& 0& 0& 0& 1
\end{bmatrix}^\intercal\\
\pi_{5,7}^1 &=
\begin{bmatrix}
1& \qqo& \qqo& \qqo& -\tto& \qqo& -\tto& \qqo& -\tto& 1& \qqo& \qqo& \qqo& -\tto
\end{bmatrix}^\intercal\\
\pi_{6,6}^4 &=
\begin{bmatrix}
  1 & \vtv & 2 & \vtv & 1 
\end{bmatrix}^\intercal\\
\pi_{6,7}^3 &=
\begin{bmatrix}
  \vtv & 2 & \vtv & 1 & 1 & 2\vtv & 3 & \vtv & 2 & 2\vtv & 2 & 2 & \vtv & 1
\end{bmatrix}^\intercal\\
\pi_{6,7}^5 &=
\begin{bmatrix}
  1 & \fft & \fto & \fto & \fft & 1
\end{bmatrix}^\intercal
\end{align*}
\caption{The maps of representations $\pi_{e,2n+r}^r:M_{2n + r}^r \twoheadrightarrow U_{2n+r}^r$ for $e = 5,6$.}
\label{Composition series e=6}
\end{figure}

\newpage
\appendix
\section{Compatibility of Representations with the Relations}
In general, we define representations above for the free algebra on generators $\cbr{T_i}$.
Recall that we may give a presentation of $\SH$ having generators $T_i$ and relations
\begin{align}
  (T_i - q)(T_i + 1) &= 0 \label{quadratic}\\
  T_iT_{i+1}T_i &= T_{i+1}T_iT_{i+1} \label{braid1}\\ 
  T_iT_j &= T_jT_i \hspace{40pt} \abs{i - j} > 1. \label{braid2}
\end{align}
We call \eqref{quadratic} the \emph{quadratic relation} and \eqref{braid1}, \eqref{braid2} the \emph{braid relations}.
It is easily seen that a representation of $\SH$ is equivalent to a representation of the free algebra $k\langle T_i \rangle$ which acts as 0 on the relations (henceforth referred to as \emph{compatibility} with the relations).
We will prove in the following sections that $V$ and $W$ are compatible with the Hecke algebra relations.

\subsection{Explicit Definition of $M$}
We will give a more careful definition of the crossingless matchings representation here.
\begin{definition}
  A \emph{crossingless matching on $2n+r$ indices with $r$ anchors} is a partition of $\cbr{1,\dots,2n+r}$ into $n$ parts of size $2$ and $r$ of size 1 such that no two parts $(a,a')$ and $(b,b')$ satisfy $a < b < a' < b'$, and no parts $(c), (a,a')$ satisfy $a < c < a'$.
  We will call these arcs and anchors, respectively.
  Then, define $M^r_{2n+r}$ to be the $k$-vector space with basis the set of crossingless matchings on $2n+r$ indices with $r$ anchors.

  In order to endow $M^r_{2n + r}$ with an $\SH$-action, consider some basis element $w_j$ and some element $(1 + T_i)$ of $\SH$.
  The elements $\cbr{1} \cup \cbr{1 + T_i | 1 \leq i < 2n + r}$ generate $\SH$, so it is sufficient to define the action of $1 + T_i$ on $w_j$.

  If $w_j$ has arc $(i,i+1)$, define $(1 + T_i)w_j := (1 + q)w_j$.
  If $w_j$ has anchors $W(i) = i$ and $W(i+1) = i+1$, define $(1 + T_i)w_j := 0$.
  If $w_j$ has anchor $W(i) = i$ and arc $W(i+1) = b$, define $(1 + T_i)w_j := q^{1/2}w_l$, where $w_l(i) = i+1$, $w_l(b) = b$, and all other arcs agree with $w_j$.
  If $w_j$ has arcs $W(i) = a$ and $W(i+1) = b$, then define $(1 + T_i)w_j := q^{1/2}w_l$, where $w_l(i) = i+1$,$w(a) = b$, and all other acts agree with $w_j$. 
  We verify that this is well-defined in Section \ref{Cross Relations}.
\end{definition}

We may alternately sharpen our topological definition;
\begin{definition}
Fix $2n + r$ distinct points $a_1,\dots,a_{2n + r}$ points along $\RR \times \cbr{0} \subset \RR^2$ and $r$ distinct points $b_1,\dots,b_r$ along $\RR \times \cbr{1}$.
Then, define $M_{2n + r}^r$ to have basis given by the isotopy classes of $n + r$ paths connecting the points $a_1,\dots,a_{2n+r},b_1,\dots,b_r$ such that no distinct $b_i,b_j$ are connected by a path.

We will take some basis element $w_j \in M_{2n +r}^r$ and define the action $(1 + T_i)w_j$.
To do so, map $w_j$ through the natural embedding $\RR \times \brk{0,1} \hookrightarrow \RR \times \brk{\frac{1}{2},1}$, and form the figure $w_j^i$ by adjoining the lines connecting $a_l$ and $a_l + \prn{0,\frac{1}{2}}$ for all $l \neq i,i+1$ as well as paths from $a_i$ to $a_{i+1}$ and $a_i + \prn{0,\frac{1}{2}}$ to $a_{i+1} + \prn{0,\frac{1}{2}}$.
This has either 0 or 1 path components which do not intersect $\RR \times \cbr{0,1}$;
these form ``loops.''

Take the figure $\tilde w_j^i$ without this component.
If $\tilde w_j^i$ is not isotopic to some $w_l$, then define $(1 + T_i)w_j := 0$.
If $\tilde w_j^i$ is isotopic to some $w_l$, define $(1 + T_i)w_j := (1 + q)w_l$ if $w_j^i$ has a loop and $(1 + T_i)w_j := q^{1/2}w_l$ otherwise.
This process is illustrated in Figure \ref{Action}.
\end{definition}

Let the length of an arc $(i,j)$ be $l(i,j) := j - i + 1$.
Note that the crossingless matchings on $2n$ indices with no anchors can all be identified with a list of $n$ integers describing the lengths of the arcs from left to right;
  This basis is illustrated for $M_{5}^1$ in Figure \ref{S5 Basis}.

\begin{remark}
  This definition gives a graphical calculus for working with our module.
  It should be clear that, if $w_j^i$ has a loop then $w_l(i) = i+1$ and $w_l = w_j$.
  Further, this easily defines an arbitrary composition:
  \[
    (1 + T_{i_1})\cdots(1 + T_{i_\ell})w_j = q^{(\ell - t)\frac{1}{2}}(1 + q)^tw_l
  \]
  if the figure we make via $(1+T_{i_1})\cdots(1+T_{i_\ell})$ is isotopic to $w_l$ after removing $t$ loops.
\end{remark}

  Note that the crossingless matchings on $2n$ indices with no anchors can all be identified with a list of $n$ integers describing the lengths of the arcs from left to right;
  call this the \emph{increasing lexicographical order basis}.
  Further, we have a surjection $M_{2n + 2r}^0 \twoheadrightarrow M_{2n + r}^r$ which takes basis elements to basis elements;
  this induces an order on the basis for $M_{2n +r}^r$, which we will henceforth refer to as the \emph{induced lexicographical order basis.}
  This basis is illustrated for $M_{5}^1$ in Figure \ref{S5 Basis}.

\subsection{Verifying the Crossingless Matchings Representations}
\label{Cross Relations}
Take some basis vector $w_i$.
We will first check \eqref{quadratic} by case work:
\begin{itemize}
  \item Suppose there is an arc $(i,i+1)$.
    Then, $(T_i-q)(T_i + 1)w = (1 + q)\brk{(1 + T_i)w - (1 + q)w} = 0$, giving \eqref{quadratic}.

  \item Suppose there is no arc $(i,i+1)$ and $i,i+1$ do not both have anchors;
    then $(T_i +  1)w = q^{1/2}w''$ for some basis vector $w'$ having arc $(i,i+1)$, and the computation follows as above for \eqref{quadratic}.
  \item Suppose $i,i+1$ are anchrors;
    then $(T_i + 1)w = 0$, giving \eqref{quadratic}.
\end{itemize}

\vspace{5pt}
Now we verify \eqref{braid1}.
Let $h := (1 + T_i)(1 + T_{i+1})(1+T_i)$, and let $g := (1 + T_{i+1})(1 + T_i)(1 + T_{i+1})$.
Note the following expansion:  
  \begin{align*}
      hw
      &= 1 + 2T_i + T_i^2 + T_{i+1} + T_iT_{i+1} + T_{i+1}T_i + T_iT_{i+1}T_i\\
      &= 1 + (1+q)T_i + T_{i+1} + T_iT_{i+1} + T_{i+1}T_i + T_iT_{i+1}T_i.
    \end{align*}
    An analogous formula gives an analogous equality in $g$.
    Hence we have
    \[
      (h-g)w = q(T_i - T_{i+1}) + T_iT_{i+1}T_i - T_{i+1}T_iT_{i+1}.
    \]
    Hence we may equivalently check that $(h-g)w = q(T_i - T_{i+1})$.
    In fact, $hw = q(1 + T_i)$ and $gw = q(1 + T_j)$ by Figure \ref{braid1arc}, giving compatibiliy.
    \begin{figure}
\[
  \begin{gathered}\EmptyNAction{3}{1/1,2/2,1/3}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\EmptyNAction{3}{1/2}\end{gathered}
  \hspace{50pt}
  \begin{gathered}\EmptyNAction{3}{2/1,1/2,2/3}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\EmptyNAction{3}{2/1}\end{gathered}
  \hspace{50pt}
  \begin{gathered}\EmptyNAction{4}{1/1,3/2}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\EmptyNAction{4}{3/1,1/2}\end{gathered}
\]
  \caption{
    The following give visual intuition for isotopy between $(1 + T_i)(1 + T_{i-1})(1 + T_i)w_j$ and $q(1 + T_i)$,between $(1 + T_{i+1})(1 + T_i)(1 + T_{i+1})w_j$ and $q(1 + T_i)$, and between $(1 + T_i)(1+T_j)w_l$ and $(1 + T_j)(1 + T_i)w_l$.
    In all cases, these isotopies imply equality between the elements listed.
  }
  \label{braid1arc}
\end{figure}

Lastly, we have the equation
\[
  (1 + T_i)(1 + T_j) - (1 + T_j)(1 + T_i) = T_iT_j - T_jT_i
\]
and hence we simply need to verify that $(1 + T_i)$ and $(1 + T_j)$ commute, as shown in Figure \ref{braid1arc}

\subsection{Verifying the Fibonacci Representations} 
\label{Fib Relations}
Similar to before, the reader may verify that \eqref{braid2} follows easily, and the others may be verified on strings of length 3 and 4.
By considering the coefficients in order of \eqref{Fib Action}, the quadratic relation \eqref{quadratic} gives the following quadratics:
\begin{equation}
  \begin{split}
    (\vara - q)(\vara + 1) &= 0\\
    (\varb - q)(\varb + 1) &= 0\\
    \varc\vard + \vard\vare &= (q-1)\vard\\
    \varc^2 + \vard^2 &= (q -1)\varc + q\\
    \vare^2 + \vard^2 &= (q -1)\vare + q
  \end{split}
\end{equation}
The first two of these are easily verified for any $q$.
Since $\vard \neq 0$, the third is equivalently given by
\[
  (q - 1) = \varc + \vare = t(q\tau - 1 + q - \tau) = (\tau^2 + \tau)(q - 1)
\]
or that $\prn{\tau^2 + \tau - 1}(q-1) = 0$.
One may verify that \[\tau^2 + \tau - 1 = q^6 + 2q^5 + q^4 + q^3 + q^2 - 1 = (-1+q+q^2)\brk{5}_q = 0.\]

The fourth is given by the quadratic
\[
  \tau^2\brk{(q\tau-1)^2 - \tau(q+1)} = \tau(q-1)(q\tau - 1) + q
\]
or equivalently,
\[
  (\tau^2 + \tau - 1)\brk{q\prn{qt^2 + 1} + t} = 0
\]
which is true for every $q$.

The fifth is similarly given by
\[
  (\tau^2 + \tau - 1)\brk{q\prn{qt + 1} + t^2} = 0 
\]
which is true for every $q$.

\vspace{7pt}
We now verify \eqref{braid1}.
We may order the basis for $V^4$ as follows:
\[
  \cbr{(\vp\vp\vp\vp),(\vs\vp\vp\vs),(\vp\vp\vp\vs),(\vs\vp\vp\vp),(\vs\vp\vs\vp),(\vp\vs\vp\vs),(\vp\vp\vs\vp),(\vp\vs\vp\vp)}.
\]
Then, in verifying the braid relation \eqref{braid1} in this order, we encounter the following quadratics (with tautologies and repetitions omitted):
\begin{align*}
    \vara\vare^2 + \varb\vard^2 &= \vara^2 \vare\\
    \vara\vard\vare + \varb\varc\vard &= \vara\varb\vard\\
    \varb\varc^2 + \vara\vard^2 &= \varb^2\varc\\
    \vara\varc^2 + \vard^2\vare &= \vara^2\varc\\
    \vard\vare^2 + \vara\varc\vard &= \vara\vard\vare
\end{align*}
Substituting in $\tau$ and dividing by $\delta$ whenever possible, these are equivalent to the vanishing of the following polynomials in $q$:
\begin{align*}
  -q (1 + q) (1 + q^2 + q^3) (2 + q + 3 q^2 + 2 q^3) \brk{5}_q &= 0\\
  (1 + 2 q + q^3 + q^4) \brk{5}_q &= 0\\
  (1+q)^2 (1+q^2+q^3) (1+3q^3 - q^4 + q^6)\brk{5}_q &= 0\\
  (1+q)^2 (1+q^2+q^3) (1+5q+5q^2+3q^3+3q^4+3q^5+q^6)\brk{5}_q &= 0\\
  (1+q) (1+q^2+q^3) (-1+2q+q^2+q^3+q^4)\brk{5}_q &= 0.
\end{align*}
Notably, each of these vanish when $e = 5$.

\section{Conjugate Subalgebras}\label{Algebra}
Throughout the text, for some representation $V$, we refer to $\Res_{\SH(S_{l})}^{\SH(S_m)} V$ without specifying exactly which subalgebra $\SH(S_{l})$.
\begin{proposition}
  Suppose $B,B'$ are subalgebras of the $k$-algebra $A$ with $B = uB'u^{-1}$, and let $V$ be a representation of $A$.
  Then, the linear isomorphism $V \xrightarrow{\phi} V$ given by $v \mapsto uv$ causes the following to commute for any $b \in B$:
  \[
    \begin{tikzcd}
      V \arrow[r,"\phi"] \arrow[d,"b"] & V \arrow[d,"ubu^{-1}"]\\
      V \arrow[r,"\phi"] & V
    \end{tikzcd}
  \]
  Hence, through the identification of $B$ and $B'$ via conjugation, we have $\Res_{B}^A V \simeq \Res_{B'}^A V$
\end{proposition}
\begin{proof}
  This is simply given by $(ubu^{-1})uv = ubv$.
\end{proof}

\begin{corollary}
  Suppose $\SH',\SH''$ are two subalgebras of $\SH(S_m)$ generated by $l$ reflections and $V$ is a representation of $\SH$.
  Then, $\Res_{\SH'}^{\SH} V \simeq \Res_{\SH''}^\SH V$.
\end{corollary}
\begin{proof}
  Let $\SH'$ and $\SH''$ be the subalgebras of $\SH(S_m)$ generated by the reflections $\cbr{T_{i_1},\dots,T_{i_l}}$ and $\cbr{T_{i_1},\dots,T_{i_{j-1}},T_{i_j + 1},T_{i_{j+1}},\dots,T_{i_l}}$ for $1 \leq i_1 < \dots < i_{j-1} < i_j + 1 < i_{j+1} < \dots < i_l \leq n$.
  It is sufficient to prove that $\SH'$ and $\SH''$ are conjugate;
  then transitivity gives conjugacy of any $S_l \subset S_m$, and the previous proposition gives isomorphisms of the representations.

  In fact, the reader can verify that $\SH'' = T_{i_j}\SH'T_{i_j}^{-1}$.
\end{proof}

\bibliographystyle{plain} 
 \bibliography{../RepBib}
\end{document}