### 2019-07-30 tex/Drafts/Draft-2.tex
incorporated kernel stuff

  \documentclass{amsart}   
  \usepackage{../RepStyle} 
  % Colors links
  \hypersetup{
  colorlinks,
  citecolor=blue,
  linkcolor=orange!50!red,
  urlcolor=blue}

  \usepackage{../NatMacros}

  % Emulates \subsection* but doesn't add to ToC (I don't know why amsart is weird 
  \newcommand{\fakesubsection}[1]{
    \vspace{7pt}
    \noindent \textbf{#1.}
  }

  \makeatletter
  %Table of Contents
  \setcounter{tocdepth}{3}

  % Add bold to \section titles in ToC and remove . after numbers
  \renewcommand{\tocsection}[3]{%
    \indentlabel{\@ifnotempty{#2}{\bfseries\ignorespaces#1 #2\quad}}\bfseries#3}
  % Remove . after numbers in \subsection
  \renewcommand{\tocsubsection}[3]{%
    \indentlabel{\@ifnotempty{#2}{\ignorespaces#1 #2\quad}}#3}
  \let\tocsubsubsection\tocsubsection% Update for \subsubsection
  %...

  \newcommand\@dotsep{4.5}
  \def\@ocline#1#2#3#4#5#6#7{\relax
    \ifnum #1>\c@tocdepth % then omit
    \else
      \par \addpenalty\@secpenalty\addvspace{#2}%
      \begingroup \hyphenpenalty\@M
      \@ifempty{#4}{%
        \@tempdima\csname r@tocindent\number#1\endcsname\relax
      }{%
        \@tempdima#4\relax
      }%
      \parindent\z@ \leftskip#3\relax \advance\leftskip\@tempdima\relax
      \rightskip\@pnumwidth plus1em \parfillskip-\@pnumwidth
      #5\leavevmode\hskip-\@tempdima{#6}\nobreak
      \leaders\hbox{$\m@th\mkern \@dotsep mu\hbox{.}\mkern \@dotsep mu$}\hfill
      \nobreak
      \hbox to\@pnumwidth{\@tocpagenum{\ifnum#1=1\bfseries\fi#7}}\par% <-- \bfseries for \section page
      \nobreak
      \endgroup
    \fi}

  \def\l@subsection{\@tocline{2}{0pt}{2.5pc}{5pc}{}}
  \makeatother

  \begin{document}

  \title[Some Graphical Realizations of Two-Row Specht Modules of Hecke Algebras]{Some Graphical Realizations of Two-Row Specht Modules of Iwahori--Hecke Algebras of the Symmetric Group}
  \author[Miles Johnson and Natalie Stewart]{Miles Johnson and Natalie Stewart\\
    Mentor : Oron Propp\\
  Project Suggested by : Roman Bezrukavnikov\\ \; \\
  July 31, 2019
  }

   \begin{titlepage}
    \maketitle
    \begin{abstract}
      We consider the Iwahori--Hecke algebra of the symmetric group on $2n + r$ letters with parameter $q$.
      Let $e$ be the smallest positive integer such that the $q$-number $\brk{e}_q = 0$, or set $e = \infty$ if none exist.
      We modify Khovanov's crossingless matchings to include $2n$ ``nodes'' and $r$ ``anchors,'' and prove in the case $e > n + r + 1$ that the associated module is isomorphic to the Specht module $S^{(n+r,n)}$ which corresponds to the partition $(n + r,n) \vdash 2n + r$.
      We then give heuristics in support of the general case, including explicit composition series for $e = n + r + 1$ and for $2n + r \leq 7$. 
      Lastly, when $e = 5$, we prove an isomorphism between the irreducible quotient $D^{(n+r,n)}$ with $r \leq 3$ and some subrepresentations of Jordan--Shor's Fibonacci representation.
      We provide explicit transition matrices between this representation and the crossingless matchings representation for $2n + r \leq 6$.
    \end{abstract}

  \begingroup
  \hypersetup{linkcolor=black} % no colored hyperlinks in toc
  \tableofcontents
  \endgroup

  \end{titlepage}

\section{Introduction} 
  Let $S_{2n+r}$ be the symmetric group on $2n+r$ letters with $2n + r \geq 2$, let $\SH := \SH_{k,q}(S_{2n+r})$ be the corresponding Hecke algebra over a field $k$ with parameter $q \in k^\times$ having a fixed square root $q^{1/2}$, and let $\cbr{T_i}$ be the simple reflections generating $\SH$.

  Let $\brk{m}_q = 1 + q + \dots + q^{m-1}$ be the $q$-number of $m$. 
  Let $e$ be the smallest positive integer such that $\brk{e}_q = 0$, and set $e = \infty$ if no such integer exists.
  Note that either $q = 1$ and $e$ is the characteristic of $k$ (with $0$ replaced by $\infty$), or $q \neq 1$ and $q$ is a primitive $e$th root of unity.

  \vspace{5pt}
  When $q = 1$, the Hecke algebra $\SH$ is isomorphic to the group algebra $k\brk{S_{2n+r}}$;
  hence the representation theory of $\SH$ generalizes the representation theory of the symmetric group.
  The Hecke algebra is also well known to be connected to the representation theory of the general linear group over a finite field.\cite{Mathas-book}

  It is a classical result that $\SH$ is semisimple precisely when $e > 2n + r$, in which case the irreducible representations of $\SH$ are given by \emph{Specht modules} $S^\lambda$, which are indexed by the partitions $\lambda$ of $2n + r$.
  Further, $\SH$ admits a cellular basis with cell modules given by $S^\lambda$.
  In particular, these admit quotients $D^\lambda$ such that the set $\cbr{D^\lambda \mid D^\lambda \neq 0, \lambda \vdash n}$ is a complete set of pairwise-nonisomorphic irreducible $\SH$-modules.
  This set is indexed by the partitions $\lambda \vdash n$ which are $e$-restricted.\cite{Murphy1,Murphy2}

  These representations $D^\lambda$ have explicit constructions, but many of their properties are unknown.
  For instance, the dimension of these modules is unknown outside of some special cases.\cite{Mathas-book}
  However, there does exist an algorithm due to Lascoux--Leclerc--Thibon--Ariki which computes the decomposition matrices of the Iwahori--Hecke Algebra $\SH_{\CC,q}(S_{2n + r})$ for $q$ an $e$th root of unity.\cite{LLT,Ariki}

  The cellular basis for $S^\lambda$ and associated construction for $D^\lambda$ tend to be complicated and often computationally intractable.
  We aim to give simple graphical realizations of $S^\lambda$ and $D^\lambda$ in some cases that $\lambda = (n+r,n)$.
  These realizations expose the structure of irreducibility and branching in an intuitive and computationally simple way.

  \vspace{5pt}
  Note that we follow the convention of Murphy and Kleshchev, which is dual to the conventions of Dipper, James, and Mathas;
  one may translate our results to the latter convention by transposing all partitions.\cite{Kleshchev,Mathas-book,Murphy1}

  Throughout this paper, we analyze the \emph{two-row partitions} $(n+r,n) \vdash 2n + r$ and their corresponding modules $S^{(n+r,n)}$ and $D^{(n+r,n)}$.

  \fakesubsection{Crossingless matchings}
  The following definition modifies the crossingless matchings defined by Khovanov.\cite{Khovanov}

  \begin{definition}
  Define a \emph{crossingless matching on $2n + r$ nodes and $r$ anchors} to be an isotopy class of $n + r$ nonintersecting paths in the slice $\RR \times \brk{0,1}$ connecting $2n + r$ disctinct points of $\RR \times \cbr{0}$ and $r$ points of $\RR \times \cbr{1}$ such that none of the latter points are conected. 
  Let $M_{2n + r}^r$ have basis given by these matchings.
  This is illustrated in Figure \ref{S5 Basis}.

  Let the length of an arc $(i,j)$ be $j - i + 1$.

  We endow $M_{2n + r}^r$ with an action by specifying $(1 + T_i)w_j$ for some basis element $w_j$ of $M_{2n + r}$.
  We do so by concatenating in ``vertical lines'' below each point other than the $i$th and $i+1$st, concatenating paths between the $i$th and $i+1$st points as well as points below them, removing any ``loops'' this forms, and taking isotopy to some matching $w_l$ if possible;
  if this is not possible, then there are anchors at $i,i+1$ and we set $(1 + T_i)w_j := 0$;
  if this is possible and there is a ``loop,'' set $(1 + T_i)w_j := (1 + q)w_j$ and otherwise set $(1 + T_i)w_j := q^{1/2}w_l$.
  This action is illustrated in Figure \ref{Action}, and we verify that this is well-defined in Appendix \ref{Cross Relations}.
  \end{definition}
  \begin{figure} 
    \def\cbasisspacing{5mm}
    \begin{adjustbox}{width=\textwidth}
    $\cbr{
      \begin{gathered}
        \GeneralizedMatching{5}{1/2, 3/4}{1}{5/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{1/4, 2/3}{1}{5/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{1/2, 4/5}{1}{3/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{2/3, 4/5}{1}{1/1}{3/4}, \hspace{\cbasisspacing}
        \GeneralizedMatching{5}{2/5, 3/4}{1}{1/1}{3/4}, \hspace{\cbasisspacing}
       \end{gathered}}$ 
     \end{adjustbox}
       \caption{The basis for $M_5^1$.}
    \label{S5 Basis}
   \end{figure}

  \begin{figure}
    \begin{adjustbox}{width=\textwidth}
      \begin{tabular}{l l}
        \GeneralizedAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{2}{1/4, 2/3}{5/1,6/2}{(1+q)}
        \hspace{20pt}
        &
        \GeneralizedAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{3}{1/2, 3/4}{5/1,6/2}{q^{1/2}}\\
        \GeneralizedAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{4}{2/3, 4/5}{1/1,6/2}{q^{1/2}}
        &
        \GeneralizedZeroAction{6}{1/4,2/3}{2}{5/1,6/2}{1}{5}
      \end{tabular}
  \end{adjustbox}

    \caption{Illustration of the actions $(1 + T_i)w_{\abs{M^2_6}}$.
      In general, we act by deleting loops, isotoping onto a new crossingless matching, and scaling by either $q^{1/2}$, $(q + 1)$, or 0.
    }
    \label{Action}
   \end{figure}

  Note that the representations $M_{0 + r}^r$ and $S^{(r)}$ are isomorphic to the sign representation.
  In fact, we will prove that $M_{2n + r}^r \simeq S^{(n+r,n)}$ whenever $e > n + r + 1$.
  Additionally, we will prove that $M_{2n + r}^r$ is irreducible everywhere that $e > n$ and $S^{(n+r,n)}$ is irreducible.
  We will conjecture that $M_{2n + r}^r \simeq S^{(n+r,n)}$ with $e$ unrestricted as well.

  This proves a graphical characterization of $S^\lambda$ in many cases, and hence $D^\lambda$ in the cases where $S^\lambda \simeq D^\lambda$.
  However, the representation theory of $\SH$ is tied to the modules $D^\lambda$ even in cases where $S^\lambda$ is not irreducible;
  we will present a characterization of the module $D^{(n+r,n)}$ when $e = 5$ and $r \leq 3$.

  \def\vara{\alpha_1}
  \def\varb{\alpha_2}
  \def\varc{\varepsilon_1}
  \def\vard{\delta}
  \def\vare{\varepsilon_2}
  \def\vs{\texttt{*}}
  \def\vp{\texttt{0}}
  \fakesubsection{Fibonacci representation}
  Now suppose that $e = 5$ and $k$ contains the algebraic number $\sqrt{q + q^4}$ for reasons which will be apparent soon.
  For convenience set $m := 2n + 4$.
  The following is a modification of Shor--Jordan's Fibonacci representation of the braid group.\cite{Shor}

  \begin{definition}
  Let $V^m$ be the $k$-vector space with basis given by the strings $\cbr{\vs,\vp}^{n+1}$ such that the bit $\vs$ never appears twice consecutively. 
  We will refer to $V^m$ as the \emph{Fibonacci representation} and suppress the superscript whenever it is clear from context.

  We wish to endow this with a $\SH$-action which acts on a basis vector only dependent on bits $i,i+1,i+2$, sending each basis vector to a combination of the other basis vectors having the same bits $1,\dots,i,i+2,\dots,n+1$ as follows:

  \begin{equation} 
    \begin{split}
      T_1 \, \prn{\vs\vp\vp} &:= \vara\prn{\vs\vp\vp}\\
      T_1 \, \prn{\vp\vp\vs} &:= \vara\prn{\vp\vp\vs}\\
      T_1 \, \prn{\vs\vp\vs} &:= \varb\prn{\vs\vp\vs}\\
      T_1 \, \prn{\vp\vs\vp} &:= \varc\prn{\vp\vs\vp} + \vard\prn{\vp\vp\vp}\\
      T_1 \, \prn{\vp\vp\vp} &:= \vard\prn{\vp\vs\vp} + \vare\prn{\vp\vp\vp}
  \end{split} \label{Fib Action} 
  \end{equation}
  for constants
  \begin{equation}
    \begin{split}
    \tau  &:= q + q^4\\
    \vara &:= -1\\
    \varb &:= q\\
    \varc &:= \tau(q\tau - 1)\\
    \vard &:= \tau^{3/2}(q + 1)\\
    \vare &:= \tau(q-\tau)
  \end{split} \label{Fib Constants} 
  \end{equation}
  with $T_i$ acting similarly on the substring $i,i+1,i+2$.
  We will verify that this is a representation of $\SH$ in Appendix \ref{Fib Relations}
  \end{definition}

  This contains four subrepresentations spanned by strings with beginning and ending with specified bits.
  Label the subrepresentation spanned by strings $(\vs\dots\vs)$ by $V_{\vs\vs}$, $V_{\vs\vp}$, $V_{\vp\vs}$, and $V_{\vp\vp}$;
  these bits are preserved by the action \eqref{Fib Action}.
  We will show that these are isomorphic to particular $D^\lambda$.

  \fakesubsection{Overview of paper}

  In Section \ref{Specht Modules Section} we give corollaries to standard theorems concerning Specht modules.
  First, James-Mathas provides a sharp characterization of the irreducibility $S^\lambda$ for $\lambda \vdash 2n + r$ which is $e$-regular, called the \emph{Carter criterion}.\cite{Mathas-book}
  We specialize this to the case that $\lambda = (n+r,n)$ to give a combinatorial condition for irreducibility of $S^{(n+r,n)}$.
  We note that this irreducibility depends only on $e$ when $e > n$;
  otherwise it depends on both $e$ and the characteristic of the field.
  Further, we use Kleshchev--Brundan's modular branching rules to prove our first significant statement: if $S^\lambda \simeq D^\lambda$ and $e > n$, then a particular length-2 composition series uniquely determines $\lambda$;
  further, an irreducible restriction to $D^{(n,n-1)}$ determines $\lambda$ as well.\cite{Kleshchev, Brundan}

  \def\thmcolor{black!60!orange}
  \vspace{5pt}
  In Section \ref{Crossingless Matchings Section}, we first prove that every basis vector in $M$ is cyclic, then that $M$ contains no sign subrepresentation.
  Then, we prove the following theorem.
  \begingroup
  \hypersetup{linkcolor=\thmcolor} % no orange theorem label
  \begin{customthm}{\ref{Irreducibility Theorem}}
    Suppose $e > n$ and $S^{(n+r,n)}$ is irreducible.
    Then $M_{2n + r}^r$ is irreducible.
  \end{customthm}
  \endgroup
  Following this, we prove the existence of a particular filtration with factors given by other crossingless matchings representations;
  using irreducibility, this becomes a composition series, and an inductive argument combined with the branching of Section \ref{Specht Modules Section} allow us to prove the following:
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{Correspondence Theorem}}
    Suppose $e > n + r + 1$.
    Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.
  \end{customthm}
  \endgroup
  Last, we finish the section by proving more general statements concerning sign subrepresentations of $M$, which contribute to conjectures later in the paper.{\color{orange} Fill in with specific information later.}

  \vspace{5pt}
  In Section \ref{Fibonacci Section}, we begin by establishing isomorphisms as in \ref{Fibonacci Theorem} for the subrepresentations of $V^2$, as well as irreducibility of $V^3_{\vs\vp}$.
  We then use these cases to prove that $V^m_{\vs\vp}$ is irreducibile for all $m$, which implies that that $V^m_{\vs\vs}$ is irreducible.
  From this, we inductively prove the following theorem:
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{Fibonacci Theorem}}
    The following isomorphisms characterize the subrepresentations of the Fibonacci representation:
    \begin{align*}  
       V^{2n}_{\vs\vs} &\simeq D^{(n,n)}\\ 
       V^{2n-1}_{\vs\vs} &\simeq D^{(n+1,n-2)}\\
       V^{2n}_{\vs\vp} &\simeq D^{(n+1,n-1)}\\
       V^{2n-1}_{\vs\vp} &\simeq D^{(n,n-1)}.
    \end{align*}
  \end{customthm}
  \endgroup
  {\color{orange} Overview of Conjecture and Empirics goes here.}

  % Acknowledgements should be given section-level heading in toc, subsection-level in text.
  \addcontentsline{toc}{section}{\textbf{Acknowledgements}}
  \fakesubsection{Acknowledgements}
  The authors thank Prof. Roman Bezrukavnikov for suggesting this project, as well as Dr. Slava Gerovitch for organizing the SPUR+ program.
  We would also like to thank Profs. David Jerison and Ankur Moitra for their role in SPUR+ as well as general advice.
  We would also thank Professor Alexander Kleshchev for our helpful conversations concerning branching theorems.
  Lastly, we would like to express our gratitude to our mentor Oron Propp for his help and advice in both acquiring background knowledge and in executing the mathematics in this paper;
  this project would not be possible without him.

\section{Preliminaries on Specht modules}\label{Specht Modules Section}
  For this section and the rest of the paper, assume $n > 0$.

  Throughout the rest of the paper, it will be useful to have precise notation;
  identify each partition with a tuple $\lambda = (\lambda_1^{a_1},\dots,\lambda_l^{a_l})$ having $\lambda_i > \lambda_{i+1}$, $a_i > 0$, and $\sum_i a_i\lambda_i = 2n + r$.
  Identify each of these with a subset $\brk \lambda \subset \NN^2$ as in \cite{Kleshchev}, and define $\lambda(i) = (\lambda_1^{a_1},\dots,\lambda_{i-1}^{a_{i-1}},\lambda_i^{a_i - 1},\lambda_i-1,\lambda_{i+1}^{a_{i+1}},\dots,\lambda_l^{a_l})$ to be the partition with the $i$th row removed.
  Say that $\lambda$ is \emph{$e$-regular} if $\lambda_i - \lambda_{i+1} < e$ for all $i$ and $\lambda_l < e$.

  In the following subsection, we cite a theorem of James--Mathas which precisely characterizes the irreducibility of $S^\lambda$ in the case that $\lambda$ is $e$-regular, and we specialize this result to the case of two-row Specht modules.
  This falls into two cases: either $e > n$, where $S^{(n+r,n)}$ is irreducible iff $e \nmid r+2,\dots,n+r+1$, or $e \leq n$, where the irreducibility of $S^{(n+r,n)}$ is more complicated and depends also on the characteristic of $k$.
  We will focus primarily on the former case.

  Following this, we reproduce the branching theorems of Kleshchev--Brundan, which allow us to fully characterize the socle of $\Res D^{\lambda}$.
  This and some combinatorial arguments yield the main result of this section, which allows us to determine certain $D^{(n+r,n)}$ via their composition series.
  This will be instrumental later for characterizing the crossingless matchings representation $M_{2n + r}^r$ as a Specht module, and it will extend to all cases with $e > n + r + 1$.

  \subsection{Irreducibility of $S^\lambda$}
  Let $\ell$ be the characteristic of $k$;
  then, set
  \[
    p := \begin{cases}
      \ell & \text{if }\ell > 0,\\
      \infty & \text{if }\ell = 0.
    \end{cases}
  \]
  Note that $p = e$ when $q = 1$.
  For $h$ a natural number, let $\nu_p(h)$ be the $p$-adic valuation of $h$.
  By convention, set $\nu_\infty(h) = 0$ for all $h$.
  Define the function $\nu_{e,p} : \NN \rightarrow \cbr{-1} \cup \NN$ by
  \[
    \nu_{e,p}(h) := \begin{cases}
      \nu_p(h) & e \text{if }\mid h\\
      -1 & e \text{if }\nmid h
    \end{cases}.
  \]
  Lastly, let $h_{ab}^\lambda$ be the hook length of node $(a,b)$ in $\brk \lambda$ as defined in \cite{Kleshchev}.
  With this language, we may express the following theorem, parts (ii)-(iii) of which are known as the \emph{Carter criterion} in the symmetric group case, due to James--Mathas.\cite{Mathas-book}
  \begin{theorem}[James--Mathas]\label{Mathas Irreducibility}
    The following are equivalent:
    \begin{enumerate}[label={(\roman*)}]
      \item $S^{\lambda} \simeq D^{\lambda}$.
      \item $\lambda$ is $e$-regular and $S^\lambda$ is irreducible.
      \item $\nu_{e,p}\prn{h_{ab}^\lambda} = \nu_{e,p}\prn{h_{ac}^\lambda}$ for all nodes $(a,b)$ and $(a,c)$ in $\brk \lambda$.
    \qed\end{enumerate}
  \end{theorem}
  \begin{proof}
    See \cite{Mathas-book} theorem 5.42.
  \end{proof}
  This result gives information solely on $e$-regular partitions, and the general irreducibility of $S^\lambda$ away from $p=2$ is not well understood.
  We will henceforth specialize slightly to the case that $(n+r,n)$ is $e$-regular.

  \begin{corollary}\label{S irreducibility}
    If $r = 0$, assume $e > 2$.
    \begin{enumerate}[label={(\roman*)}]
      \item Suppose $e > n$.
        Then, $S^{(n+r,n)}$ is irreducible iff $e \nmid r + 2,r+3,\dots,n + r + 1$.
      \item Suppose $e \leq n$.
        If $S^{(n+r,n)}$ is irreducible, then $e \mid r + 1$.
    \end{enumerate}  
  \end{corollary}
  Note that the condition $e \nmid r+2,r+3,\dots,n+r+1$ implies that $e > n$.
  \begin{proof}
    Our initial assumption on $e$ implies that $\lambda$ is $e$-regular, which we will use below.

    \textbf{(i)}
    Note that $\nu_p(h) \neq -1$ for all naturals $h$ and only hook lengths in the top row may vanish mod $e$ by Figure \ref{Hooks};
    hence we may equivalently prove that $e$ divides no hook lengths in the leftmost $n$ columns of the second row by Theorem \ref{Mathas Irreducibility}.
    These hook lengths are precisely $r + 2,\dots,n+r+1$.

    \textbf{(ii)}
    Note that we have $\nu_{e,p}\prn{h^\lambda_{2,n-e+1}} \neq -1$, and $\nu_{e,p}\prn{h^\lambda_{2,n+r}} = -1$ so $\nu_{e,p}$ acquires at least two values.
    Suppose that $e \nmid r + 1$.
    Then, \[\nu_{e,p}\prn{h^\lambda_{1,n-e+1}} = \nu_{e,p}\prn{h^\lambda_{2,n-e+1} + r + 1} = -1,\]
    giving $S^{(n+r,n)}$ reducible by Theorem \ref{Mathas Irreducibility}.
  \end{proof}

  \begin{figure}
  \ytableausetup{mathmode,boxsize=4.5em}
  \begin{ytableau}
    n+r+1 & n + r & \dots & r + 2 & r & r-1 & \dots & 1\\ 
    n & n-1 & \dots & 1
  \end{ytableau}
  \ytableausetup{mathmode,boxsize=normal}
  \caption{The young diagram corresponding to the partition $(n+r,n)$. The hook lengths are in the center of the corresponding cells.}\label{Hooks}
\end{figure}

  From part (i) we see that irreducibility at $e > n$ is not dependent on $p$, and we may cover many modular cases without reference to the characteristic of $k$.
  We will finish our discussion of irreducibility of $S^\lambda$ via sharp characterization of the $e \leq n$ case.
  \begin{corollary}
    If $r = 0$, assume $e > 2$.
    Suppose $e \leq n$, and suppose $p > n + r + 1$.
    Then, $S^{(n+r,n)}$ is irreducible if and only if $e \mid r+1$.
  \end{corollary}
  \begin{proof}
    This follows from the proof of Corollary \ref{S irreducibility} part (ii) and the fact that $\nu_p(h) = \nu_p(h')$ for all naturalls $h,h'$. 
  \end{proof}
  \subsection{Branching theorems of Specht modules}
  In this section as well as later sections, we will consider the restriction of representations of $\SH$ to particular subalgebras isomorphic to $\SH_{k,q}\prn{S_{2n + r - 1}}$.
  We verify in Appendix \ref{Algebra} that any two subalgebras of $\SH$ generated by $2n + r - 2$ simple transpositions are canonically isomorphic, and the corresponding restrictions are canonically isomorphic via this isomorphism of algebras.
  We will hence abuse notation, pick one such subalgebra $\SH'$, and notate $\Res_{\SH'}^\SH W$ by $\Res W$ for any $\SH$-module $W$.

  Fixing some partition $\lambda \vdash 2n + r$, for $1 \leq i \leq j \leq l$, let $\beta_\lambda(i,j)$ and $\gamma_\lambda$ be the quantities
  \begin{align*}
    \beta_\lambda(i,j) &= \lambda_i - \lambda_j + \sum_{t = i}^j a_t\\
    \gamma_\lambda(i,j) &= \lambda_i - \lambda_j + \sum_{t = i+1}^j a_t.
  \end{align*}
  Note that $\beta_\lambda(i,j)$ is the hook length of cell $(a_1 + \dots + a_{i-1}+1,\lambda_j)$.

  Results due to Kleshchev and Brundan refer to \emph{normal} and \emph{good} numbers;
  for these, we will use the facts that 1 is always normal and that $j$ is normal when $\beta_\lambda(i,j) \not\equiv 0 \pmod e$ for all $i \leq j$.
  Further, we will use that $j$ is good if and only if $j$ is normal and $\gamma_\lambda(j,j') \not\equiv 0 \pmod e$ for all $j' \geq j$ normal.\cite{Brundan,Kleshchev}
  When $\lambda(i) = \mu$ for $i$ normal, write $\mu \xrightarrow{\text{normal}} \lambda$, and similar in the good case.

  The following statements, collectively known as \emph{modular branching rules} of $D^\lambda$, were originally written by Kleshchev for Specht modules of the group algebra $k\brk{S_n}$, then generalized to the Hecke algebra case by Brundan.\cite{Kleshchev,Brundan} 
  They entirely characterize the socle of $\Res D^\lambda$, as well as semisimplicity od $\Res D^\lambda$.
  \begin{theorem}[Kleshchev-Brundan]
    We have the following isomorphisms of vector spaces
    \begin{align*}
      \Homarg{\SH'}{S^\mu,\Res D^\lambda} 
      &\simeq \begin{cases}
        k & \text{if }\mu \xrightarrow{\text{normal}} \lambda,\\
        0 & \text{otherwise}.
      \end{cases}\\
      \Homarg{\SH'}{D^\mu,\Res D^\lambda} 
      &\simeq \begin{cases}
        k & \text{if } \mu \xrightarrow{\text{good}} \lambda\\
        0 & \text{otherwise}
      \end{cases}
    \end{align*}
    and $\Res D^\lambda$ is semisimple if and only if every normal number in $\lambda$ is good.\qed
  \end{theorem}

  Using this, we immediately see that, for any rectangular partition $(m^\ell)$, we have
  \[
    \Res D^{(m^\ell)} \simeq D^{(m^{\ell-1},m-1)}.
  \]
  The non-rectangular two-row case is more complicated, but we may still describe it fully as follows.
  \begin{corollary}\label{D Restrictions}
    Suppose $r > 0$.
    Then, we may characterize the socle of $\Res \, D^\lambda$ as follows:
    \[
      \soc\prn{\Res \, D^{(n+r,n)}} \simeq \begin{cases}
        D^{(n + r - 1,n)} & \text{if }e \mid r+2\\
        D^{(n + r,n-1)} &e \text{if }e\nmid r+2, \; e \mid r\\
        D^{(n+r-1,n)} \oplus D^{(n+r,n-1)} & \text{if } e \nmid r+2,r
      \end{cases}
    \]
    Further, when $e \nmid r$ or $e \mid r + 2$, $\Res D^{(n+r,n)}$ is semisimple.
  \end{corollary}
  \begin{proof}
    This amounts to computations of the hook lengths $\beta(1,2)$ and $\gamma(1,2)$:
    \begin{align*}
      \beta_\lambda(1,2) &= r + 2\\
      \gamma_\lambda(1,2) &= r
    \end{align*}
    Since $2$ is the largest removable number, $D^{(n+r,n-1)} \subset D^{(n+r,n)}$ if and only if $e \nmid r + 2$.
    Further, if $e \nmid r + 2$, then $D^{(n+r-1,n)} \subset D^{(n+r,n)}$ if and only if 1 is good;
    this is equivalent to $e \nmid r$.
  \end{proof}

  Now that we've characterized how $D^\lambda$ restrict, we can describe how strongly these restrictions characterize irreducibles.
  Namely, we will prove that some $D^\lambda$ having the same composition series as  $D^{(n+r,n)}$ is sufficient to determine that $\lambda = (n+r,n)$ in the case that condition (i) of Corollary \ref{S irreducibility} holds and either $r \neq 0$ or $e \neq 4$.
  \begin{proposition}\label{Combinatorics}
    Let $\lambda$ be an $e$-regular partition of $2n + r$.
    \begin{enumerate}[label={(\roman*)}]
      \item Suppose $r > 0$, suppose $e \nmid r+1, r + 2,\dots,n + r + 1$, and suppose either $e \mid r$ or $e \nmid r-2$.
        If $D^\lambda$ has the composition series 
        \begin{equation}\label{D Composition Series}
          0 \subset D^{(n+r-1,n)} \subset \Res \, D^{\lambda}
        \end{equation}
        with factor $\Res \, D^{\lambda} / D^{(n+r-1,n)} \simeq D^{(n+r,n)}$, then $\lambda = (n+r,n)$.
      \item Suppose $r = 0$, suppose $e \nmid 4$, and suppose $D^{(n,n-1)} \simeq \Res \, D^\lambda$.
        Then $\lambda = (n,n)$. 
    \end{enumerate}
  \end{proposition}
  \begin{proof}
    Note that $e > n$.
    Further, note that the above characterizations are necessary regardless of $e$-regularity;
    in the case that $\mu$ below fails too be $e$-regular, this proposition will prove that $\lambda$ does not satisfy \ref{D Composition Series}, a contradiction.

    \textbf{(i)}
    Let $\varpi := (n+r-1,n,1)$, let $\varsigma := (n+r-1,n+1)$, and let $\mu := (n + r,n)$.
    Since $D^{(n+r-1,n)} \subset \Res \, D^\lambda$, we have $(n + r - 1,n) \longrightarrow \lambda$, implying $\lambda \in \cbr{\varpi,\varsigma,\mu}$.
    We will show that $\varpi, \varsigma$ do not have socle compatible with \eqref{D Composition Series}, allowing us to conclude $\lambda = \mu$.

    If $\varpi$ or $\varsigma$ are not $e$-regular, then $D^\varpi = 0$ or $D^\varsigma = 0$, and we may immediately rule these out;
    henceforth assume that these are each $e$-regular.

    First suppose that that $\lambda = \varpi$.
    We will break into cases with $r$.
    \begin{itemize}
      \item Suppose that $r > 1$.
        Note that $e \nmid r + 1 = \beta_\varpi(1,2)$, so $2$ is normal.
        Further, $\gamma_\varpi(2,3) = n \not\equiv 0 \pmod e$, so 2 is good and $D^{(n+r-1,n-1,1)} \subset D^\varpi$, which is not a composition factor in \eqref{D Composition Series}.
        Hence, by the Jordan-H\"older theorem \cite{Etingof}, we have $\lambda \neq \varpi$.

      \item Suppose that $r = 1$.
        Then, $\varpi = (n,n,1)$ has $\gamma_{\varpi}(1,2) = n \not\equiv 0 \pmod e$, giving $D^{(n,n-1,1)} \subset D^\varpi$ and hence $\lambda \neq \varpi$ as in the previous case.
    \end{itemize} 
  \iffalse % keeping just in case
   \begin{align*}
    \beta_{\varpi}(1,2)     &= r + 1 \\
    \beta_{\varpi}(1,3)     &= n + r + 1 \\
    \beta_{\varpi}(2,3)     &= n + 1\\
    \gamma_{\varpi}(1,2)    &= r\\
    \gamma_{\varpi}(1,3)    &= n + r\\
    \gamma_{\varpi}(2,3)    &= n\\
  \end{align*}
  \fi

  \vspace{5pt}
  Now suppose that $\lambda = \varsigma$ and break into cases with $r$:
  \begin{itemize}
    \item Suppose $r > 2$.
      Then, by Corollary \ref{D Restrictions}, we require that $e \nmid r$ and $e \mid r - 2$;
      these are not satisfied, so $\lambda \neq \varsigma$.
    \item Suppose $r = 2$.
      Then $\Res \, D^\varsigma \simeq D^{(n+1,n)}$ is irreducible, as $\varsigma = (n+1,n+1)$ has rows of the same length;
      this contradicts \eqref{D Composition Series}. 
    \item Suppose $r < 2$. 
      Then $\varsigma$ is not a partition.
  \end{itemize}
  This completes the proof.

  \textbf{(ii)}
  Since the socle of $D^\lambda$ is irreducible, we require that 1 is the only normal number and $\lambda(1) = (n,n-1)$.
  This reduces to the cases of $\varsigma := (n + 1, n - 1)$ and $\mu := (n,n)$;
  if $\lambda = \varsigma$, then we have that $e \mid \beta_\varsigma(1,2) = 4$, a contradiction.
  Hence $\lambda = \mu$.
\end{proof}

\iffalse
\begin{proposition}
  Suppose that $n > 0$ and $e \nmid n + r$.
  \begin{enumerate}[label={(\roman*)}]
    \item Suppose $r > 0$, $n > 1$, and either $e \nmid n + 1$ or $e \nmid n + r + 1, n + r - 1$.
      If $D^\lambda$ has the composition series 
      \begin{equation}\label{D Composition Series}
        0 \subset D^{(n+r-1,n)} \subset \Res \, D^{\lambda}
      \end{equation}
      with factor $\Res \, D^{\lambda} / D^{(n+r-1,n)} \simeq D^{(n+r,n)}$, then $\lambda = (n+r,n)$.
    \item Suppose $n = 1$ and $e \nmid r+2$, and suppose that we have composition series \eqref{D Composition Series}.
      Then $\lambda=(r+1,r+2)$.
    \item Suppose $r = 0$, either $e \nmid n+2$ or $e > 3$, and $D^{(n,n-1)} \simeq \Res \, D^\lambda$.
      Then $\lambda = (n,n)$.
  \end{enumerate}
  The above conditions are unilaterally met whenever $e > n + r + 1$ and $(n,r) \neq (1,0)$.
\end{proposition}
\begin{proof}
  \textbf{(i)}
  Let $\lambda_1 := (n+r-1,n,1)$, let $\lambda_2 := (n+r-1,n+1)$, and let $\lambda_3 := (n + r,n)$.
  Since $D^{(n+r-1,n)} \subset \Res \, D^\lambda$, we have $(n + r - 1,n) \longrightarrow \lambda$, implying $\lambda = \lambda_i$ for some $i = 1,2,3$.
  We will show that $\lambda_1,\lambda_2$ each have good numbers other than 1, leading to elements of their socle that are not factors in \eqref{D Composition Series};
  then, we will have $\lambda = \lambda_3$.

  We may straightforwardly compute this.
  First assume $n \geq 2$, so that
  \begin{align*}
    \beta_{\lambda_1}(1,2) &= n + 1\\
    \beta_{\lambda_1}(1,3) &= n + r + 1\\
    \beta_{\lambda_1}(2,3) &= n + r - 1\\
    \beta_{\lambda_2}(1,2) &= n + r
  \end{align*}

  \textbf{(ii)}
  The choices are the same as above, but now $\lambda_1$ only has two removable nodes, so that
  \begin{align*}
    \beta_{\lambda_1}(1,3) &= r + 2\\
    \beta_{\lambda_2}(1,2) &= r + 1
  \end{align*}

  \textbf{(iii)}
  Let $\lambda_1 := (n,n-1,1)$, let $\lambda_2 := (n+1,n-1)$, and let $\lambda_3 := (n,n)$;
  as before we know that $\lambda = \lambda_i$ for $i = 1,2,3$, so we will eliminate $\lambda_1,\lambda_2$.
  First, we will show that $\lambda_1$ has a good number other than 1:
  \begin{align*}
    \beta_{\lambda_1}(1,2) &= n\\
    \beta_{\lambda_1}(1,3) &= n + 2\\
    \beta_{\lambda_1}(2,3) &= n\\
  \end{align*}
  Hence we may rule out $\lambda_1$ when $e | n$.

  Suppose $D^{\lambda_2}$ satisfies the isomorphism;
  then the number 2 is normal.
  We may make the following computations:
   \begin{align*}
    \beta_{\lambda_2}(1,2) &= n + 2\\
    \gamma_{\lambda_2}(1,2) &= 3.
  \end{align*}
  When $e | n + 2$, this brings contradiction;
  when $e \nmid n + 2$ and $e > 3$, 2 is a good number (recalling that $e > 2$), giving non-semisimple restriction, a contradiction.
  Hence $\lambda = \lambda_3$ and the proposition is proven.
\end{proof}
\fi

\section{Crossingless matchings and Specht modules}\label{Crossingless Matchings Section}
In this section, we will analyze the crossingless matchings representation $M := M_{2n + r}^r$ with the goal of proving $M_{2n + r}^r \simeq S^{(n+r,n)}$ under certain conditions in $e$.
We begin by proving irreducibility of $M_{2n + r}^r$ whenever $e \nmid r+2,r+3,\dots,n+r+1$;
when $e>n$, this is true if and only if $S^{(n+r,n)}$ is irreducible by Corollary \ref{S irreducibility}.
This is proven via an inductive process; if $e \nmid n + r + 1$, then $M$ contains no sign subrepresentation, and this allows us to ``project'' down to the case $M_{2(n-1) + r}^r$ and deduce irreducibility of $M$ from irreducibility of this representation.

We will use the following base case to the correspondence throughout:
\begin{lemma}\label{Base cases}
  Note that $S^{(1^n)}$ is the sign representation and $S^{(n)}$ the trivial representation.
  We have the following isomorphisms
  \begin{enumerate}[label={(\roman*)}]
    \item $M_2^0 \simeq S^{(2)}$,
    \item $M_r^r \simeq S^{(1^r)}$.
  \end{enumerate}
  Each of these are 1-dimensional, so they are irreducible.
\end{lemma}
\begin{proof}
  (i) follows from the fact that $(1 + T_1)w = (1 + q)w$, and hence $T_1w = qw$ for any nonzero vector $w \in M_2^0$.
  Similarly, (ii) follows from the fact that $(1 + T_i)w = 0$, so $T_iw = -w$ for any $i$ and any $w \in M_r^r$.
\end{proof}

\subsection{Irreducibility of $M$}
We refer to a vector $w \in M$ satisfying $\SH w = M$ as \emph{cyclic}.
It is a classical result that a representation $W$ is irreducible if and only if every nonzero element of $W$ is cyclic.\cite{Etingof}
We will prove irreducibility by showing that every nonzero $w \in M$ is cyclic;
the following lemma is integral in showing this.

\begin{lemma}
  \label{Cyclic}
  Every basis vector in $M_{2n + r}^r$ is cyclic.
\end{lemma}
\begin{proof}
  We have already proven this in the $r = 0$ case, so suppose that $r > 0$.

  Note that, between anchors at indices $a<a'$ having no arc at index $b$ with $a < b < a'$, the $M_{a'-a}^0$ case allows us to generate the basis vector with all length-2 arcs between $a,a'$ and identical arcs/anchors outside of this sub-matching.
  At the ends, we apply the $M_a^0$ case or the $M_{2n + r - a}^{0}$ case in the same way for the first $a$ or last $2n + r - a$ indices.

  Applying this between each arc gives us a vector with anchors and length-2 arcs, and we may use the appropriate $(1+T_i)$ to move anchors to any positions.
  Then, we may use the reverse process from above to generate the correct matchings between arcs and generate any other basis vector.
\end{proof}

Let $K_{2n+r}^r := \bigcap_{i=1}^{2n+r-1} \ker (1 + T_i) = \ker \bigoplus_{i = 1}^{2n + r - 1} (1 + T_i)$.
This will be an important technical tool in our proof of irreducibility.

We give two proofs of the following proposition. The first, below, intuits the necessary graphical intuition to construct a sub-matrix with nonzero determinant and perform row reduction. The second, given in section \ref{kernel}, formalizes the graphical intuition and gives results necessary for the following sections.

\begin{proposition}\label{Trivial}
  Suppose $e \nmid n + r + 1$.
  Then, $K := K_{2n + r}^r = 0$.
\end{proposition}
\begin{proof}[Proof 1.]
  Consider the matrix $A = \bigoplus (1 + T_i)$ having kernel $K$.
  It is sufficient by lemma \ref{rainbow nonzero} to show that $A$ includes a row $[0,\dots,0,1,0,\dots,0]$ with a nonzero entry only on the row $j$.

  Now, we may characterize the rows of $A$ as follows;
  if the row corresponding to $(1 + T_i)$ and mapping onto the element $w_l \in W$ is nonzero, then it is of the form $\brk{a_1,\dots,a_{\abs W}}$ where $a_l = 1 + q$, $a_m = q^{1/2}$ whenever $(1 + T_i)w_m = q^{1/2}w_l$, and $a_m = 0$ otherwise.

  Seeing this, the row corresponding to $(1 + T_{n + r})$ and $w_j$ has nonzero entries $q^{1/2}$ at $w_j$ and $(1 + q)$ at the vector $w$ agreeing with $w_j$ at all indices except having arcs at $(n+r-1,n+r)$ and $(n+r+1,n+r+2)$.
  Similar justification leads the row corresponding to $(1 + T_{n+r-1})$ at $w$ to have nonzero entries $q^{1/2}$ at $w$ and $(1 + q)$ at $w_j$ and the vector with anchors $1,\dots,r$ , arc $(n+r-3,n+r-2)$, and all other arcs maximum length.

  \begin{figure}
   \[
     \Action{10}{1/2,3/10,4/9,5/8,6/7}{2}{1/10,2/3,4/9,5/8,6/7}{q^{1/2}}
   \]
   \[
     \Action{10}{1/10,2/3,4/9,5/8,6/7}{2}{1/10,2/3,4/9,5/8,6/7}{q+1}
   \]
   \[
     \Action{10}{1/10,2/9,3/4,5/8,6/7}{2}{1/10,2/3,4/9,5/8,6/7}{q^{1/2}}
   \]
   \caption{Illustrated is the row constructed for transposition $(1 + T_2)$; clearly these are the only basis elements mapping to multiples of the desired element, and they relate to each other.
     replacing the outermost and/or innermost arc with an anchor typifies the rows constructed with three nonzero coefficients.}
 \label{Submatrix}
\end{figure}

  We may iterate this process as illustrated in Figure \ref{Submatrix}, eventually ending at a row with two nonzero entries, either an arc $(1,2)$ or an arc $(2,3)$, and all anchors otherwise left-aligned and arcs of maximum length.
  These rows together form an $(n + r) \times \abs{M_{2n + r}^r}$ submatrix of $A$ which has a nonzero row in the row corresponding to $j$, and has (by removing zero rows) the same row space as the following square matrix:
  \newcommand*\bigzero{\makebox(0,15){\text{\Huge0}}}
  \newcommand*\bigzerotwo{\makebox(-20,15){\text{\Huge0}}}
  \begin{equation}\label{Bnr}
    B_{n + r} := \begin{bmatrix}
      \begin{matrix}
      q + 1 & q^{1/2} & \\
      q^{1/2} & q + 1 & q^{1/2}\\
      & q^{1/2} & q + 1 & q^{1/2}\\
      &  & \ddots & \ddots
      \end{matrix}
      &   \bigzero\\
      \bigzerotwo & \begin{matrix}
        q^{1/2} & q+1 & q^{1/2}\\
        &  q^{1/2} & q + 1
       \end{matrix}
    \end{bmatrix}.
  \end{equation}
  We will show that this matrix is invertible;
  then, a sequence of elementary row operations will yield the identity, and in particular, when applied to $A$, will yield a row with a nonzero entry only on row $j$, giving $K = 0$.

  We may prove invertibility of this matrix by proving that $\det B_{n + r} = \brk{n + r + 1}_q$ inductively on $n+r$.
  This is satisfied for our base case $n + r = 1$, so suppose that it is true for each $m < n + r$.
  Then,
  \begin{align*}
    \det B_{n+r} &= (q + 1) \det B_{n + r - 1} - q \det B_{n + r - 2}\\
    &= (q + 1)(1 + \dots + q^{n + r - 1}) - (q + \dots + q^{n + r - 1})\\
    &= 1 + \dots + q^{n + r}\\
    &= \brk{n + r + 1}_q.
  \end{align*}
  Hence $\det B_{n + r} \neq 0$, and $K = 0$.
\end{proof}
\begin{theorem}\label{Irreducibility Theorem}
  Suppose that $e \nmid r+2,r+3,\dots,n+r+1$.
  Then the representation $M_{2n + r}^r$ is irreducible. 
\end{theorem}
\begin{proof}
  We proceed by induction on $n$.
  The base case the base case $n = 0, r \neq 0$ and $n = 1, r = 0$ follow from Lemma \ref{Base cases}.

  Take an arbitrary vector $w \in M$.
  By Proposition \ref{Trivial} there exists some $(1 + T_i) \in \SH$ such that $(1 + T_i)w \neq 0$.
  Note that \[\ima (1 + T_i) = \Span\cbr{w_j \mid w_j \text{ contains the arc }(i,i+1)}.\]
  Hence, as vector spaces, there is an isomorphism $\varphi:\ima(1 + T_i) \rightarrow M_{2(n-1) + r}^r$ ``deleting'' the arc $(i,i+1)$.

  We will show that, for every element $(1 + T_j') \in \SH(S_{2(n-1) + r})$, there is some element $h_j \in \SH(S_{2n + r})$ such that the following commutes:
  \[
    \begin{tikzcd}
      \ima (1 + T_i) \arrow[r,"\varphi" above, "\sim" below] \arrow[d,"h_j"] & M_{2(n-1) + r}^r \arrow[d,"1 + T_j'"]\\
      \ima (1 + T_i) \arrow[r,"\varphi" above, "\sim" below] & M_{2(n-1) + r}^r
    \end{tikzcd}
  \]
  Indeed, when $i+1 \neq j$ this is given by $h_j = 1 + T_j$, and we have $h_{i+1} = q^{-1}(1 + T_i)(1 + T_{i+1})(1 + T_{i-1})$, as given by Figure \ref{bigloop}.

  Note that, by definition, $e \nmid r+2,\dots,n+r$ as well, and hence $(n+r-1,n-1)$ satisfies the hypotheses of the proposition as well.
  Then, by the inductive hypothesis, there is some element $h' \in \SH(S_{2(n-1) + r})$ sending $\varphi((1 + T_i)w)$ to the image of a basis vector of $M_{2n + r}^r$ via $\varphi$;
  then, the action $\SH$ generates the endomorphism $\varphi^{-1}h'\varphi$ of $M$, which sends $(1 + T_i)w$ to a basis vector in $M_{2n + r}^r$.
  This implies that $w$ is cyclic, and hence $M_{2n + r}^r$ is irreducible.
  \begin{figure}
  \begin{adjustbox}{width=\textwidth}
    \Action{6}{1/6, 2/5, 3/4}{2}{1/6, 2/3, 4/5 }{q^{1/2}} \hspace{20pt} 
    \IsoAction{8}{1/8, 2/7, 3/4, 5/6}{3}{1/8, 2/5, 3/4, 6/7}{q^{1/2}}
  \end{adjustbox}
  \begin{adjustbox}{width=\textwidth}
    \Action{4}{1/4, 2/3}{2}{1/4,2/3}{(1 + q)} \hspace{20pt}
    \IsoAction{6}{1/6, 2/5, 3/4}{3}{1/6, 2/5, 3/4}{(1 + q)}
\end{adjustbox}
  \caption{The correspondence between the action of $(1 + T_2)$ on $w'_5 \in M^0_6$ and the action of $q^{-1}(1 + T_3)(1 + T_4)(1 + T_2)$ on the corresponding vector in $M^0_8$ having arc $(3,4)$ first, then on $w'_2 \in M^0_4$.
  This demonstrates that the action works with or without creating a loop.
  }
  \label{bigloop}
  \end{figure}
\end{proof}

\begin{figure}
\[
  \begin{gathered}\CapNAction{4}{1/1,3/2,2/3}{2}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\BigNAction\end{gathered}
\]
  \caption{The isotopy demonstrating the correspondence between $h_{i+1}$ and $T_i'$;
  the scaling issue with having three simple reflections is handled by the $q^{-1}$ factor in $h_{i+1}$.}
\end{figure}

\subsection{Correspondence with Specht modules}
The following theorem due to Mathas (theorem 5.5 in \cite{Mathas-article}) generalizes the classical branching theorem of the symmetric group.
It will not be necessary for our present proof of the correspondence, but analogy with $M$ is suggestive.
\begin{theorem}[Characteristic-free classical branching theorem]
  Let $\lambda$ be a partition of $m$ with $\ell$ removable nodes.
  Then, $\Res S^\lambda$ has an $\SH_{k,q}(S_{m-1})$-module filtration
  \[
    0 = S^{0,\lambda} \subset S^{1,\lambda} \subset \cdots \subset S^{\ell,\lambda} = \Res S^\lambda
  \]
  such that $S^{t,\lambda} / S^{t-1,\lambda} \simeq S^{\lambda(t)}$ for all $1 \leq t \leq \ell$.\qed
\end{theorem}
In particular, this holds in cases where $S^\lambda$ fails to be irreducible.
If we replace $S^\lambda$ with the appropriate $M_{2n + r}^r$ above, we find the statement of the following proposition.
\begin{proposition}Suppose that $n > 0$.

  \begin{enumerate}[label={(\roman*)}]
    \item
    Suppose that $r > 0$.
    Then, a filtration of $\Res M_{2n + r}^r$ is given by
    \begin{equation}
      0 \subset M_{2n + r - 1}^{r-1} \subset \Res M_{2n + r}^r, \label{Filtration}
     \end{equation}
     with $\Res M_{2n + r}^r / M_{2n + r - 1}^{r-1} \simeq M_{2n + r - 1}^{r + 1}$. 
    \item
      We have the following isomorphism of representations:
      \begin{equation}
        M_{2n - 1}^1 \simeq \Res M_{2n}^0. \label{0 Restriction}
       \end{equation}
  \end{enumerate}
  In the case that $e \nmid r+1,\dots,n + r + 1$, \ref{Filtration} and \ref{0 Restriction} are composition series.
\end{proposition}
\begin{proof}
  \textbf{(i)}
  Note that we may identify the subrepresentation of $\Res M_{2n + r}^r$ having anchor at index $2n+r$ with $M_{2n + r - 1}^{r-1}$.

  Let $U := \Res M_{2n + r}^r / M_{2n + r - 1}^{r - 1}$, and let $\pi: \Res M_{2n + r}^r \twoheadrightarrow U$ be the associated projection to $U$.
  Let $\phi:U \rightarrow M_{2n + r - 1}^{r + 1}$ be the $k$-linear map which regards the arc $(i,2n + r)$ in $U$ as an anchor at $i$ in $M_{2n + r - 1}^{r + 1}$.
  It is not hard to verify that this is a well-defined isomorphism of vector spaces, so we must show that it is $\SH$-linear.

  Given a basis vector $w_j$ containing arc $(i,2n + r)$, $\phi$ is clearly compatible with $(1 + T_{i'})$ with $i' \neq i,i-1$.
  Further, it is easy to verify that $\phi$ is compatible with $T_i$ by definition of the relevant modules, as well as with $(1 + T_{i-1})$ when $w_j$ does not have anchor $i-1$.
  The case that $w_j$ has anchor $i-1$ is illustrated in the following commutative diagram:
  \[
    \begin{tikzcd}
      M_{2n + r}^r \arrow[r,"1 + T_i"] \arrow[d,"\phi \circ \pi"] & M_{2n + t}^r \arrow[r,"\pi"] & U \arrow[d,"\phi"]\\
      M_{2(n-1) + r + 1}^{r+1} \arrow[rr,"1 + T_i"] & & M_{2(n-1) + r + 1}^{r+1}
    \end{tikzcd}
  \]
  which we may diagram chase as follows
    \[
    \QuotDiagram
  \]

  In each of these cases, linearity of $\SH$-action and $\phi$ give 
  \begin{align*}
    \phi(T_jw_i) 
    &= \phi(-w_j + (1+T_j)w)\\
    &= -\phi(w_j) + (1 + T_j)\phi(w_j)\\
    &= T_j\phi(w_j).
  \end{align*}
  Hence $\phi$ is an isomorphism of representations, and the statement is proven.

  \textbf{(ii)}
  This follows from an analogous proof:
  now, $\phi:\Res M_{2n}^0 \rightarrow M_{2(n-1)+1}^1$ is an isomorphism of representaitons, which is proven to be $\SH$-linear by the same proof.
\end{proof}

We've now assembled the basic pieces necessary to prove our correspondence in the case $e > n + r + 1$.

\begin{theorem}\label{Correspondence Theorem}
  Suppose $e > n + r + 1$.
  Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.
\end{theorem}
\begin{proof}
  The case $n = 0$ is already proven, by lemma \ref{Base cases}, so suppose $n > 0$.
  In order to use Proposition \ref{Combinatorics}, suppose for now that either $e \nmid 4$ or $r > 0$.

  We will prove this inductively; the base case $2n + r = 2$ is implied by \ref{Base cases}, so suppose that $M_{2n + s}^s \simeq S^{(m+s,s)}$ whenever $2m + s < 2n + r$ and $m + s \leq n + r$, so that $e > m + s + 1$.

  Suppose $r > 0$.
  By Theorem \ref{Irreducibility Theorem}, we know that $M_{2n + r}^r \simeq D^\lambda$ for some $e$-restricted partition $\lambda$.
  By the inductive hypothesis and Corollary \ref{S irreducibility}, we have a composition series given by the short exact sequence
  \begin{equation}
    \label{Composition Series} 0 \longrightarrow D^{(n + r - 1,n)} \longrightarrow \Res \, D^\lambda \longrightarrow D^{(n+r,n-1)} \longrightarrow 0
   \end{equation}
  Hence the theorem is given by Proposition \ref{Combinatorics} (i).

  Now suppose $r = 0$ and $e \neq 4$.
  Similarly, by Theorem \ref{Irreducibility Theorem}, we know that $M_{2n + r}^r \simeq D^\lambda$ for some $e$-restricted partition $\lambda$, and by the inductive hypothesis and Corollary \ref{S irreducibility}, we have the irreducible restriction $\Res D^\lambda \simeq D^{(n,n-1)}$.
  Then, the theorem is given by Proposition \ref{Combinatorics} (ii).

  Now, suppose $e = 4$ and $r = 0$;
  then $4 > n + 1$, so $n \leq 2$.
  We've already proven the $n = 1$ case via the trivial representation, so suppose $n = 2$.
  Then, from the proof of Proposition \ref{Combinatorics}, we know that $M_{4}^0 \simeq D^\lambda$, where $\lambda \in \cbr{(n,n), (n+1,n-1)}$.
  We have already proven that $M_4^2 \simeq D^{(n+1,n-1)}$, and we may verify that $\dim M_4^0 = 2 \neq 3 = \dim M_4^2$, so we have that $\lambda = (n,n)$ and the theorem is proven for $e = 4$.
\end{proof}

\subsection{Kernels and further work}

In this section, we characterize $K_{2n+r}^r := \bigcap_{i=1}^{2n+r-1} \ker (1 + T_i) = \ker \bigoplus_{i = 1}^{2n + r - 1} (1 + T_i)$ for $e \nmid n + r + 1$ and for $e=n+r+1$. When the representation is clear, we will simply denote this kernel as $K$ or $\cap\text{ker}(1+T_i)$. For compactness, in this section we use $\sim $ to denote "proportional to". For convenience, define $M_0^0$ to be the zero representation.

\subsubsection{Proposition \ref{Trivial}}
\label{kernel}
\begin{definition}

	Fix some basis element $\psi\in M_{2n+r}^r$. For $1 \leq a,b \leq 2n + r$ define $\psi(a):=b$ iff $a$ and $b$ are matched in $\psi$, $\psi(a):=a$ if $a$ is an anchor in $\psi$. Given that $\psi$ has $r'$ anchors in the range $a,..,b$, define a \textbf{sub-matching} $\psi(a,b)$ of $\psi$ to be the basis element $\sigma\in M_{b-a+1}^{r'}$ specified by $\sigma(i)=\psi(i+a-1)-a+1$. This sub-matching is defined for $a<b$ when $\psi(i)\in \{a,a+1,...,b\}$ for all $i\in \{a,a+1,...,b\}$. See \ref{sub-matching example}.

	If $b\geq a$ and nodes $1\leq a,...,b\leq 2n+r$, $\psi(a,b)$ will always refer to those nodes and they're matchings, though it may not be a sub-matching unless specified. For any other $a,b$, we define $\psi(a,b)$ to be an element of the representation with no nodes.

	\label{sub-matchings}
\end{definition}

\begin{figure} 
	\def\cbasisspacing{5mm}
	$\cbr{
		\begin{gathered}
		\Matching{6}{1/2, 3/6, 4/5}, \hspace{\cbasisspacing}
		\Matching{4}{1/4, 2/3}
		\end{gathered}}$
	\caption{$\psi\in M_6^0$ is pictured on the left, $\sigma\in M_4^0$ is pictured on the right. $\psi(3,6)=\sigma$. $\psi(2,5)$ is not a sub-matching.}
	\label{sub-matching example}
\end{figure}

Define the rainbow element $R\in M_{2n+r}^r$ to be the basis element specified by $R(i)=2n+2r-i+1$ for $i>r$, $R(i)=i$ for $i\leq r$. In other words, the basis element with all anchors to the left then a rainbow.

\begin{proposition}
	Let $w$ be an arbitrary vector in $M_{2n+r}^r$. I claim that if $w\in \cap\text{ker}(1+T_i)$, the coordinate $c$ of the rainbow element $R$ in $w$ is nonzero.

	\label{rainbow nonzero}
\end{proposition}

\begin{proof}
	Let $Y$ be the set of basis elements with nonzero coordinate in $w$. Let $z$ be the greatest integer such that there exists $y\in Y$ where $y(1)-1=...=y(z)-z=0$, and let $U\subset Y$ be the set of such $y$. In other words, $U$ is the set of basis elements in $Y$ which have the most anchors to the far left.

	Suppose $z<r$. Then for each $y\in U$ there exists a minimal $i_y>z+2$ such that $y(i_y)=0$. In other words, $i_y$ is the position of the next leftmost anchor in $y$.  Fix $\tilde{y}$ such that $i_{\tilde{y}}\leq i_y$ for all $y$. Then I claim the basis element $y':=q^{-1/2}(1+T_{i_{\tilde{y}}-1})\tilde{y}$ has nonzero coordinate in $(1+T_{i_{\tilde{y}}-1})w$, implying $w\not\in \cap\text{ker}(1+T_i)$. To see this, we can show that $\tilde{y}$ is the only basis element in $Y$ such that $(1+T_{i_{\tilde{y}}-1})\tilde{y}\sim y'$. $y'$ still has $z$ anchors on the left, and $i_{y'}<i_{\tilde{y}}$, so $y'\not\in Y$. If $x\in Y,\not\in U$, the basis element $x'$ proportional to $(1+T_{i_{\tilde{y}}-1})x$ will have $k$ anchors at the far left only if the next anchor is at a position $i_{x'}>i_{\tilde{y}}$, so it cannot be $y'$. If $x\in U$, $x'$ will have anchor at $i_{y'}$ if and only if $i_x=i_{\tilde{y}}$ and $x(i_{\tilde{y}})=\tilde{y}(i_{\tilde{y}})$. Since this is the only match altered by action of $(1+T_{i_{\tilde{y}}-1})$ on $x$, if $(1+T_{i_{\tilde{y}}-1})x\sim y'$ this implies $x=\tilde{y}$. So if $z<r$ $w$ is not in the desired kernel.

	Suppose $z=r$ but $R\not\in U$ (so $R\not\in Y$). Let us define a sequence of subsets of $U$ in the following way: $U_0:=U$, $U_{i+1}:=\{u\in U_i| u(r+i+1)=2n+2r-i+1 \}$. Since $R\not\in U$, $\exists t<n-1$ such that $U_{t+1}=\emptyset$. Choose $\tilde{u}\in U_t$ such that $\tilde{u}(r+t+1)\geq u(r+t+1)$ for all $u\in U_t$. Consider the basis element $u':=q^{-1/2}(1+T_{\tilde{u}(r+t+1)})\tilde{u}$. I claim that $\tilde{u}$ is the only element in $Y$ such that $(1+T_{\tilde{u}(r+t+1)})\tilde{u}\sim u'$, again implying that $w$ is not in the desired kernel. $u'$ still has $r$ anchors on the left, $u'(r+i)=2n+2r-i+2$, $1\leq i\leq t$, and $u'(r+t+1)>\tilde{u}(r+t+1)$, so $u'\not\in Y$. If $x\in Y,\not\in U$,  the basis element $x'$ proportional to $(1+T_{\tilde{u}(r+t+1)})x$ will have $r$ leftmost anchors only if $x'(r+t+1)<\tilde{u}(r+t+1)$, so $x'\not=u'$. Similarly, if $x\in U,\not\in U_t$, the basis element $x'$ will have the property $x'(r+i)=2n+2r-i+2$ for all $i\leq t$ only if $x'(r+t+1)<\tilde{u}(r+t+1)$, so $x'\not=u'$. If $x\in U_t$, $x'(r+t+1)=u'(r+t+1)$ if and only if $x(r+t+1)=\tilde{u}(r+t+1)$ and $x(x(r+t+1)+1)=\tilde{u}(\tilde{u}(r+t+1)+1)$ (since $u'\not \in Y$). These are the only matches altered by the action $(1+T_{\tilde{u}(r+t+1)})$, so this implies $x=\tilde{u}$. Thus we have proved that if $R\not\in Y$, $w$ is not in the kernel.

\end{proof}

Given a rainbow element $R$, define the basis elements $R_{R,i},R_{L,i}$ to be those where you move the middle hump across $i$ lines to the right or left, respectively. Examples are pictured in \ref{shifted rainbow}. Formally, $R_{R,i}:=q^{-i/2}(1+T_{r+n+i})...(1+T_{r+n+1})R$, $R_{L,i}:=q^{-i/2}(1+T_{r+n-i})...(1+T_{r+n-1})R$.

\begin{figure}
	\def\cbasisspacing{5mm}
	$\cbr{
		\begin{gathered}
		\Matching{8}{1/8, 2/7, 3/6, 4/5}, \hspace{\cbasisspacing}
		\Matching{8}{1/8, 2/7, 3/4, 5/6}, \hspace{\cbasisspacing}
		\Matching{8}{1/8, 2/3, 4/7, 5/6}, \hspace{\cbasisspacing}
		\Matching{8}{1/2,3/8,4/7,5/6}
		\end{gathered}}$
	\caption{$R_{L,0},...,R_{L,3}$ pictured from left to right}
	\label{shifted rainbow}
\end{figure}

Define $Q_n:=(q^n+...+1)/q^{n/2}(-1)^n$ for $n\in \{0,1,...\}$. The following proposition says that, for any element in the kernel, if some basis element $y$ has coordinate $c$ in that element, and if $y$ has a rainbow sub-matching, the basis elements where you replace that sub-matching by the shifted rainbow matchings $R_{L,i}$ or $R_{R,i}$ both have coordinate $Q_ic$ in the kernel element.

\begin{proposition}
	Let $w$ be an element in the kernel intersection $\cap (1+T_z)$ in some generalized crossingless matchings representation. Let $y$ be a basis element with coordinate $c$ in $w$. Suppose $\exists a,b$ such that $y(a,b)=R$, the rainbow element. Define the basis elements $\theta_i,\phi_i$ by $\theta_i(1,a-1)=\phi(1,a-1)=y(1,a-1)$, $\theta_i(b+1,2n)=\phi(b+1,2n)=y(b+1,2n)$, $\theta_i(a,b)=R_{R,i}$, $\phi_i(a,b)=R_{L,i}$ (leave $\theta_i$ or $\phi_i$ undefined for any $i$ where $R_{R,i},R_{L,i}$ are undefined, respectively). The coordinates of $\phi_i$ and $\theta_i$ in $w$ are both $Q_ic$.

	\label{shifted rainbow coeffs}
\end{proposition}

Proof of this proposition requires a simple algebraic fact that will be used throughout this document, so I state it as a lemma.

\begin{lemma}
	$Q_1Q_n-Q_{n-1}=Q_{n+1}$

	\label{Q alg}
\end{lemma}

\textit{Proof of lemma.}

\begin{align*}
Q_1Q_n-Q_{n-1}=&\frac{-(q+1)}{q^{1/2}}\frac{(-1)^n(q^n+...+1)}{q^{n/2}}-\frac{(-1)^{n-1}(q^{n-1}+...+1)}{q^{(n-1)/2}}\\
=&\frac{(-1)^{n+1}(q^{n+1}+2q^n+...+2q+1)}{q^{(n+1)/2}}-\frac{(-1)^{n+1}(q^n+...+q)}{q^{(n+1)/2}}\\
=&\frac{(-1)^{n+1}(q^{n+1}+...+1)}{q^{(n+1)/2}}=Q_{n+1}
\end{align*}

Now let us prove the proposition.

\begin{proof}
	Consider acting on $w$ by an element $(1+T_z)$. The coordinate of $\phi_i$ in $(1+T_z)w$ will be a linear combination of the coordinates of basis elements sent to $\phi_i$ by the element $(1+T_z)$. Specifically, it will be $(1+q)c\iota+(q^1/2)\sum c_\psi$ where $\iota=1$ if $y(z)=z+1$, $\iota=0$ otherwise, and $c_\psi$ are the coordinates of all basis elements $\psi$ where $(1+T_z)\psi\sim y$.

	Let $n:=a+b-1$ and $r$ be the number of anchors in $y(a,b)$. Consider the coordinate of $\phi_i$ in $(1+T_{a-1+r+n/2-i})w$. This is the transposition that acts on the "moved middle hump" in $\phi_i(a,b)=R_{L,i}$, as shown in \ref{shifted relations}. I claim the following:

	\begin{center}
		\textit{claim}:
		The only basis elements $\psi$ where $(1+T_{a-1+r+n/2-i})\psi\sim \phi_i$ are $\phi_i$ and $\phi_{i-1},\phi_{i+1}$ when they exist\footnote{we defined $R_{L,i}$ as far out as we can move the hump, so for $0\leq i<n+r$, and take the analogous domain for $\phi_i$}.

	\end{center}

	Note that the action of any $(1+T_z)$ on a basis element $\psi$ creates exactly two lines: an arc of length two connecting $z$ and $z+1$, and either an anchor or an arc of length $\geq 2$ connecting $\psi(z)$ and $\psi(z+1)$.\footnote{Although this fact follows directly from our definition of the representation, it will be used throughout the document, so it is important that the reader understands it.\label{two humps created}} The easiest way to see the claim is to see that the given transposition is surrounded by arcs on both sides, so any basis element sent to the same element can vary from $\phi_i$ by at most one of those arcs and nothing else.

	Let us prove the claim formally: It is easy to see that the action of $(1+T_{a-1+r+n/2-i})$ will bring $\phi_{i-1},\phi_i,\phi_{i+1}$ to $\sim \phi$, as shown in \ref{shifted relations}. Suppose there was another basis element $\psi$ sent to $\phi_i$ by the given transposition. Me note that if $\psi$ contains the arcs or anchors directly to the right and left of the arc $(a-1+r+n/2-i,a-1+r+n/2-i+1)$ in $\phi_i$ (formally, it contains the arc $(a-1+r+n/2-i-1,a-1+r+n/2+i+2)$ or an anchor at $a-1+r+n/2-i-1$ and the arc $(a-1+r+n/2-i+2,a-1+r+n/2+i+1)$ or an anchor at $a-1+r+n/2-i+2$), it must contain the arc $(a-1+r+n/2-i,a-1+r+n/2-i+1)$ to be a crossingless matching. Thus, if $\psi$ contains both of these arcs/anchors, $(1+T_{a-1+r+n/2-i})$ acts as the constant $(1+q)$, so $(1+T_{a-1+r+n/2-i})\psi\sim \phi_i=>\psi\sim \phi$. If $\psi$ does not contain the left arc/anchor and $(1+T_{a-1+r+n/2-i})\psi\sim \phi_i$, the action of $(1+T_{a-1+r+n/2-i})$ must create that arc/anchor, so $\psi(a-1+r+n/2-i-1)=a-1+r+n/2-i$ and $\psi(a-1+r+n/2-i+1)=a-1+r+n/2+i+2$ in the case of an arc or $a-1+r+n/2-i+1$ is an anchor. All other matchings remain unchanged, so this implies $\psi=\phi_{i+1}$. Likewise, if the right arc $((a+b-1)/2-i+2,(a+b-1)/2+i+1)$ does not exist, $\psi=\phi_{i-1}$. For boundary cases, note that for $\phi_0=\theta_0$, the only other basis element sent to this by the middle transposition is $\phi_1=\theta_1$. Also note that at the edge case $\phi_{n+r-1}$ there is not necessarily a left arc, so other elements may be sent to $\phi_{n+r-1}$ by the given transposition, and this case gives no new information. Lastly, note that our argument was completely symmetric and thus applies to the $\theta_i$ case, except that for $\theta_i$ we do not have to deal with anchors. Thus the claim is proved.

	Given this claim and lemma \ref{Q alg}, the proposition follows quickly through induction:

	Acting by $(1+T_{a-1+r+n/2})$ on $w$, the new coordinate of $\phi_0=y$ is $(q+1)c+q^{1/2}c_{\phi_1}$ where $c_{\phi_1}$ is the coordinate of $\phi_1$ in $w$. Since $w$ is in the kernel, we have $(q+1)c+q^{1/2}c_{\phi_1}=0=>c_{\phi_1}=Q_1c$. $\phi_1=\theta_1$ so this gives us all our base cases.

	Acting by $(1+T_{a-1+r+n/2-i})$ on $w$, the new coordinate of $\phi_i$ is $q^{1/2}c_{\phi_{i+1}}+q^{1/2}c_{\phi_{i-1}}+(q+1)c_{\phi_{i}}=0$. By the inductive hypothesis, $q^{1/2}c_{\phi_{i+1}}+q^{1/2}Q_{i-1}c+(q+1)Q_i=0$ so $c_{\phi_{i+1}}=Q_1Q_i-Q_{i-1}=Q_{i+1}$ by lemma \ref{Q alg}. $\theta_i$ is an identical proof, so the proposition follows.

	\begin{figure}
		\[
		\GeneralizedAction{8}{3/4,5/8,6/7}{2}{1/1,2/2}{1}{3}{3/4,5/8,6/7}{1/1,2/2}{(1+q)}
		\]
		\[
		\GeneralizedAction{8}{2/3,5/8,6/7}{2}{1/1,4/2}{1}{3}{3/4,5/8,6/7}{1/1,2/2}{q^{1/2}}
		\]
		\[
		\GeneralizedAction{8}{3/8,4/5,6/7}{2}{1/1,2/2}{1.25}{3}{3/4,5/8,6/7}{1/1,2/2}{q^{1/2}}
		\]
		\caption{The action of $(1+T_{a-1+r+n/2-i})$ on $\phi_i,\phi_{i=1},\phi_{i+1}$ (ordered from top to bottom), shown as the case where $y$ is the rainbow vector in $M_{8}^2$ and $i=2$.}
		\label{shifted relations}
	\end{figure}
\end{proof}

We are now ready to prove proposition \ref{Trivial}.

\begin{proof}
	Suppose $\cap\text{ker}(1+T_i)=K\not=0$. Take nonzero $w\in K$. By Proposition \ref{rainbow nonzero}, the coordinate of the rainbow vector $R$ is nonzero; suppose the coordinate is $c$. By proposition \ref{shifted rainbow coeffs}, the coordinates of the basis elements $R_{L,n+r-1}$ and $R_{L,n+r-2}$ are $Q_{n+r-1}c$ and $Q_{n+r-2}c$ respectively.

	Consider the coordinate of $R_{L,n+r-1}$ in $(1+T_1)w$. Using the same logic as in the proof of proposition \ref{shifted rainbow coeffs}, we note that if a basis element $\psi$ has no anchor at position $3$ and is not equal to $R_{L,n+r-2}$, $(1+T_1)\psi\not\sim R_{L,n+r-1}$. Thus the desired coordinate is equal to $(1+q)Q_{n+r-1}c+q^{1/2}Q_{n+r-2}c=-q^{1/2}Q_{n+r}c$ by lemma \ref{Q alg}. Since $w\in K$, we must have $-q^{1/2}Q_{n+r}c=0$. Me have that $c$ is nonzero, and we assume $q$ nonzero, and $Q_{n+r}$ is zero iff $q$ is a root of $q^{n+r}+...+1$, implying $e|n+r+1$. Thus we have arrived at contradiction, and $K=0$.
\end{proof}

\subsubsection{Kernel Basis}

In this section, we determine an explicit basis for $K$ when $e=n+r+1$, assuming $K\not=0$. In the next section, we prove $K\not=0$.

First let us formalize a useful property of sub-matchings.

\begin{definition}
	Given a basis element $\psi\in M_{2n+r}^r$, specify some sub-matching $\psi(a,b)$. Let $\Res_{\SH_{b-a+1}(q)}^{\SH_{2n+r}(q)}M_{2n+r}^r$ be the restriction to the sub-algebra generated by transpositions $T_a,...,T_{b-1}$. Define $Y_\psi\subset\Res_{\SH_{b-a+1}(q)}^{\SH_{2n+r}(q)}M_{2n+r}^r$ to be the subrepresentation generated by the set of basis elements $\{\sigma| \sigma(1,a-1)=\psi(1,a-1),\ \sigma(b+1,2n+r)=\psi(b+1,2n+r)\}$.
\end{definition}

\begin{lemma}
	Take a basis element $\psi\in M_{2n+r}^r$. Suppose $\psi$ has some sub-matching $\psi(a,b)$ with $r'$ anchors. Define $Y_\psi$ with respect to this sub-matching.

	The map $\rho:Y_\psi\rightarrow M_{b-a+1}^{r'}$ defined by $$\rho(\sigma)=\sigma(a,b)$$ is an isomorphism of representations.

	\label{sub-matching isomorphism}
\end{lemma}

\begin{proof}

	The map is clearly bijective. Thus it is sufficient to prove the following: $$\rho(T_{i+a-1}\sigma)=T_i\rho(\sigma)$$

	As mentioned in the previous section, the action of a transposition $T_i$ can change at most 4 nodes, so we need to show that the transpositions end up changing the same nodes in the same way in  $\rho(T_{i+a-1}\sigma)$ and $T_i\rho(\sigma)$.

	Suppose $\sigma(i+a-1)=s,\sigma(i+a)=t$. Then $(T_{i+a-1}\sigma)(i+a-1)=i+a$, $(T_{i+a-1}\sigma)(s)=t$, so $\rho(T_{i+a-1}\sigma)(i)=i+1$, $\rho(T_{i+a-1}\sigma)(s-a+1)=t-a+1$. Separately, $\rho(\sigma)(i)=s-a+1$ and $\rho(\sigma)(i+1)=t-a+1$, so $T_i\rho(\sigma)(i)=i+1$ and $T_i\rho(\sigma)(s-a+1)=t-a+1$ as desired. So the map is an isomorphism and the lemma is proved.
\end{proof}

\vspace{5mm}
The lemma above motivates a recursive characterization of the kernel. To do this, it will be convenient to define some notation.

\begin{definition}
	Recall $Q_i:=(q^i+...+q+1)/q^{i/2}(-1)^i$ (lemma \ref{Q alg}). For $a>0$ define $Q(0,b):=1$. For $b>a>0$ define $$\Q_b^a:=\frac{Q_{b-1}...Q_{b-a}}{Q_1...Q_{a-1}}$$

\end{definition}

\begin{definition}
	For $\psi \in M_0^0$, define the function $x_\psi(q):=1$.

	For all other basis elements $\psi\in M_{2n+r}^r$, we define $x_\psi$ recursively:

	$$x_\psi(q):=x_{\psi(2,a-1)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{\lfloor a/2\rfloor}_{n+r}$$

	I will refer to $x_\psi$ as the \textbf{coordinate function} of $\psi$.

	\label{coeff def}
\end{definition}

\vspace{2mm}
The following proposition states the forward direction of our characterization.
\begin{proposition}

	Let $M_{2n+r}^r$ be a crossingless matchings representation, and suppose $Q_1,...Q_{n+r-1},\not=0$. Let $w\in\cap\text{ker}(1+T_i)$. MLOG the rainbow element $R$ has coordinate 1 in $w$ (by proposition \ref{rainbow nonzero}). Then the coordinate of any basis element $\psi\in M_{2n+r}^r$ in $w$ is $x_{\psi}(q)$.

	\label{kernel characterization}

\end{proposition}

An illustration of this proposition is shown in figure \ref{coeff example}.

\begin{figure}
	\def\cbasisspacing{5mm}

	$\cbr{
		\begin{gathered}
		\GeneralizedMatching{11}{1/6, 2/3,4/5,7/10,8/9}{1}{11/1}{11/8}, \hspace{\cbasisspacing}
		\Matching{4}{1/2, 3/4}, 
		\hspace{\cbasisspacing}
		\GeneralizedMatching{5}{1/4, 2/3}{1}{5/1}{3/4}, \hspace{\cbasisspacing}
		\end{gathered}}$ 
	\caption{Suppose the second and third elements have coordinates $x_2(q_1)$ and $x_3(q_2)$ in their respective kernel elements, where for $q_1$, $e=3$ and for $q_2$ $e=4$. The coordinate of the first element is $x(q)=x_2(q)x_3(q)\frac{Q_5Q_4Q_3}{Q_1Q_2}$, where for $q$, $e=7$}
	\label{coeff example}
\end{figure}

\begin{proof}
	Suppose $\psi(1)=a$. The proof is structured as follows: use proposition \ref{shifted rainbow coeffs} to find the coefficient of the basis element with $\lfloor a/2\rfloor$ humps then a rainbow element; use the same proposition in a reversed manner to find the coefficient of the basis element consisting of the rainbow for the first $a$ nodes, then the rainbow for the final $2n+r-a$ nodes; finally, we finish the proof through induction using lemma \ref{sub-matching isomorphism}.

	\vspace{2mm}
	By proposition \ref{shifted rainbow coeffs} the element $R_1:=R_{L,n+r-1}$ has coordinate $Q_{n+r-1}$ in $w$. Then $R_1(3,2n+r)$ is the rainbow element in  $M_{2(n-1)+r}^r$, so the element $R_2$ defined by $R_2(1,2):=R_1(1,2)$, $R_2(3,2n+r):=R_{L,n+r-2}\in M_{2(n-1)+r}^r$ has coordinate $Q_{n-1}Q_{n-2}$. Generally, define $R_i$ by $R_i(1,2(i-1)):=R_{i-1}(1,2(i-1))$, $R_i(2i-1,2n+r):=R_{L,n+r-i}\in M_{2(n-i+1)+r}^r$. Then the coefficient of $R_i$ is $Q_{n+r-1}...Q_{n+r-i}$. These elements are shown in figure 5.

	\begin{figure}
		\def\cbasisspacing{5mm}

		$\cbr{
			\begin{gathered}
			\GeneralizedMatching{8}{3/8, 4/7,5/6}{2}{1/1,2/2}{5/4}, \hspace{\cbasisspacing}
			\GeneralizedMatching{8}{1/2, 5/8,6/7}{2}{3/1,4/2}{5/4}, 
			\hspace{\cbasisspacing}
			\GeneralizedMatching{8}{1/2, 3/4,7/8}{2}{5/1,6/2}{5/4}, \hspace{\cbasisspacing}
			\end{gathered}}$ 
		\caption{In order, the rainbow element, $R_1$, and $R_2$. The coordinate of the rainbow element is 1. The coordinate of $R_1$ is $Q_{4}$. The coordinate of $R_2$ is $Q_4Q_3$. Generally, $R_i$ is the element with $i$ humps then a rainbow element, and has coordinate $Q_{n+r-1}...Q_{n+r-i}$.}
	\end{figure}

	Now define basis elements $E_i$ by $E_i(2i+1,2n+r):=R_i(2i+1,2n+r)$, $E_i(1,2i):=R$, the appropriate rainbow element. By the same argument as above, if $E_i$ has coordinate $c$ in $w$, $R_i$ has coordinate $Q_{i-1}...Q_{1}c$. One way to make this more clear is to consider intermediate basis elements $\sigma_j^{E_i}$ defined by $\sigma_j^{E_i}(2i+1,2n+r):=E_i(2i+1,2n+r)$ and $\sigma_j^{E_i}(1,2i):=R_{L,j}$. Then the coordinates of $\sigma_j^{E_i}(2i+1,2n+r)$ in terms of the coordinate $c$ of $E_i$ are $Q_{i-1}...Q_{i-j}$, and $R_i=\sigma_{i-1}^{E_i}$.

	Since we assume $Q_i\not=0$ for $i< n+r$, this implies the coefficient of $E_i$ is $\frac{Q_{n+r-1}...Q_{n+r-i}}{Q_1...Q_{i-1}}=\Q^i_{n+r}=x_{E_i}$. In particular, returning to our desired basis element $\psi$, the coordinate of $E_{\lfloor a/2\rfloor}$ is $\Q^{\lfloor a/2\rfloor}_{n+r}=x_{E_{\lfloor a/2\rfloor}}$.

	Note that the above logic only uses proposition \ref{shifted rainbow coeffs}, which requires only that a sub-matching be a rainbow element. So, suppose some basis element $\sigma$ has sub-matching $\sigma(s,t)=R$ with $n'$ nodes and $r'$ anchors, and that the coordinate of $\sigma$ in $w$ is $c$. Then it follows that the basis element $\theta_i$ defined by $\theta_i(1,s-1):=\sigma(1,s-1)$, $\theta_i(t+1,2n+r):=\sigma(t+1,2n+r)$, and $\theta_i(s,t):=E_i$ has coefficient $\Q^i_{n'+r'}c$. In other words, defining $Y_\psi$ with respect to the sub-matching $\sigma(s,t)$, the operation of finding the coordinate of $E_i$ given the coordinate of $R=\sigma(s,t)$ commutes with the isomorphism to $Y_\psi$.  An example is given in figure \ref{kernel induct characterization example}.

	\vspace{2mm}
	The above technique specifies an algorithm for determining the coordinate of $\psi$.

	As a base case, for the zero element have the algorithm return 1.

	Suppose inductively that the algorithm returns the coordinate for any $\sigma\in M_{2n'+r'}^{r'}$, $2n'+r'<2n+r$, and that that coordinate is equal to the coordinate function $x_\sigma$. Also suppose that the algorithm commutes with any isomorphism defined by lemma \ref{sub-matching isomorphism}. These statements are clearly true for the base case.

	Given $\psi\in M_{2n+r}^r$, if $\psi(1)=a$, we may find the coordinate of $E_{\lfloor a/2\rfloor}$ as before. Note that this operation commutes with any isomorphism defined by lemma \ref{sub-matching isomorphism}. Me may define $Y_{E_{\lfloor a/2\rfloor}}$ with respect to the sub-matching $E_{\lfloor a/2\rfloor}(2,a-1)$. By the inductive hypothesis, we may apply the algorithm to this sub-matching and commute with the isomorphism with $Y_{E_{\lfloor a/2\rfloor}}$. In this way, we find that the coordinate of $\tilde{\psi}$ defined by $\tilde{\psi}(1,a):=\psi(1,a)$ and $\tilde{\psi}(a+1,2n+r)=R$ is $x_{\psi(2,a-1)}(q)\Q^{\lfloor a/2\rfloor}_{n+r}=x_{\tilde{\psi}}(q)$. Similarly, define $Y_{\tilde{\psi}}$ with respect to the sub-matching $\tilde{\psi}(a+1,2n+r)$, and commute the algorithm with the isomorphism. In the same way, we obtain that the coordinate of $\psi\in Y_{\tilde{\psi}}$ is $x_{\psi(2,a-1)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{\lfloor a/2\rfloor}_{n+r}=x_\psi(q)$ as desired.

	Note that we only added a single operation to the algorithm in the inductive step, which also commutes with any isomorphism defined by lemma \ref{sub-matching isomorphism}. Thus the inductive step holds and the proposition is proved.

	\begin{figure}
		\def\cbasisspacing{1mm}
		\begin{adjustbox}{width=\textwidth}
			\GeneralizedMatching{14}{3/12,4/11,5/10,6/9,7/8,13/14}{2}{1/1,2/2}{12/4}, \hspace{\cbasisspacing}
			\GeneralizedMatching{14}{1/2,3/4,5/6,9/12,10/11,13/14}{2}{7/1,8/2}{12/4}, \hspace{\cbasisspacing}
			\GeneralizedMatching{14}{1/6,2/5,3/4,9/12,10/11,13/14}{2}{7/1,8/2}{12/4}
		\end{adjustbox}

		\caption{The figure on the left has sub-matching $R$ ignoring the last two nodes. The middle figure has submatching $R_3$ ignoring the last two nodes. The figure on the right has sub-matching $E_3$ also ignoring the last two nodes. Since the last two nodes have the same structure for all elements, if the coordinate of the first element is $c$, the coordinate of the second is $Q_6Q_5Q_4c$, and the coordinate of the third is $\frac{Q_6Q_5Q_4}{Q_1Q_2}c$.}

		\label{kernel induct characterization example}
	\end{figure}

\end{proof}

\vspace{5mm}
The following few corollaries will help to simplify some later arguments.

\begin{corollary}
	Let $w\in\cap\text{ker}(1+T_i)$, $w\not=0$. Suppose $\psi(1,a)$ is a sub-matching with no anchors. Then:

	$$x_\psi=x_{\psi(1,a)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{a/2}_{n+r}$$

	\label{characterization generalization}
\end{corollary}

\begin{proof}
	Define $a_1=\psi(1)$, $a_i=\psi(a_{i-1}+1)$. Then for some $j$ we have $a_j=a$. If $j=1$, the statement is the same as the proposition. Suppose that the statement is true for any matching with $a_v=a$, $v<j$. Then the statement holds for the sub-matching $\psi(a_1+1,2n+r)$, and we have:
	\begin{align*}
	x_\psi(q)=&x_{\psi(1,a_1)}(q)x_{\psi(a_1+1,2n+r)}(q)\Q^{a_1/2}_{n+r}\\
	=&x_{\psi(1,a_1)}(q)x_{\psi(a_1+1,a)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{a_1/2}_{n+r}\Q_{n+r-a_1/2}^{a/2-a_1/2}\\
	=&x_{\psi(1,a_1)}(q)x_{\psi(a_1+1,a)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{a_1/2}_{n+r}\Q_{n+r-a_1/2}^{a/2-a_1/2}\prn{\frac{Q_{a/2-1}...Q_{a/2-a_1/2}}{Q_{a/2-1}...Q_{a/2-a_1/2}}}\\
	=&x_{\psi(1,a_1)}(q)x_{\psi(a_1+1,a)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{a_1/2}_{a/2}\Q^{a/2}_{n+r}\\
	=&x_{\psi(1,a)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{a/2}_{n+r}
	\end{align*}

\end{proof}

\vspace{5mm}
\begin{corollary}
	If $\psi\in M_{2n+r}^r$, then $x_\psi(q)\not=0$ if $e>n+r$.

	\label{coeff nonzero}
\end{corollary}

\begin{proof}
	For our base cases, if $2n+r=2$ all coefficients are 1, which is nonzero for any $q$.

	Assume the statement is true for all $2n'+r'<2n+r$. Given $\psi(1)=a$ we have $$x_\psi(q)=x_{\psi(2,a-1)}(q)x_{\psi(a+1,2n+r)}(q)\Q^{\lfloor a/2\rfloor}_{n+r}$$

	If $e>n+r$, non of the $Q_i$ term appearing in $\Q^{\lfloor a/2\rfloor}_{n+r}$ are zero, and $n'+r'<n+r<e$ for any of the sub-matchings that appear, so those coordinates are nonzero and the corollary holds.

\end{proof}

\vspace{5mm}
The proposition fully characterizes any possible kernel element when $Q_1...Q_{n+r-1}\not=0$. In particular, the following corollary holds:

\begin{corollary}
	When $Q_1..Q_{n+r-1}\not=0$ and the kernel is nontrivial, the kernel is one dimensional.

	\label{1-D}
\end{corollary}

This corollary follows from the fact that we may write the coordinate of any basis element as proportional to the coordinate of the rainbow basis element.

\subsubsection{Nontrivial Kernel}

To verify the kernel element, we will need to know exactly which basis elements are mapped to a specific basis element by a given $(1+T_i)$. The next two lemmas help address this question.

\begin{lemma}
	Take some basis element $\psi\in M_{2n+r}^r$.

	\begin{enumerate}[label={(\roman*)}]

		\item Suppose $\psi(a)=b$ for some $b>a+1$, and that $(1+T_i)\psi=(1+q)\psi$ for some $a<i<b-1$. Me then have a subrepresentation $\psi(a,b)$ and define $Y_\psi$ with respect to this subrepresentation. Then for all basis elements $\sigma$ such that $(1+T_i)\sigma=q^{1/2}\psi$, we have that $$\sigma\in Y_\psi$$.

		\item Suppose $\psi$ has some anchor at position $u$, and $(1+T_i)\psi=(1+q)\psi$ for some $i>u$, we again have a subrepresentation $\psi(u,2n+r)$ and define $Y_\psi$ with respect to this subrepresentation. Then for all basis elements $\sigma$ such that $(1+T_i)\sigma=q^{1/2}\psi$, we have that $\sigma\in Y_\psi$ again.

	\end{enumerate}
	\label{preimage under hump}
\end{lemma}

\begin{proof}
	This lemma follows from an observation I made in section \ref{two humps created}: a transposition can only create two arcs or an arc and an anchor.

	(i) If $\sigma\not\in Y_\psi$ either $\sigma(1,a-1)\not=\psi(1,a-1)$ or $\sigma(b+1,2n+r)\not=\psi(b+1,2n+r)$. Suppose it is the first case. Then for some $s,t\in [1,a-1]$, $s<t$, we have $\psi(s)=t$ and $\sigma(s)\not=t$. To have $(1+T_i)\sigma=q^{1/2}\psi$ we must have $\sigma(t)=i+1$, $\sigma(s)=i$. But then $\sigma(a)\not=b$ and $\sigma(a)\not=i$ or $i+1$, so $((1+T_i)\sigma)(a)\not=b$ and $(1+T_i)\sigma\not=q^{1/2}\psi$. The same argument proves the $\sigma(b+1,2n+r)\not=\psi(b+1,2n+r)$ case.

	(ii) An analogous argument proves the anchor case. Specifically, the anchor cannot exist at position $u$ and is not created by action of $(1+T_i)$ if $\sigma(s)=i$ and $\sigma(t)=i+1$.

\end{proof}

\vspace{5mm}
It is important to note that lemma \ref{preimage under hump} only references cases where a transposition acts under an arc or to the right of an anchor. An example is given in figure \ref{preimage under hump example}.

The next lemma characterizes cases where the transposition is not under any arcs and all anchors are to the right.

\begin{figure}[b]
	\[
	\begin{tabular}{l l l l}
	\GeneralizedAction{13}{1/4,2/3,6/7,8/13,9/10,11/12}{1}{5/1}{5/4}{9}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(1+q)}\\

	\GeneralizedAction{13}{1/4,2/3,6/7,8/9,10/13,11/12}{1}{5/1}{5/4}{9}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(q^{1/2})}\\

	\GeneralizedAction{13}{1/4,2/3,6/7,8/13,9/10,11/12}{1}{5/1}{5/4}{6}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(1+q)}\\

	\GeneralizedAction{13}{1/4,2/3,5/6,8/13,9/10,11/12}{1}{7/1}{5/4}{6}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(q^{1/2})}\\

	\end{tabular}
	\]

	\caption{In the first line we act under an arc, so if another element without that arc is sent to that element, it must fix the arc as shown in the second line. In the third line we act to the right of an anchor, so if another element without that anchor is sent to that element, it must fix the anchor as shown in the fourth line. 
	}
	\label{preimage under hump example}
\end{figure}

Essentially, this lemma states that the only elements sent to the same element are those which break at most one of the top level arcs to the left of the leftmost anchor, or that break the leftmost anchor. An illustration is given in figure \ref{preimage under nothing example}.

\vspace{5mm}
\begin{lemma}
	Take a basis element $\psi\in M_{2n+r}^r$. Suppose the leftmost anchor in $\psi$ is at index $b$, or let $b=2n+r+1$ if there is no anchor. Define $a_j$ such that $\psi(a_j)=a_{j-1}+1$ and $\psi(a_1)=1$ for all $j$ such that $a_j<b$.

	Suppose $(1+T_i)\psi=(1+q)\psi$ for some $i<b-1$ where $\nexists s,t$ such that $\psi(s)=t$ and $s<i,t>i+1$. Suppose there is some basis element $\sigma$ such that $(1+T_i)\sigma=q^{1/2}\psi$. Then:
	\\

	\begin{enumerate}[label={(\roman*)}]
		\item  $\psi(a_{j-1}+2,a_j-1)=\sigma(a_{j-1}+2,a_j-1)$ for all $j$.

		\item $\psi(b+1,2n+r)=\sigma(b+1,2n+r)$

		\item If $b$ is not an anchor in $\sigma$, $\psi(a_j)=\sigma(a_j)$ for all $j$ such that $a_j\not=i+1$.

		\item If $b$ is an anchor in $\sigma$, there exists exactly one value of $j$ such that $\sigma(a_j)\not=\psi(a_j)$ and $a_j\not=i+1$
	\end{enumerate}
	\label{preimage under nothing}
\end{lemma}

\begin{proof}

	(i) Suppose that, for some $j$ there exists $s,t\in [a_{j-1}+2,a_j-1]$ such that $\psi(s)=t$ but $\sigma(s)\not=t$. Then if $(1+T_i)\sigma=q^{1/2}\psi$ we must have $\sigma(i)=s$ or $t$ and $\sigma(i+1)=s$ or $t$. But, by definition, $i,i+1\not\in[a_{j-1}+1,a_j]$, so this implies $\sigma(a_j)\not=a_{j-1}+1,i,i+1$, so $((1+T_i)\sigma)(a_j)\not=a_{j-1}+1$ and $(1+T_i)\sigma\not=q^{1/2}\psi$. So (i) is proved.

	\vspace{5mm}
	(ii) The proof of (ii) is analogous to the proof of (i). Me cannot have $\psi(b+1,2n+r)\not=\sigma(b+1,2n+r)$ and $\psi(b+1,2n+r)=((1+T_i)\sigma)(b+1,2n+r)$ if $((1+T_i)\sigma)(b)=b$.

	\vspace{5mm}
	(iii) If $b$ is not an anchor in $\sigma$ and $(1+T_i)\sigma=q^{1/2}\psi$, we must have $i$ an anchor in $\sigma$, and $\sigma(i+1)=b$. No other nodes in $\sigma$ are changed, so this proves (iii).

	\vspace{5mm}
	(iiii) From (i)-(iii) we have that the only remaining matchings that can differ are the $(a_{j-1}+1,a_j)$ matchings. If one of them differs, by the same argument as before it must be fixed by the action of $(1+T_i)$, and no other nodes are changed, so (iiii) is proved.

\end{proof}

\begin{figure}[b]
	\[
	\begin{tabular}{l l l l}
	\GeneralizedAction{15}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{1}{11/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(1+q)}\\

	\GeneralizedAction{15}{1/8,2/3,4/5,6/7,9/10,12/15,13/14}{1}{11/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(q^{1/2})}\\

	\GeneralizedAction{15}{1/6,2/3,4/5,7/10,8/9,12/15,13/14}{1}{11/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(q^{1/2})}\\

	\GeneralizedAction{15}{1/6,2/3,4/5,8/11,9/10,12/15,13/14}{1}{7/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(q^{1/2})}\\

	\end{tabular}
	\]

	\caption{The action of $(1+T_7)$ fixes the first basis element. Shown are all the basis vectors sent to the same element by the same transposition. Note that in all of them nodes 2-5 and 12-15 are the same. This illustrates (i) and (ii) in lemma \ref{preimage under nothing}. Note that in the last case where the anchor is in a different place, 1,6 and 9,10 are still matched. This illustrates (iii). In the middle two cases where the anchor is in the same place, only one of 1,6 or 9,10 are not paired. This illustrates (iiii). 
	}
	\label{preimage under nothing example}
\end{figure}

\vspace{5mm}
Lastly, we will need a small combinatorial result.

\begin{lemma}
	Suppose $n>b\geq a>0$ and $e>n$. Then $$Q_{n-a}Q_b-Q_{n-b-1}Q_{a-1}=Q_{n}Q_{b-a}$$

	\label{Q lemma}
\end{lemma}

\begin{proof}
	If $b=1$, the only possibility for $a$ is 1, in which reduces to lemma \ref{Q alg}.

	Suppose the lemma is true for all $\tilde{b}<b+1$. Then for $a< b$ we have 
	\begin{align*}	
	Q_{n-a}Q_b -Q_{n-b-1}Q_{a-1}=&Q_{n}Q_{b-a}\\
	Q_1Q_{n-a}Q_b-Q_1Q_{n-b-1}Q_{a-1}=&Q_1Q_{n}Q_{b-a}
	\end{align*}

	from lemma \ref{Q alg}, we have	
	$$Q_{n-a}(Q_{b+1}+Q_{b-1})-(Q_{n-b}+Q_{n-b-2})Q_{a-1}=Q_{n}(Q_{b-a-1}+Q_{b-a+1})$$

	and from the inductive hypothesis we have

	$$Q_{n-a}Q_{b+1}-Q_{n-b}Q_{a-1}=Q_{n}Q_{b-a+1}$$

	as desired.

	For $a=b$ we have 
	\begin{align*}
	Q_{n-b}Q_b -Q_{n-b-1}Q_{b-1}=&Q_{n}\\
	Q_1Q_{n-b}Q_b-Q_1Q_{n-b-1}Q_{b-1}=&Q_1Q_{n}
	\end{align*}

	from lemma \ref{Q alg}, we have
	\begin{align*}
	Q_{n-b}(Q_{b+1}+Q_{b-1})-(Q_{n-b}+Q_{n-b-2})Q_{b-1}=&Q_1Q_{n}\\
	Q_{n-b}Q_{b+1}-Q_{n-b-2}Q_{b-1}=&Q_1Q_{n}\\
	\end{align*}

	as desired.

	For $a=b+1$, we continue:
	\begin{align*}
	Q_1Q_{n-b}Q_{b+1}-Q_1Q_{n-b-2}Q_{b-1}=&Q_1Q_1Q_{n}\\
	(Q_{n-b-1}+Q_{n-b+1})Q_{b+1}-Q_{n-b-2}(Q_{b}+Q_{b-2})=&(1+Q_2)Q_{n}\\
	\end{align*}

	So by the inductive hypothesis
	$$Q_{n-b-1}Q_{b+1}-Q_{n-b-2}Q_{b}=Q_{n}$$

	as desired, and the proof is finished by induction.

\end{proof}

\vspace{5mm}
Me are now ready to prove existence of a kernel element. To prove this, we will show that if $w\in M_{2n+r}^r$ is as characterized above, the coordinate of any basis element in $(1+T_i)w$ is zero. This will split into various cases related to the previous lemmas.

\vspace{5mm}
\begin{theorem}
	Suppose $e=n+r+1$. Then $\cap\text{ker}(1+T_i)\not=0$.
	\label{kernel existence}
\end{theorem}

\begin{proof}

	As a base case, when $2n'+r'\leq 2$, the representation is at most one dimensional. If the one basis element has only anchors, it is sent to zero by any $(1+T_i)$, and is in the kernel. If the single basis element is a single arc, it is sent to $(1+q)$ times itself, and we take $e=n+r+1=2$ so $1+q=0$ and the base case holds.

	\vspace{5mm}

	Assume inductively that the statement holds for all $M_{2n'+r'}^{r'}$ where $2n'+r'<2n+r$. Take $w$ as defined by proposition \ref{kernel characterization}.

	Given $\psi\in (1+T_i)M_{2n+r}^r$ let $E_\psi\subset M_{2n+r}^r$ be the pre-image of $\psi$ under the action of $(1+T_i)$. To prove $w$ is in the kernel, we must show the following: 
	\begin{equation}
	(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=0\text{ for all basis elements }\psi
	\label{weak kernel}
	\end{equation}
	Inductively, we assume this equation holds for basis elements in smaller representations $M_{2n'+r'}^{r'}$, but only for $q$ such that $e=n'+r'+1$. Clearly this is true in the base case. For the following proof we will need a slightly stronger inductive assumption. Take $\psi'\in M_{2n'+r'}^{r'}$, and suppose either that $\psi'(1)=2n'+r'$, and that $T_i\psi'=(1+q)\psi'$, $1<i<2n'+r'-1$, or that 1 is an anchor in $\psi$. Defining $E_{\psi'}$ as before, we assume

	\begin{equation}
	(1+q)x_{\psi'}+\sum_{\sigma\in E_{\psi'},\sigma\not=\psi'}q^{1/2}x_\sigma=0\text{ for any } q \text{ with }e>n'+r'
	\label{strong kernel}
	\end{equation}

	Note that \ref{strong kernel} does not apply in the base case. Our proof of the inductive step will be split into cases, and each case will only depend on sub-cases in which certain inductive hypotheses apply, so this will not lead to any problems.

	Before exploring the cases, let us formally define $E_\psi$ to be the pre-image of $\psi$ under the action of $(1+T_i)$, and $E_{\psi(a,b)}$ to be the pre-image of $\psi(a,b)$ under action of $(1+T_{i-a+1})$:

	\begin{enumerate}[label={case \arabic*:}]
		\item Suppose $\psi\in (1+T_i)M_{2n+r}^r$ for some $i$, and that $\exists s,t$ such that $s<i<t-1$, $s>1$ or $t<2n+r$, and $\psi(s)=t$. Also suppose the leftmost anchor is at some index $u>t$, or that there are no anchors. Then we have a sub-matching $\psi(s,t)$, and by lemma \ref{preimage under hump} $E_\psi\subset Y_\psi$. Then, using corollary \ref{characterization generalization}, the following equality holds:

		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)\\
		=&\prn{x_{\psi(1,s-1)}(q)\Q_{n+r}^{(s-1)/2}}\prn{(1+q)x_{\psi(s,2n+r)}(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_{\sigma(s,2n+r)}(q)}\\
		=&\prn{x_{\psi(1,s-1)}(q)\Q_{n+r}^{(s-1)/2}}\prn{x_{\psi(t+1,2n+r)}(q)\Q_{n+r-(s-1)/2}^{(t-s+1)/2}}\prn{(1+q)x_{\psi(s,t)}(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_{\sigma(s,t)}(q)}
		\end{align*}

		Me have that $e>j$ for any $Q_j$ term appearing in the equation above, and $e>n'+r'$ for any sub-matching coordinate appearing above, so by corollary \ref{coeff nonzero}:

		$$(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=0$$

		if and only if

		$$(1+q)x_{\psi(s,t)}(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_{\sigma(s,t)}(q)=0$$

		Note that $(\psi(s,t))(1)=t-s+1$. So by our inductive hypothesis (ii), we have

		$$(1+q)x_{\psi(s,t)}(q)+\sum_{\sigma\in E_{\psi(s,t)},\sigma\not=\psi(s,t)}q^{1/2}x_{\sigma}(q)=0$$

		By lemma \ref{sub-matching isomorphism}, if $\sigma\in Y_\psi$, $(1+T_i)\sigma=q^{1/2}\psi$ if and only if $(1+T_{i-s+1})\sigma(s,t)=q^{1/2}\psi(s,t)$, so the previous equation implies $$(1+q)x_{\psi(s,t)}(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_{\sigma(s,t)}(q)=0$$

		as desired, and this case is proved.

		\vspace{5mm}
		\item Again take $\psi\in (1+T_i)M_{2n+r}^r$ for some $i$, but suppose the leftmost anchor is at some position $u$ where $1<u<i$. Then, as before, we have a sub-matching $\psi(u,2n+r)$ and by lemma \ref{preimage under hump} $E_\psi\subset Y_\psi$.

		Note that both corollary \ref{characterization generalization} and our inductive hypothesis \ref{strong kernel} still apply in this case, where we consider a left anchor instead of a matching. This allows the exact same logic from the proof of the first case to prove this case.
		\\

		It is important to note that, for both case 1 and case 2, the inductive hypothesis depends only on cases in which \ref{strong kernel} holds. Thus, if we show these cases rely on valid base cases, case 1 and 2 follow. This will be done in case 4.

		\vspace{5mm}
		\item Suppose $\psi\in (1+T_i)M_{2n+r}^r$ for some $i$, the leftmost anchor is at a position $u>i+1$ or there are no anchors, and $\nexists s,t$ such that $\psi(s)=t$ and $s<i<t-1$. Lemma \ref{preimage under nothing} characterizes all $\sigma\in E_\psi$. Me would like to prove the following for arbitrary $q$ where $e>n+r$:

		$$(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=-q^{1/2}x_{\psi(1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r+1}^{(i+1)/2}$$

		\begin{figure}[b]
			\[
			\begin{tabular}{l l l l}
			\GeneralizedAction{7}{1/2,3/4,5/6}{1}{7/1}{3/4}{3}{1/2,3/4,5/6}{7/1}{(1+q)}\\

			\GeneralizedAction{7}{1/4,2/3,5/6}{1}{7/1}{3/4}{3}{1/2,3/4,5/6}{7/1}{(q^{1/2})}\\

			\GeneralizedAction{7}{1/2,3/6,4/5}{1}{7/1}{3/4}{3}{1/2,3/4,5/6}{7/1}{(q^{1/2})}\\

			\GeneralizedAction{7}{1/2,4/7,5/6}{1}{3/1}{3/4}{3}{1/2,3/4,5/6}{7/1}{(q^{1/2})}\\

			\end{tabular}
			\]

			\caption{The four elements sent to the first element by $(1+T_3)$ are listed. The coordinate of the first element is $Q_3Q_2Q_1$. The coordinate of the second is $Q_3Q_2$. The coordinate of the third is $Q_3Q_2$. The coordinate of the fourth is $Q_3$. Call the first element $\psi$. Then $x_{\psi(1,2)}=1$, $x_{\psi(5,7)}=Q_1$, so $-q^{1/2}x_{\psi(1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r+1}^{(i+1)/2}=-q^{1/2}Q_4Q_3$. Me also have $(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=(q+1)(Q_3Q_2Q_1)+q^{1/2}(Q_3Q_2+Q_3Q_2+Q_3)=-q^{1/2}(Q_3Q_2Q_1^2-2Q_3Q_2+Q_3)=-q^{1/2}Q_4Q_3$ as desired (one can verify the last equality by hand or simplify using lemma \ref{Q lemma}).}
			\label{case 3}
		\end{figure}

		See figure \ref{case 3} for an example of this equality. Note that if $e=n+r+1$, $Q_{n+r}$ is the only zero component in the right side of this equation, so proving this equation is sufficient to prove case three.
		\\

		Me will prove this equality through yet another inductive proof, this time inducting on the number of top level humps, including the leftmost anchor.

		Formally, as we have in earlier lemmas, we will define $a_j$ by $a_1:=\psi(1)$, $a_j:=\psi(a_{j-1}+1)$. Then define $b_\psi$ such that $a_{b_\psi}=u$ if there is an anchor or $a_{b_\psi}=2n+r$ otherwise. Me induct on $b_\psi$.
		\\

		If $b_\psi=1$, we must be in $M_2^0$ to be in case 3 (otherwise $s<i<t-1$ for some $s,t$ where $\psi(s)=t$), which is trivially satisfied. Thus the base case holds.
		\\

		Suppose for all basis elements $\sigma$ such that $b_\sigma<b_\psi$, the equality holds. Suppose $i\not=1$. Then $a_1<i$ and lemma \ref{preimage under nothing} gives that there is a unique $\upsilon\in E_\psi$ such that $\upsilon(1)\not=a_1$. Thus we have the following equality:

		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_\sigma(q)\\
		=&\prn{x_{\psi(1,a_1)}(q)\Q_{n+r}^{a_1/2}}\prn{(1+q)x_{\psi(a_1+1,2n+r)}(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_{\sigma(a_1+1,2n+r)}(q)}
		\end{align*}

		Define $Y_\psi$ with respect to the sub-matching $\psi(a_1+1,2n+r)$. Then $\sigma\in E_\psi$, $\sigma\not=\upsilon$ implies $\sigma\in Y_\psi$. By our inductive hypothesis, we have that 
		\begin{align*}
		&(1+q)x_{\psi(a_1+1,2n+r)}(q)+\sum_{\sigma\in E_{\psi(a_1+1,2n+r)},\sigma\not=\psi(a_1+1,2n+r)}q^{1/2}x_{\sigma}(q)\\
		=&-q^{1/2}x_{\psi(a_1+1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r-a_1/2+1}^{(i+1-a_1)/2}
		\end{align*}

		By lemma \ref{sub-matching isomorphism}, $\sigma\subset Y_\psi$, $\sigma\in E_\psi$ if and only if $\sigma(a_1+1,2n+r)\in E_{\psi(a_1+1,2n+r)}$. This implies:

		\begin{align*}
		&(1+q)x_{\psi(a_1+1,2n+r)}(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_{\sigma(a_1+1,2n+r)}(q)\\
		=&-q^{1/2}x_{\psi(a_1+1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r-a_1/2+1}^{(i+1-a_1)/2}
		\end{align*}

		So, combining with the aforementioned equality, we have 
		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_\sigma(q)\\
		=&\prn{x_{\psi(1,a_1)}(q)\Q_{n+r-a_1/2+1}^{(i+1-a_1)/2}}\prn{-q^{1/2}x_{\psi(a_1+1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r-a_1/2+1}^{(i+1-a_1)/2}}\\
		=&\prn{x_{\psi(1,a_1)}(q)\Q_{n+r-a_1/2+1}^{(i+1-a_1)/2}}\prn{-q^{1/2}x_{\psi(a_1+1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r-a_1/2+1}^{(i+1-a_1)/2}}\prn{\frac{Q_{(i-1)/2-1}...Q_{(i-1-a_1)/2}}{Q_{(i-1)/2-1}...Q_{(i-1-a_1)/2}}}\\
		=&-q^{1/2}x_{\psi(1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r+1}^{(i+1)/2}\prn{\frac{Q_{n+r-a_1/2}Q_{(i-1)/2}}{Q_{n+r}Q_{(i-1-a_1)/2}}}
		\end{align*}

		Separately, note that $\upsilon$ is defined by $\upsilon(2,a_1-1)=\psi(2,a_1-1)$, $\upsilon(a_1+1,i-1)=\psi(a_1+1,i-1)$, $\upsilon(i+2,2n+r)=\psi(i+2,2n+r)$, and $\upsilon(1)=i+1$, $\upsilon(a_1)=i$. Thus we may determine $x_\upsilon$, again utilizing corollary \ref{characterization generalization}:

		\begin{align*}
		x_\upsilon=&x_{\psi(i+2,2n+r)}x_{\upsilon(2,i)}\Q_{n+r}^{(i+1)/2}\\
		=&x_{\psi(i+2,2n+r)}\prn{x_{\psi(2,a_1-1)}x_{\upsilon(a_1,i)}\Q_{(i-1)/2}^{(a_1-2)/2}}\Q_{n+r}^{(i+1)/2}\\
		=&x_{\psi(i+2,2n+r)}\prn{x_{\psi(1,a_1)}x_{\psi(a_1+1,i-1)}\Q_{(i-1)/2}^{(a_1-2)/2}}\Q_{n+r}^{(i+1)/2}\\
		=&x_{\psi(i+2,2n+r)}x_{\psi(1,i-1)}\Q_{n+r+1}^{(i+1)/2}\prn{\frac{Q_{n+r-(i+1)/2}Q_{a_1/2-1}}{Q_{n+r}Q_{(i-1-a_1)/2}}}
		\end{align*}

		Adding this into our previous equation, we have:
		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)\\
		=&-q^{1/2}x_{\psi(1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r+1}^{(i+1)/2}\prn{\frac{Q_{n+r-a_1/2}Q_{(i-1)/2}-Q_{n+r-(i+1)/2}Q_{a_1/2-1}}{Q_{n+r}Q_{(i-1-a_1)/2}}}
		\end{align*}

		Applying lemma \ref{Q lemma} to the portion of the equation above in parenthesis, the above is equivalent to

		$$(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=-q^{1/2}x_{\psi(1,i-1)}(q)x_{\psi(i+2,2n+r)}(q)\Q_{n+r+1}^{(i+1)/2}$$

		as desired. Note that if $e\geq n+r+1$ the only term above that can be zero is $Q_{n+r}$ (by corollary \ref{coeff nonzero}). Thus we have proved the inductive step for the case where $i\not=1$.

		\vspace{5mm}
		If $i=1$, we instead look at the sub-matchings $\psi(1,a_{(b_\psi-1)})$, $\psi(a_{(b_\psi-1)}+1,2n+r)$. Again lemma \ref{preimage under nothing} gives that there is a unique $\upsilon\in E_\psi$ such that $\upsilon(a_{(b_\psi-1)}+1)\not=\psi(a_{(b_\psi-1)}+1)$. Taking $Y_\psi$ with respect to the sub-matching $\psi(a_{(b_\psi-1)}+1,a_{b_\psi})$ again we have that $\sigma\in E_\psi$, $\sigma\not=\upsilon$ implies $\sigma\in Y_\psi$. Thus, following the same logic as before, we arrive at the following equality:
		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_\sigma(q)\\
		=&\prn{x_{\psi(a_{(b_\psi-1)}+1,2n+r)}(q)\Q_{n+r}^{a_{(b_\psi-1)}/2}}\prn{-q^{1/2}x_{\psi(3,a_{(b_\psi-1)})}(q)Q_{a_{(b_\psi-1)}/2}}\\
		=&-q^{1/2}x_{\psi(3,2n+r)}\frac{Q_{n+r-1}Q_{a_{(b_\psi-1)}/2}}{Q_{a_{(b_\psi-1)}/2-1}}
		\end{align*}

		Again, we know the structure of $\upsilon$ from lemma \ref{preimage under nothing}. Suppose for now that $a_{b_\psi}$ is not an anchor, so it is $2n+r$. Then $\upsilon$ is defined by $\upsilon(3,a_{(b_\psi-1)})=\psi(3,a_{(b_\psi-1)})$, $\upsilon(a_{(b_\psi-1)}+2,2n+r-1)=\psi(a_{(b_\psi-1)}+2,2n+r-1)$, and $\upsilon(1)=2n+r$, $\upsilon(2)=a_{(b_\psi-1)}+1$. So we may again find $x_\upsilon$:
		\begin{align*}
		x_\upsilon=x_{\upsilon(2,2n+r-1)}=&x_{\psi(3,a_{(b_\psi-1)})}x_{\psi(a_{(b_\psi-1)}+2,2n+r-1)}\Q_{n+r-1}^{a_{(b_\psi-1)}/2}\\
		=&x_{\psi(3,a_{(b_\psi-1)})}x_{\psi(a_{(b_\psi-1)}+1,2n+r)}\Q_{n+r-1}^{a_{(b_\psi-1)}/2}\\
		=&x_{\psi(3,2n+r)}\frac{Q_{n-r-1-a_{(b_\psi-1)}/2}}{Q_{a_{(b_\psi-1)}/2-1}}
		\end{align*}

		Alternatively, if $a_{b_\psi}$ is an anchor, the definition of $\upsilon$ is now $\upsilon(3,a_{b_\psi}-1)=\psi(3,a_{b_\psi}-1)$, $\upsilon(a_{b_\psi}+1,2n+r)=\psi(a_{b_\psi}+1,2n+r)$, and $\upsilon(1)=1$, $\upsilon(2)=a_{b_\psi}$, so we have:

		$$x_\upsilon=x_{\upsilon(2,2n+r)}=x_{\psi(3,a_{(b_\psi-1)})}x_{\psi(a_{b_\psi}+1,2n+r)}\Q_{n+r-1}^{a_{(b_\psi-1)}/2}=x_{\psi(3,2n+r)}\frac{Q_{n-r-1-a_{(b_\psi-1)}/2}}{Q_{a_{(b_\psi-1)}/2-1}}$$

		so for our purposes $x_\upsilon$ is the same in either case.

		Incorporating into the above equation, we have:

		$$(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=$$

		$$-q^{1/2}x_{\psi(3,2n+r)}\frac{Q_{n+r-1}Q_{a_{(b_\psi-1)}/2}-Q_{n-r-1-a_{(b_\psi-1)}/2}}{Q_{a_{(b_\psi-1)}/2-1}}$$

		By lemma \ref{Q lemma}, this is simply $-q^{1/2}x_{\psi(3,2n+r)}Q_{n+r}$ as desired, and we have finished proving case 3.
		\\

		\item The only cases we have not yet dealt with are those where either $1$ is an anchor or $\psi(1)=2n+r$. These are those cases related to our inductive hypothesis (ii).

		To not be in case 1 or 2, we must have that there are no anchors between index 1 and $i$, and that there is no integer $s$ such that $1<s<i<\psi(s)-1$. It follows from the same argument that proved lemma \ref{preimage under hump} that there exists exactly one $\upsilon\in E_\psi$ such that $\upsilon(1)\not=\psi(1)$. Define $N$ to be $2n+r$ if 1 is an anchor, or $2n+r-1$ if 1 is not an anchor. Then, defining $Y_\psi$ with respect to the sub-matching $\psi(2,N)$, we have that $\sigma\in E_\psi, \sigma\not=\upsilon$ if and only if $\sigma(2,N)\in E_{\psi(2,N)}$. Note that for $E_{\psi(2,N)}$ we may apply the inductive hypothesis from case 3, so we have:
		\begin{align*}
		&(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi,\upsilon}q^{1/2}x_\sigma(q)\\
		=&(1+q)x_{\psi(2,N)}(q)+\sum_{\sigma\in E_{\psi(2,N)},\sigma\not=\psi(2,N)}q^{1/2}x_{\sigma}(q)\\
		=&-q^{1/2}x_{\psi(2,i-1)}x_{\psi(i+2,N)}\Q{n+r}^{i/2}
		\end{align*}

		As in case 3, we can also determine $x_\upsilon$. $\upsilon$ is defined by $\upsilon(2,i-1)=\psi(2,i-1)$, $\upsilon(i+2,N)=\psi(i+2,N)$, $\upsilon(1)=i$, and $\upsilon(i+1)=2n+r$ if 1 is not an anchor or $i+1$ if 1 is an anchor, and we have:

		$$x_\upsilon=x_{\psi(2,i-1)}x_{\upsilon(i+2,N)}\Q{n+r}^{i/2}$$

		Thus we have

		$$(1+q)x_\psi(q)+\sum_{\sigma\in E_\psi,\sigma\not=\psi}q^{1/2}x_\sigma(q)=$$

		$$-q^{1/2}x_{\psi(2,i-1)}x_{\psi(i+2,N)}\Q{n+r}^{i/2}\prn{1-1}=0$$

		as desired, and the last case is proved. Note that this only relies on the inductive hypothesis from case 3, for which we showed the base case holds.
	\end{enumerate}

	Thus our inductive hypotheses have all been proven, and those that apply in the base case hold in the base case, so by induction the theorem is proved.

\end{proof}

\begin{corollary}
	If $e=n+r+1$, $M_{2n+r}^r$ is reducible, and has a unique sign subrepresentation.
\end{corollary}

\section{Fibonacci representations and quotients of Specht modules}\label{Fibonacci Section}
We have difficulty characterizing the Specht module above when it is irreducible.
We may instead attempt to characterize the irreducible quotient $D^\lambda$ of the Specht module.
In particular, note that, whenever $e \leq r + 2$, the restriction $\Res D^{(n+r,n)}$ decomposes into a direct sum of partitions $(m+s,m)$ such that $e \leq s + 2$ and $2m + s = 2n + r - 1$.
This forms a recurrence among these representations, allowing a combinatorial description of their dimension.
In this section we henceforth assume $e = 5$ and note that this recurrence resembles the Fibonacci recurrence;
we follow this to characterize the restriction $D^{(n+r,n)}$ with $r \leq 3$.

\begin{remark}
  Let $d_{m}^{0,3}$ be the dimension $\dim D^{(n,n)}$ with $2n = m$ when $m$ is even, and $\dim D^{(n+3,n)}$ with $2n + 3 = m$ when $m$ is odd.
  Similarly define $d_m^{1,2}$.
  Then, Corollary \ref{D Restrictions} gives
  \begin{align*}
    d_m^{0,3} &= d_{m-1}^{1,2}\\
    d_m^{1,2} &= d_{m-1}^{1,2} + d_{m-1}^{0,3}\\
    &= d_{m-1}^{0,3} + d_{m-2}^{0,3}.
  \end{align*}
  Carefully following this and noting the base cases $d_2^{0,3} = d_2^{1,2} = 1$, one may note that this recurrences proves that $d_m^{1,2} = d_{m+1}^{0,3} = f_n$, where $f_n$ is the $n$th \emph{Fibonacci number}.
  This matches the dimension our Fibonacci subrepresentations, motivating their definition.\qed
\end{remark}

We can start our study of $V$ by studying low-dimensional cases.
First, note that $V_{\vs\vp}^2$ is the sign representation $D^{(1^2)}$ and $V_{\vs\vp}^2$ is the trivial representation $D^{(2)}$.

$V_{\vp\vp}^2$ is a 2-dimensional representation of a semisimple commutative algebra, and hence decomposes into a direct sum of two subrepresentations.
In particular, we can use the basis $\cbr{(\vp\vs\vp),(\vp\vp\vp)}$ and explicitly write the matrix
\[
  \rho_{T_1} = \begin{bmatrix}
    \varc & \vard\\
    \vard & \vare
  \end{bmatrix}
\]
having characteristic polynomial $(\varc - \lambda)(\vare - \lambda) - \vard^2 = \lambda^2 - (\varc + \vare)\lambda +(\varc\vare - \vard^2)$.
We may verify that, for $\lambda = -1$, this evaluates to
\[
  -((-1 + q + q^2) (1 + q^3 + q^4 + q^5 + 2 q^6 + q^7))\brk{5}_q = 0
\]
and for $\lambda = q$ this evaluates to 
\[
-(q^2 (-1 + q + q^2) (1 + q + q^2 + q^3 + 2 q^4 + q^5)) \brk{5}_q = 0
\]
hence $\rho_{T_1}$ has eigenvalues $-1$ and $q$.

The eigenspaces with eigenvalues $-1$ and $q$ are subrepresentations isomorphic to the sign and trivial representation, hence $V_{\vp\vp}$ is isomorphic to a direct sum of the trivial and sign representations: $V^2_{\vp\vp} \simeq V^2_{\vs\vp} \oplus V^2_{\vs\vs}$.

Now we may prove that $V^3_{\vs\vp}$ is irreducible;
this has basis $\cbr{\prn{\vs\vp\vs\vp}, \prn{\vs\vp\vp\vp}}$, and the following matrices:
\[
  \rho_{T_1} = \begin{bmatrix}
    \varb & 0\\
    0 & \vara
  \end{bmatrix}; \hspace{20pt}
  \rho_{T_2} = \begin{bmatrix}
    \varc & \vard\\
    \vard & \vare
  \end{bmatrix}.
\]
A proper nontrivial subrepresentation of $V_{\vs\vp}^3$ must be one-dimensional, and hence an eigenspace of each of these matrices;
since $\varb \neq \vara$, the first has eigenspaces given by the spans of basis elements, and since $\vard \neq 0$, these are not eigenspaces of the second.
Hence $V^3_{\vs\vp}$ is irreducible.

These establish the low-dimensional behavior that we will use as base cases below.
The rest of this section will proceed first by proving that $V_{\vs\vp}$ and $V_{\vs\vs}$ are irreducible;
then, we will use combinatorial arguments to prove that these are isomorphic to the correct two-row Specht module quotients.

\begin{proposition}
  The representation $V_{\vs\vp} := V^m_{\vs\vp}$ is irreducible.
\end{proposition}
\begin{proof}
  We will prove this inductively in $m$.
  We've already proven it for $V^2_{\vs\vp}$ and $V^3_{\vs\vp}$, so suppose that $V^{m-2}_{\vs\vp}$ is irreducible.

  Let $\cbr{v_i}$ be the basis for $V_{\vs\vp}$.
  Then, each $v_i$ is cyclic; indeed, we can transform every basis vector into $(\vs\vp \dots \vp)$ by multiplying by the appropriate $\frac{1}{\vard-\varc}(T_i - \varc)$, and we can transform $(\vs\vp \dots \vp)$ into any basis vector by multiplying be the appropriate $\frac{1}{\vard-\vare}(T_i - \vare)$.
  Hence it is sufficient to show that each $v \in V_{\vs\vp}$ generate some basis element.

  Let $v'$ be the basis element $(\vs\vp\vs\vp\dots \vp)$, which is many copies of $\vs\vp$, followed by an extra $\vp$ if $m$ is odd.
  We will show that each $v \in F$ generates $v'$.

  Suppose that no elements beginning $(\vs\vp\vs\vp)$ are represented in $v_i$;
  then, all such elements are represented in $T_3v$, so we may assume that at least one is represented in $v$.

  Note that $\ima (T_2 - \vara) = \Span\cbr{\text{Basis vectors beginning }(\vs\vp\vs\vp)}$ and $(T_2 - \vara)v \neq 0$.
  Further, note that $\Res_{\SH(S_{m-2})}^{\SH(S_m)} \ima (T_2 - \vara) \simeq V^{m-2}_{\vs\vp}$ as representations.
  Hence irreducibility of $V^{m-2}_{\vs\vp}$ implies that $v'$ is generated by $(T_2 - \vara)v$, and $V^m_{\vs\vp}$ is irreducible.
\end{proof}

Now, we may begin considering restrictions:
\begin{lemma}
  The following branching rules hold:
  \begin{align*}
    V^{m-1}_{\vp\vp} \simeq \Res V_{\vs\vp}^m &\simeq V^{m-1}_{\vs\vs} \oplus V^{m-1}_{\vs\vp}\\
    \Res V_{\vs\vs}^m &\simeq V_{\vs\vp}^{m-1}.
  \end{align*}
\end{lemma}
\begin{proof}
  The first line follows by considering the last two $m-2$ transpositions for the left isomorphism, then the first two for the right isomorphism.
  This is well-behaved by Appendix \ref{Algebra}.

  Similarly, the second line follows by considering the last $m-2$ transpositions.
\end{proof}
This immediately gives a rather strong characterization of $V$.
\begin{corollary}
  The representation $V_{\vs\vs}$ is irreducible.\qed
\end{corollary}
\begin{corollary}
  The representation $V$ decomposes into a direct sum of irreducible representations as follows:
  \[
    V \simeq 3V_{\vs\vp} \oplus 2V_{\vs\vs}.
  \]\qed
\end{corollary}

Now we may use these in order to apply Young Tableau to characterize $V$.
\begin{theorem}\label{Fibonacci Theorem}
  The irreducible components of $V$ are given by the following isomorphisms:
     \begin{align*}  
      V^{2n}_{\vs\vs} &\simeq D^{(n,n)}\\ 
      V^{2n-1}_{\vs\vs} &\simeq D^{(n+1,n-2)}\\
      V^{2n}_{\vs\vp} &\simeq D^{(n+1,n-1)}\\
      V^{2n-1}_{\vs\vp} &\simeq D^{(n,n-1)}.
     \end{align*}
\end{theorem}
\begin{proof}
  We will prove this by induction on $n$;
  we have already proven the base case $V^{2}$, so suppose that we have proven these isomorphisms for $V^{2n-2}$.
  We will prove the isomorphisms for $V^{2n-1}$ and $V^{2n}$.

  By irreducibility, $V^{2n-1}_{\vs\vs} \simeq D^{\lambda}$ and $V^{2n-1}_{\vs\vp} \simeq D^{\mu}$ for some diagrams $\lambda$ and $\mu$.
  We will show that $\lambda = (n+1,n-2)$ and $\mu = (n+1,n-1)$.

  First, note that we have \[\Res \; D^{\lambda} \simeq D^{(n,n-2)} \simeq \Res \; D^{(n+1,n-2)}\] and \[\Res \; D^{\mu} \simeq D^{(n,n-2)} \oplus D^{(n-1,n-1)} \simeq \Res \; D^{(n,n-1)}.\]

  By irreducibility of $\Res D^{\lambda}$, the only normal number in $\lambda$ is 1.
  Further, the only tableaux which can remove a cell to yield $D^{(n,n-2)}$ are $(n+1,n-2)$, $(n,n-1)$, and $(n,n-2,1)$ as illustrated in Figure \ref{OddRes};
  we have already seen that $D^{(n,n-1)}$ does not have irreducible restriction, so we are left with $(n+1,n-2)$ and $\varsigma = (n,n-2,1)$.
  We may directly check that $\varsigma$ doesn't satisfy this, as we have the following:
  \begin{align*} 
    \beta_\varsigma(1,2) &= 3 - 2 + (n-2) = n-1\\
    \beta_\varsigma(1,3) &= 3 - 1 + n = n+2\\
    \beta_\varsigma(2,3) &= 2 - 1 + 3 = 4.
   \end{align*} 
  At least one of $\beta_\varsigma(1,2)$ and $\beta_\varsigma(1,3)$ is nonzero, since $\beta_\varsigma(1,3) - \beta_\varsigma(1,2) = 3 \not\equiv 0 \pmod e$, and hence at least one of $M_2$ and $M_3$ is empty.
  Hence at least one of 2 or 3 is normal in $\varsigma$, and $\lambda = (n+1,n-2)$.

  For $\mu$, we immediately see from Figure \ref{OddRes} that the only option is $(n,n-1)$.
  \begin{figure}
  \[
      \cbr{
        \begin{gathered}
        \ydiagram{6,3}. \; \;
        \ydiagram{5,4}, \; \;
        \ydiagram{5,3,1}
      \end{gathered} 
    }\longrightarrow \begin{gathered}
      \ydiagram{5,3}
  \end{gathered}\]
      \[\cbr{
        \begin{gathered}
          \ydiagram{5,4}
        \end{gathered}
      }\longrightarrow
      \begin{gathered}
        \ydiagram{5,3},\;\;
        \ydiagram{4,4}.
      \end{gathered}
    \]
      \caption{
      Illustration of the partitions of $9$ which can, via row removal, yield $(n,n-2)$ alone, or both $(n,n-2)$ and $(n-1,n-1)$.
    }\label{OddRes}
\end{figure}

  We can perform a similar argument for the $V^{2n}$ case, finding now that 
  \[\Res \; D^{\lambda'} \simeq D^{(n,n-1)} \simeq \Res \; D^{(n,n)}\] and 
  \[\Res \; D^{\mu'} \simeq D^{(n,n-1)} \oplus D^{(n+1,n-2)} \simeq \Res \; D^{(n+1,n-1)}.\]

  Through a similar process, we see that $\mu' = (n+1,n-1)$.
  We narrow down $\lambda'$ to one of $(n,n)$ or $\varpi := (n,n-1,1)$, and note that
  \begin{align*} 
    \beta_\varpi(1,2) &= 3 - 2 + (n-1) = n\\
    \beta_\varpi(1,3) &= 3 - 1 + n = n+2\\
    \beta_\varpi(2,3) &= 2 - 1 + 2 = 3
  \end{align*}
  and hence at least one of 2 or 3 is normal, $\Res D^{\varpi}$ is not irreducible, and $\lambda' = (n,n)$, finishing our proof.
\end{proof}
\begin{corollary}
  We have the following isomorphisms of representations:
  \begin{align*} 
    V^{2n} &\simeq 3D^{(n+1,n-1)} \oplus 2D^{(n,n)}\\
    V^{2n - 1} &\simeq 3D^{(n,n-1)} \oplus 2D^{(n+1,n-2)}
  \end{align*}
\end{corollary}

\section{Conjecture}\label{Conjecture Section}
Recall that $K_{2n + r}^r := K$ is the direct sum of all copies of the sign representation in $M$.
Hence the following characterises sign subrepresentations of $M$ completely:
\begin{proposition}
  $K \subset M_{2n + r}^r$ is trivial when $e \neq n + r + 1$, and $\dim \, K = 1$ when $e = n + r + 1$.\qed{\color{red} We know this is not accurate anymore, right?}
\end{proposition} 
\begin{proposition}
  Suppose $e < n + r + 1$, and suppose $n'$ is such that $e = n' + r + 1$.
  Note that $h := (1 + T_1)(1 + T_3)\dots(1 + T_{n - n'})$ maps $M_{2n + r}^r$ onto $M_{2n' + r}^r$.
  Then, the preimage $h^{-1}(K_{2n + r}^r)$ is a subrepresentation of $M_{2n + r}^r$, and the series
  \[
    0 \subset h^{-1}(K_{2n + r}^r) \subset M_{2n + r}^r
  \]
  is a composition series of $M_{2n + r}^r$.\qed
\end{proposition}
\begin{proposition}
  Denote the composition factor $M_{2n + r}^r / h^{-1}(K_{2n + r}^r)$ by $U_{2n + r}^r$.
  Then, there exist some naturals $m,s$ satisfying $2m + s = 2n + r$ and $m + s > n + r$ such that the following is an isomorphism of $\SH$-modules \[h^{-1}\prn{K_{2n + r}^r} \simeq U_{2m + s}^s\].
  \qed
\end{proposition}
\begin{proposition}
  For the same $m,s$ as above, we have the following composition series of Specht modules:
  \[
    0 \longrightarrow D^{(m+s,m)} \longrightarrow S^{(n+r,n)} \longrightarrow D^{(n+r,n)} \longrightarrow 0.
  \]
\end{proposition}
\begin{proposition}
  $M_{2n + r}^r \simeq S^{(n+r,n)}$ and $U_{2m +s}^s \simeq D^{(m+s,m)}$.
\end{proposition}

\newpage
\appendix
\section{Compatibility of Representations with the Relations}
In general, we defined the representations $V := V^{2n + r}$ and $M := M_{2n + r}$ for the free algebra on generators $\cbr{T_1,\dots,T_{2n + r - 1}}$.
Recall that we may give a presentation of $\SH$ having generators $T_i$ and relations
\begin{align}
  (T_i - q)(T_i + 1) &= 0 \label{quadratic}\\
  T_iT_{i+1}T_i &= T_{i+1}T_iT_{i+1} \label{braid1}\\ 
  T_iT_j &= T_jT_i \hspace{40pt} \abs{i - j} > 1. \label{braid2}
\end{align}
We call \eqref{quadratic} the \emph{quadratic relation} and \eqref{braid1}, \eqref{braid2} the \emph{braid relations}.
It is easily seen that a representation of $\SH$ is equivalent to a representation of the free algebra $k\langle T_i \rangle$ which are compatible with the relations.
We will prove in the following sections that $V$ and $M$ are compatible with the Hecke algebra relations.

\subsection{Explicit Definition of $M$}
We will give a more careful definition of the crossingless matchings representation here.
\begin{definition}
  A \emph{crossingless matching on $2n+r$ indices with $r$ anchors} is a partition of $\cbr{1,\dots,2n+r}$ into $n$ parts of size $2$ and $r$ of size 1 such that no two parts $(a,a')$ and $(b,b')$ satisfy $a < b < a' < b'$, and no parts $(c), (a,a')$ satisfy $a < c < a'$.
  We will call these arcs and anchors, respectively.
  Then, define $M^r_{2n+r}$ to be the $k$-vector space with basis the set of crossingless matchings on $2n+r$ indices with $r$ anchors.

  In order to endow $M^r_{2n + r}$ with an $\SH$-action, consider some basis element $w_j$ and some element $(1 + T_i)$ of $\SH$.
  The elements $\cbr{1} \cup \cbr{1 + T_i | 1 \leq i < 2n + r}$ generate $\SH$, so it is sufficient to define the action of $1 + T_i$ on $w_j$.

  If $w_j$ has arc $(i,i+1)$, define $(1 + T_i)w_j := (1 + q)w_j$.
  If $w_j$ has anchors $W(i) = i$ and $W(i+1) = i+1$, define $(1 + T_i)w_j := 0$.
  If $w_j$ has anchor $W(i) = i$ and arc $W(i+1) = b$, define $(1 + T_i)w_j := q^{1/2}w_l$, where $w_l(i) = i+1$, $w_l(b) = b$, and all other arcs agree with $w_j$.
  If $w_j$ has arcs $W(i) = a$ and $W(i+1) = b$, then define $(1 + T_i)w_j := q^{1/2}w_l$, where $w_l(i) = i+1$,$w(a) = b$, and all other acts agree with $w_j$. 
  We verify that this is well-defined in Section \ref{Cross Relations}.
\end{definition}

We may alternately sharpen our topological definition;
\begin{definition}
Fix $2n + r$ distinct points $a_1,\dots,a_{2n + r}$ points along $\RR \times \cbr{0} \subset \RR^2$ and $r$ distinct points $b_1,\dots,b_r$ along $\RR \times \cbr{1}$.
Then, define $M_{2n + r}^r$ to have basis given by the isotopy classes of $n + r$ paths connecting the points $a_1,\dots,a_{2n+r},b_1,\dots,b_r$ such that no distinct $b_i,b_j$ are connected by a path.

We will take some basis element $w_j \in M_{2n +r}^r$ and define the action $(1 + T_i)w_j$.
To do so, map $w_j$ through the natural embedding $\RR \times \brk{0,1} \hookrightarrow \RR \times \brk{\frac{1}{2},1}$, and form the figure $w_j^i$ by adjoining the lines connecting $a_l$ and $a_l + \prn{0,\frac{1}{2}}$ for all $l \neq i,i+1$ as well as paths from $a_i$ to $a_{i+1}$ and $a_i + \prn{0,\frac{1}{2}}$ to $a_{i+1} + \prn{0,\frac{1}{2}}$.
This has either 0 or 1 path components which do not intersect $\RR \times \cbr{0,1}$;
these form ``loops.''

Take the figure $\tilde w_j^i$ without this component.
If $\tilde w_j^i$ is not isotopic to some $w_l$, then define $(1 + T_i)w_j := 0$.
If $\tilde w_j^i$ is isotopic to some $w_l$, define $(1 + T_i)w_j := (1 + q)w_l$ if $w_j^i$ has a loop and $(1 + T_i)w_j := q^{1/2}w_l$ otherwise.
This process is illustrated in Figure \ref{Action}.
\end{definition}

Let the length of an arc $(i,j)$ be $l(i,j) := j - i + 1$.
Note that the crossingless matchings on $2n$ indices with no anchors can all be identified with a list of $n$ integers describing the lengths of the arcs from left to right;
  This basis is illustrated for $M_{5}^1$ in Figure \ref{S5 Basis}.

\begin{remark}
  This definition gives a graphical calculus for working with our module.
  It should be clear that, if $w_j^i$ has a loop then $w_l(i) = i+1$ and $w_l = w_j$.
  Further, this easily defines an arbitrary composition:
  \[
    (1 + T_{i_1})\cdots(1 + T_{i_\ell})w_j = q^{(\ell - t)\frac{1}{2}}(1 + q)^tw_l
  \]
  if the figure we make via $(1+T_{i_1})\cdots(1+T_{i_\ell})$ is isotopic to $w_l$ after removing $t$ loops.
\end{remark}

  Note that the crossingless matchings on $2n$ indices with no anchors can all be identified with a list of $n$ integers describing the lengths of the arcs from left to right;
  call this the \emph{increasing lexicographical order basis}.
  Further, we have a surjection $M_{2n + 2r}^0 \twoheadrightarrow M_{2n + r}^r$ which takes basis elements to basis elements;
  this induces an order on the basis for $M_{2n +r}^r$, which we will henceforth refer to as the \emph{induced lexicographical order basis.}
  This basis is illustrated for $M_{5}^1$ in Figure \ref{S5 Basis}.

\subsection{Compatibility for the Crossingless Matchings Representations}
\label{Cross Relations}
We verify the relations on the crossingless matchings representation $M$.
Take some basis vector $w_i \in M$.
We will first check \eqref{quadratic} by case work:
\begin{itemize}
  \item Suppose there is an arc $(i,i+1)$.
    Then, \[(T_i-q)(T_i + 1)w = (1 + q)\prn{(1 + T_i)w - (1 + q)w} = 0,\] giving \eqref{quadratic}.
  \item Suppose there is no arc $(i,i+1)$ and indices $i,i+1$ do not both have anchors;
    then $(T_i +  1)w = q^{1/2}w'$ for some basis vector $w'$ having arc $(i,i+1)$, and the computation follows as above for \eqref{quadratic}.
  \item Suppose $i,i+1$ are anchors;
    then $(T_i + 1)w = 0$, giving \eqref{quadratic}.
\end{itemize}

\vspace{5pt}
Now we verify \eqref{braid1}.
Let $h := (1 + T_i)(1 + T_{i+1})(1+T_i)$, and let $g := (1 + T_{i+1})(1 + T_i)(1 + T_{i+1})$.
Note the following expansion:  
  \begin{align*}
      hw
      &= 1 + 2T_i + T_i^2 + T_{i+1} + T_iT_{i+1} + T_{i+1}T_i + T_iT_{i+1}T_i\\
      &= 1 + (1+q)T_i + T_{i+1} + T_iT_{i+1} + T_{i+1}T_i + T_iT_{i+1}T_i.
    \end{align*}
    This equality, with $i$ and $i+1$ interchanged, holds for $g$.
    Hence we have
    \[
      (h-g)w = q(T_i - T_{i+1}) + T_iT_{i+1}T_i - T_{i+1}T_iT_{i+1}.
    \]
    Hence we may equivalently check that $(h-g)w = q(T_i - T_{i+1})$.
    In fact, $hw = q(1 + T_i)$ and $gw = q(1 + T_{i+1})$ by Figure \ref{braid1arc}, giving compatibility.
    \begin{figure}
\[
  \begin{gathered}\EmptyNAction{5}{2/1,3/2,2/3}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\EmptyNAction{5}{2/1}\end{gathered}
  \hspace{30pt}
  \begin{gathered}\EmptyNAction{5}{3/1,2/2,3/3}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\EmptyNAction{5}{3/1}\end{gathered}
\]
  \caption{
    The above give visual intuition for isotopies giving rise to equalities between $(1 + T_i)(1 + T_{i-1})(1 + T_i)w_j$ and $q(1 + T_i)$, and between $(1 + T_{i+1})(1 + T_i)(1 + T_{i+1})w_j$ and $q(1 + T_i)$.
  }
  \label{braid1arc}
\end{figure}

\begin{figure}
\[  
  \begin{gathered}\EmptyNAction{7}{2/1,5/2}\end{gathered}
  \longsquigrightarrow
  \begin{gathered}\EmptyNAction{7}{5/1,2/2}\end{gathered}
\]
  \caption{
    The above give visual intuition for isotopies giving rise to equalities between $(1 + T_i)(1+T_j)w_l$ and $(1 + T_j)(1 + T_i)w_l$.}
    \label{braid2arc}
\end{figure}

Lastly, we have the equation
\[
  (1 + T_i)(1 + T_j) - (1 + T_j)(1 + T_i) = T_iT_j - T_jT_i
\]
and hence we simply need to verify that $(1 + T_i)$ and $(1 + T_j)$ commute, as shown in Figure \ref{braid2arc}

\subsection{Compatibility for the Fibonacci Representations} 
\label{Fib Relations}
We verify the relations on the Fibonacci representation $V$.
Note that \eqref{braid2} follows easily from the ``local'' nature of $V$, and the others may be verified explicitly on strings of length 3 and 4.
By considering the coefficients in order of \eqref{Fib Action}, the quadratic relation \eqref{quadratic} gives the following polynomials in $q$:
\begin{equation}
  \begin{split}
    (\vara - q)(\vara + 1) &= 0,\\
    (\varb - q)(\varb + 1) &= 0,\\
    \varc\vard + \vard\vare &= (q-1)\vard,\\
    \varc^2 + \vard^2 &= (q -1)\varc + q,\\
    \vare^2 + \vard^2 &= (q -1)\vare + q
  \end{split}
\end{equation}
The first two of these are easily verified.
Since $\vard \neq 0$, the third is equivalently given by
\[
  (q - 1) = \varc + \vare = \tau(q\tau - 1 + q - \tau) = (\tau^2 + \tau)(q - 1)
\]
or that $\prn{\tau^2 + \tau - 1}(q-1) = 0$.
One may verify that \[\tau^2 + \tau - 1 = q^6 + 2q^5 + q^4 + q^3 + q^2 - 1 = (-1+q+q^2)\brk{5}_q = 0.\]

The fourth is given by the quadratic
\[
  \tau^2\brk{(q\tau-1)^2 - \tau(q+1)} = \tau(q-1)(q\tau - 1) + q
\]
or equivalently,
\[
  (\tau^2 + \tau - 1)\brk{q\prn{qt^2 + 1} + t} = 0.
\]

The fifth is similarly given by
\[
  (\tau^2 + \tau - 1)\brk{q\prn{qt + 1} + t^2} = 0. 
\]
All of these vanish for $e = 5$, giving compatibility with \eqref{quadratic}.

\vspace{7pt}
We now verify \eqref{braid1}.
We may order the basis for $V^4$ as follows:
\[
  \cbr{(\vp\vp\vp\vp),(\vs\vp\vp\vs),(\vp\vp\vp\vs),(\vs\vp\vp\vp),(\vs\vp\vs\vp),(\vp\vs\vp\vs),(\vp\vp\vs\vp),(\vp\vs\vp\vp)}.
\]
Then, in verifying the braid relation \eqref{braid1} in this order, we encounter the following quadratics (with tautologies and repetitions omitted):
\begin{align*}
    \vara\vare^2 + \varb\vard^2 &= \vara^2 \vare\\
    \vara\vard\vare + \varb\varc\vard &= \vara\varb\vard\\
    \varb\varc^2 + \vara\vard^2 &= \varb^2\varc\\
    \vara\varc^2 + \vard^2\vare &= \vara^2\varc\\
    \vard\vare^2 + \vara\varc\vard &= \vara\vard\vare
\end{align*}
Substituting in $\tau$ and dividing by $\delta$ whenever possible, these are equivalent to the vanishing of the following polynomials in $q$:
\begin{align*}
  -q (1 + q) (1 + q^2 + q^3) (2 + q + 3 q^2 + 2 q^3) \brk{5}_q &= 0\\
  (1 + 2 q + q^3 + q^4) \brk{5}_q &= 0\\
  (1+q)^2 (1+q^2+q^3) (1+3q^3 - q^4 + q^6)\brk{5}_q &= 0\\
  (1+q)^2 (1+q^2+q^3) (1+5q+5q^2+3q^3+3q^4+3q^5+q^6)\brk{5}_q &= 0\\
  (1+q) (1+q^2+q^3) (-1+2q+q^2+q^3+q^4)\brk{5}_q &= 0.
\end{align*}
Notably, each of these vanish when $e = 5$, giving compatibility with \eqref{braid1}.

\section{Restrictions to Conjugate Subalgebras}\label{Algebra}
Throughout the text, for some representation $V$, we refer to $\Res_{\SH(S_{l})}^{\SH(S_m)} V$ without specifying exactly which subalgebra $\SH(S_{l})$.
For instance, in section \ref{Fibonacci Section}, we explicitly state that the subrepresentations $V_{\vs\vp} \oplus V_{\vs\vs}$ and $V_{\vp\vp}$ are isomorphic because they both may be characterized by such a restriction.
We will verify that this is justified, using a more general fact about resctrictions to conjugate subalgebras.
\begin{proposition}
  Suppose $B,B'$ are subalgebras of a $k$-algebra $A$ with $B = uB'u^{-1}$ for some unit $u \in A^\times$, and let $V$ be a left $A$-module.
  Let $\phi:V \rightarrow V$ be the linear automorphism specified by $v \mapsto uv$.
  Then, the following commutes for any $b \in B$:
  \[
    \begin{tikzcd}
      V \arrow[r,"\phi"] \arrow[d,"b"] & V \arrow[d,"ubu^{-1}"]\\
      V \arrow[r,"\phi"] & V
    \end{tikzcd}
  \]
  Hence, through the identification of $B$ and $B'$ via conjugation by $u$, we have $\Res_{B}^A V \simeq \Res_{B'}^A V$
\end{proposition}
\begin{proof}
  It suffices to note that $(ubu^{-1})uv = ubv$.
\end{proof}

\begin{corollary}
  Suppose $\SH',\SH''$ are two subalgebras of $\SH(S_m)$ generated by $l$ simple reflections and $V$ is a representation of $\SH$.
  Then, $\Res_{\SH'}^{\SH} V \simeq \Res_{\SH''}^\SH V$.
\end{corollary}
\begin{proof}
  Let $\SH'$ and $\SH''$ be the subalgebras of $\SH(S_m)$ generated by the reflections $\cbr{T_{i_1},\dots,T_{i_l}}$ and $\cbr{T_{i_1},\dots,T_{i_{j-1}},T_{i_j + 1},T_{i_{j+1}},\dots,T_{i_l}}$ for $1 \leq i_1 < \dots < i_{j-1} < i_j + 1 < i_{j+1} < \dots < i_l \leq n$.
  It is sufficient to prove that $\SH'$ and $\SH''$ are conjugate;
  then transitivity gives conjugacy of any $S_l \subset S_m$, and the previous proposition gives isomorphisms of the representations.

  We will show that $\SH'' = T_{i_j}\SH'T_{i_j}^{-1}$.
  It suffices to show that $T_{i_j}T_wT_{i_j}^{-1} \in \SH''$ for $w$ a word generated by simple transpositions $s_{i_1},\dots,s_{i_{l}} \in S_m$.
  First, note that $l(w) < l(s_{i_j}w)$, implying $T_{i_j}T_w = T_{s_{i_j}w}$ by lemma 1.12 in \cite{Mathas-book}.
  Further, by the same lemma, we have
  \begin{align*}
    T_{s_{i_j}w}T_{i_j}^{-1} 
    &= q^{-1}\prn{T_{s_{i_j}w}T_{i_j} + (1-q)T_{s_{i_j}w}}\\
    &= q^{-1}\prn{T_{qs_{i_j}ws_{i_j}} + (q-1)T_{s_{i_j}w} + (1-q)T_{s_{i_j}w}}\\
    &= T_{s_{i_j}ws_{i_j}}
  \end{align*}
  which is in $\SH''$.
\end{proof}
\begin{figure}[h]
\iffalse
\def\tho{\Phi_3(q^{1/2})}
\def\tht{\brk{3}_{q^{1/2}}}
\def\to{q^{1/2}\brk{2}_{q^{1/2}}}
Define the following coefficients:
\begin{align*}
  \tho &:= q^{3/2} + 1\\
  \tht &:= q^{3/2} + q + q^{1/2} + 1\\
  \to &:= q + q^{1/2}
\end{align*}
\fi

\def\tho{q^{3/2} + 1}
\def\tht{\brk{4}_{q^{1/2}}}
\def\to{q^{1/2}\prn{q^{1/2} + 1}}
\begin{align*}
  \varphi_6^0, \varphi_5^1 &= \begin{bmatrix}
    0 & 0 & -\tho & 0 & 0\\
    0 & -\tho & \tht & 0 & 0\\
    0 & 0 & \tht & 0 & -\tho\\
    -\tht & \tht & \to & 0 & \tht\\
    \tht & 0 & \tht & -\tht & 0
  \end{bmatrix}\\
  \varphi_6^2 &= \begin{bmatrix}
    0 & 0 & - q - 1 & 0 & 0 & 0 & 0 & 0\\
    0 & -q - 1 & -q^{1/2} & 0 & 0 & 0 & 0 & 0\\
    0 & q + 1 & 0 & q^{1/2} & 0 & 0 & 0 & 0\\
    0 & 0 & -q^{1/2} & 0 & 0 & 0 & -q^2 - 1 & 0\\
    q^{1/2} & -q^{1/2} & -\brk{3}_{q^{1/2}} & 0 & 0 & 0 & -q^{1/2} & 0\\
   -q^{1/2} & q^{1/2} & 0 & \brk{3}_{q^{1/2}} &0 & 0 & 0 & q^{1/2}\\
   -q^{1/2} & 0 & -q^{1/2} & 0 & q^{1/2} & 0 & 0 & 0\\
    q^{1/2} & 0 & 0 & 0 & -q^{1/2} & -\brk{3}_{q^{1/2}} & 0 & -q^{1/2}
  \end{bmatrix}\\
  \varphi_5^3 &= \begin{bmatrix}
    -\to & \tht & q^{1/2}\\
    q^{1/2} & q^{3/2} & 0\\
    0 & -q - 1
  \end{bmatrix}  \hspace{30pt} 
  \varphi_4^0 = \begin{bmatrix}
    0 & -\tht\\
    -1 & 1
  \end{bmatrix}\\
  \varphi_3^1 &= \begin{bmatrix}
    0 & -\tht\\
    \to & -\to
  \end{bmatrix} \hspace{27pt}
  \varphi_4^2 = \begin{bmatrix}
    0 & \tht & 0\\
    1 & -1 & 0\\
    -1 & 0 & 1
  \end{bmatrix}
\end{align*}

\caption{We define representations $\varphi_{2n + r}^r:M_{2n + r}^r/K_{2n + r}^r \rightarrow V^{2n + r}_{s}$, where $s = \vs\vs$ if $r = 0,3$ and $s = \vs\vp$ otherwise.
These are given with respect to the basis on $M_{2n + r}^r$ induced by the increasing lexicographic order basis on $M_{2n + 2r}^0$ and the quotient $M_{2n + 2r}^0 \twoheadrightarrow M_{2n + r}^r$.
These are also given with respect to the basis on $V$ given by increasing lexicographic order $\vs < \vp$.
  }\label{Fib Data}
\end{figure}

\begin{figure}[h]
  \[
    \ENRGrid{13/7}{2/1,5/1,8/1,11/1,1/2,4/2,7/2,10/2}
  \]\[
    \ENRGrid{13/7}{3/1,7/1,11/1,2/2,6/2,10/2,1/3,5/3,9/3}
  \]
  \caption{
    Illustration of the modules $M_{2n + r}^r$ having sign submodules for $p = \infty$ and $e = 3,4$ respectively.
    The value $n + r + 1$ is filled in the squares, and modules having sign submodules are colored magenta.
    For $p = \infty$ and $2n + r \leq 14$, it has been verified, through a combination of theorems here and empirical computations, that $K_{2n + r}^r$ is nontrivial if and only if $e | n+r+1$ and $e < n$.
  }\label{Empirical Kernel}
\end{figure}

\def\qqo{\alpha}
\def\tto{\beta}
\def\vtv{\gamma}
\def\fft{\delta}
\def\fto{\varepsilon}
\setcounter{MaxMatrixCols}{20}

\begin{figure}[h]
\begin{align*} 
  \iota_{3,3}^1 = \iota_{3,4}^0 &=  
\begin{bmatrix}
  1 & 1
\end{bmatrix}^\intercal\\
\iota_{3,5}^1 = \iota_{3,6}^0&=  
\begin{bmatrix}
  1 & 0 & 0 & 0 & 0\\
  0 & 0 & 1 & 1 & 0\\
  0 & 1 & 0 & 1 & 0\\
  0 & 0 & 0 & 1 & 1
\end{bmatrix}^\intercal\\
\iota_{3,6}^4&=  
\begin{bmatrix}
  -1 & -1 & 0 & 1 & 1
\end{bmatrix}^\intercal\\
\iota_{3,7}^3&=  
\begin{bmatrix}
 1 & 0 &-1 &-1 & 1 & 0 & -1 & -1 & 0 & 0 & 0 & 0 & 1 & 1
\end{bmatrix}^\intercal\\
\iota_{3,7}^1&=  
\begin{bmatrix}
 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 &-1\\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1
\end{bmatrix}^\intercal
\end{align*}
\caption{The maps of representations $\iota_{e,2n+r}^r:U_{2n +r}^r \hookrightarrow M_{2n + r}^r$ for $e = 3$.}
\label{Composition series e=3}
\end{figure}

\begin{figure}[h]
\begin{align*}
  \iota_{4,4}^2&=  
\begin{bmatrix}
  1 & \qqo & 1
\end{bmatrix}^\intercal\\
\iota_{4,4}^1&=  
\begin{bmatrix}
  1 & \frac{1}{2}\qqo & \frac{1}{2}\qqo & 1 & \frac{1}{2}\qqo
\end{bmatrix}^\intercal\\
\iota_{4,6}^2 &=
\begin{bmatrix}
  1& 0& 0& 0& 0& 0& \qqo & 1& 0\\
  0& 1& 0& 0& 0& -1& -1& -\qqo &0\\
  0& 0& 1& 0& 0& \qqo& 0 &1 &0\\
  0& 0& 0& 1& 0& 0& -1& -\qqo &-1\\
  0& 0& 0& 0& 1& \frac{1}{2}\qqo& \frac{1}{2}\qqo &1 &\frac{1}{2}\qqo
\end{bmatrix}^\intercal\\
\iota_{4,6}^0 &=
\begin{bmatrix}
  \qqo & 1 & 1 & \qqo & 1
\end{bmatrix}^\intercal\\
\iota_{4,7}^1 &=
\begin{bmatrix}
1& 0& 0& 0& \frac{1}{2}\qqo& 0& -\frac{1}{2}& -\frac{1}{2}& 0& 0& 1& 1/2& \qqo& \frac{1}{2}\\
0& 1& 0& 0& -1& 0& \frac{1}{2}\qqo& \frac{1}{2}\qqo& 1& 0& \qqo& \frac{1}{2}\qqo& -1 & \frac{1}{2}\qqo\\
0& 0& 1& 0& -1& 0& 0& \qqo& 1& 0& -\qqo& 0& -1& 0\\
0& 0& 0& 1& \qqo& 0& 0& -1& -\qqo& 0& 1& -1& 0& 0\\
0& 0& 0& 0& 0& 1& \frac{1}{2}\qqo& \frac{1}{2}\qqo& 0& 0& 0& \frac{1}{2}\qqo& -1& \frac{1}{2}\qqo\\
0& 0& 0& 0& 0& 0& 0& 0& 0& 1& -\frac{1}{2}\qqo& \frac{1}{2}\qqo& 1& \frac{1}{2}\qqo
\end{bmatrix}^\intercal\\
\end{align*}
\caption{The maps of representations $\iota_{e,2n+r}^r:U_{2n +r}^r \hookrightarrow M_{2n + r}^r$ for $e = 4$.}
\label{Composition series e=4}
\end{figure}

\begin{figure}[h]
\begin{align*}
  \iota_{5,5}^3 &=
\begin{bmatrix}
  1 & \tto & \tto & 1
\end{bmatrix}^\intercal\\
\iota_{5,6}^2 &=
\begin{bmatrix}
  \tto& \tto& 1& 1& \tto+1& \tto & \tto & \tto & 1
\end{bmatrix}^\intercal\\
\iota_{5,7}^3 &=
\begin{bmatrix}
1& \tto& \tto& 1& 0& 0& 0& 0& 0& 0& 0& 0& 0& 0\\
0& 0& 0& 0& 1& \tto& \tto& 1& 0& 0& 0& 0& 0& 0\\
1& 0& 0& 0& 0& 0& 0& 0& \tto& \tto& 1& 0& 0& 0\\
\tto& \tto& 1& 0& 1& \tto+1& \tto& 0& \tto& \tto& 0& 1& 0& 0\\
-\tto-1& -\tto& -\tto& 0& -\tto& -\tto-1& -\tto-1& 0& -\tto-1& -\tto-1& 0& 0& 1& 0\\
0& 0& 1& 0& 0& 0& \tto& 0& 0& \tto& 0& 0& 0& 1
\end{bmatrix}^\intercal\\
\iota_{5,7}^1 &=
\begin{bmatrix}
1& \qqo& \qqo& \qqo& -\tto& \qqo& -\tto& \qqo& -\tto& 1& \qqo& \qqo& \qqo& -\tto
\end{bmatrix}^\intercal\\
\iota_{6,6}^4 &=
\begin{bmatrix}
  1 & \vtv & 2 & \vtv & 1 
\end{bmatrix}^\intercal\\
\iota_{6,7}^3 &=
\begin{bmatrix}
  \vtv & 2 & \vtv & 1 & 1 & 2\vtv & 3 & \vtv & 2 & 2\vtv & 2 & 2 & \vtv & 1
\end{bmatrix}^\intercal\\
\iota_{6,7}^5 &=
\begin{bmatrix}
  1 & \fft & \fto & \fto & \fft & 1
\end{bmatrix}^\intercal
\end{align*}
\caption{The maps of representations $\iota_{e,2n+r}^r:U_{2n +r}^r \hookrightarrow M_{2n + r}^r$ for $e = 5,6$.}
\label{Composition series e=6}
\end{figure}

\section{Data}\label{Empirics Section}
In this section we aim both to support conjecture with data and to provide transition matrices wherever possible between our two graphical representations.\cite{Github}

\begin{remark}
  These results are restricted by two things; first and with less recourse, this was computed in Magma, a closed source language with bugs that could not be resolved in certain cases.
  Second, the memory necessary to store the Hecke algebra via basis elements and structure constants grows with $(n!)^3$;
  this is prohibitively large when $m > 7$.
  An implementation as a quotient by a free algebra or only specifying multiplication by generators may fix this, however, this is not possible with the Magma language, and the closed source renders modification of the language impossible.
\end{remark}

We give in Figure \ref{Fib Data} some isomorphisms between $M/K$ and $V$ for $e \geq n + r + 1$ with $p = \infty$;
all but one of these are cases with $e > n + r + 1$, so $K=0$ and this is an isomorphism with our crossingless matchings representation.
The case $n = 1$, $r = 3$ gives an example of an isomorphism not proven in general, but proven via our computation.
All of these computations are done for $q$ a primitive 5th root of unity in the algebraic extension of the Cyclotomic field $\QQ(\zeta_{10})$ by a root of the polynomial $x^2 - \tau$.

We give in Figure \ref{Empirical Kernel} some data supporting a conjecture concerning sign subrepresentations of $M_{2n + r}^r$.
The computations to support this were done over $\CC$ with $q$ a primitive 5th root of unity.

It is known that, for small $2n + r$, each Specht module $S^{(n+r,n)}$ has a composition series of length 2.\cite{Mathas-book}
We give in Figures \ref{Composition series e=3} through \ref{Composition series e=6} the map $\iota_{e,2n+r}^r:U_{2n + r}^r \hookrightarrow M_{2n + r}^r \twoheadrightarrow U_{2n + r}^r$, which conjecturally illustrates the inclusion of the first composition factor of $S^{(n+r,n)}$ into $S^{(n+r,n)}$ for all $2n + r \leq 7$.

\bibliographystyle{numeric} 
 \bibliography{../RepBib}
\end{document}